\documentclass[paper=a6,pagesize=pdftex,9pt,twocolumn]{scrbook}
% Required for marginpar verse references
\usepackage[maxfloats=100]{morefloats}

\usepackage[french]{babel}
\usepackage[protrusion=true,final]{microtype}
\usepackage[newcommands]{ragged2e}

% For references
\usepackage{bibleref-french}
\biblerefstyle{colombefull}
%\biblerefstyle{text}
\addbiblebook{1 Ne.}{1 Nephi}%
\addbiblebook{2 Ne.}{2 Nephi}%
\addbiblebook{Jacob}{Jacob}%
\addbiblebook{Enos}{Enos}%
\addbiblebook{Jarom}{Jarom}%
\addbiblebook{Omni}{Omni}%
\addbiblebook{W of M}{Words of Mormon}%
\addbiblebook{Mosiah}{Mosiah}%
\addbiblebook{Alma}{Alma}%
\addbiblebook{Hel.}{Helaman}%
\addbiblebook{3 Ne.}{3 Nephi}%
\addbiblebook{4 Ne.}{4 Nephi}%
\addbiblebook{Morm.}{Mormon}%
\addbiblebook{Ether}{Ether}%
\addbiblebook{Moro.}{Moroni}%


% fonts
\usepackage{fontspec}
%\usepackage{xunicode}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont{EB Garamond}

\usepackage{pgffor}
\usepackage{ifthen}
\usepackage{lettrine}
\usepackage[all]{nowidow}

\usepackage{trimspaces}

% Balance twocolumn
\usepackage{balance}

\title{Bible de Genève\\1564}

\setlength\marginparsep{8pt}
\setlength\marginparwidth{8pt}

% Only put a sep if not at the beginning of a line
\newcommand{\vsep}{\unskip%
  \cleaders\copy\vsepbox\hskip\wd\vsepbox%
  \ignorespaces}
\newsavebox\vsepbox
\sbox\vsepbox{\vrule width 0.5pt height 5pt}
%\sbox\vsepbox{|}

% Romans 3 32
\def\bref{}
\newcounter{verse}
\makeatletter
\newcommand\bverse{%
  \addtocounter{verse}{1}
  \ifthenelse{\equal{\theverse}{1}}%
    {\lettrine{\liningnums{\arabic{section}}}{}}%
    {\vsep\marginpar{\footnotesize \theverse}}
  \def\bref{\protect\bibleverse{\curbook}(\arabic{section}:\theverse)}
  \extramark{bref}
}
\makeatother

\newcommand\curbook{}
\newcommand{\bbook}[2]{%
  \chapter{#1}
  \setcounter{section}{0}
  \renewcommand\curbook{#2}
}

\newcommand\bchapter{%
  \addtocounter{section}{1}
  \setcounter{verse}{0}
}

% Adjust headings
\usepackage{titlesec}
\usepackage[extramarks]{titleps}
\newfontfamily\chapheadfont{EB Garamond}
\newcommand{\chapheadstyle}{\centering\chapheadfont\Huge\textsc}
\titleformat{\chapter}[display]
  {}
  {}{0pt}
  {\chapheadstyle}

% http://tex.stackexchange.com/questions/26392/titlesec-toptitlemarks-and-bottitlemarks-broken-when-section-not-called-explic
% could also be interesting

\newcommand*{\cadr}{\nobreak\,\nobreak\mbox{---}\nobreak\,\nobreak}
\newmarkset{bref}
\newextramark{bref}{\bref}
\newpagestyle{Biblestyle}{
  \setheadrule{.009pt}
  \sethead[\thepage][\textsc{\topextramarks{bref}\bref\cadr\botextramarks{bref}\bref}]
  []{}{\textsc{\topextramarks{bref}\bref\cadr\botextramarks{bref}\bref}}{\thepage}
}
\pagestyle{Biblestyle}

\begin{document}

\maketitle
\balance

\bbook{The Book of Mosiah}{Mosiah}

\bchapter
\bverse \iffalse And now there was no more contention in all the land of Zarahemla, among
all the people who belonged to king Benjamin, so that king Benjamin had
continual peace all the remainder of his days.\fi
\bverse \iffalse And it came to pass that he had three sons; and he called their names
Mosiah, and Helorum, and Helaman. And he caused that they should be taught in
all the language of his fathers, that thereby they might become men of
understanding; and that they might know concerning the prophecies which had been
spoken by the mouths of their fathers, which were delivered them by the hand of
the Lord.\fi
\bverse \iffalse And he also taught them concerning the records which were engraven on
the plates of brass, saying: My sons, I would that ye should remember that were
it not for these plates, which contain these records and these commandments, we
must have suffered in ignorance, even at this present time, not knowing the
mysteries of God.\fi
\bverse \iffalse For it were not possible that our father, Lehi, could have remembered
all these things, to have taught them to his children, except it were for the
help of these plates; for he having been taught in the language of the Egyptians
therefore he could read these engravings, and teach them to his children, that
thereby they could teach them to their children, and so fulfilling the
commandments of God, even down to this present time.\fi
\bverse \iffalse I say unto you, my sons, were it not for these things, which have been
kept and preserved by the hand of God, that we might read and understand of his
mysteries, and have his commandments always before our eyes, that even our
fathers would have dwindled in unbelief, and we should have been like unto our
brethren, the Lamanites, who know nothing concerning these things, or even do
not believe them when they are taught them, because of the traditions of their
fathers, which are not correct.\fi
\bverse \iffalse O my sons, I would that ye should remember that these sayings are true,
and also that these records are true. And behold, also the plates of Nephi,
which contain the records and the sayings of our fathers from the time they left
Jerusalem until now, and they are true; and we can know of their surety because
we have them before our eyes.\fi
\bverse \iffalse And now, my sons, I would that ye should remember to search them
diligently, that ye may profit thereby; and I would that ye should keep the
commandments of God, that ye may prosper in the land according to the promises
which the Lord made unto our fathers.\fi
\bverse \iffalse And many more things did king Benjamin teach his sons, which are not
written in this book.\fi
\bverse \iffalse And it came to pass that after king Benjamin had made an end of teaching
his sons, that he waxed old, and he saw that he must very soon go the way of all
the earth; therefore, he thought it expedient that he should confer the kingdom
upon one of his sons.\fi
\bverse \iffalse Therefore, he had Mosiah brought before him; and these are the words
which he spake unto him, saying: My son, I would that ye should make a
proclamation throughout all this land among all this people, or the people of
Zarahemla, and the people of Mosiah who dwell in the land, that thereby they may
be gathered together; for on the morrow I shall proclaim unto this my people out
of mine own mouth that thou art a king and a ruler over this people, whom the
Lord our God hath given us.\fi
\bverse \iffalse And moreover, I shall give this people a name, that
thereby they may be distinguished above all the people which the Lord God hath
brought out of the land of Jerusalem; and this I do because they have been a
diligent people in keeping the commandments of the Lord.\fi
\bverse \iffalse And I give
unto them a name that never shall be blotted out, except it be through
transgression.
\fi
\bverse \iffalse Yea, and moreover I say unto you, that if this highly
favored people of the Lord should fall into transgression, and become a wicked
and an adulterous people, that the Lord will deliver them up, that thereby they
become weak like unto their brethren; and he will no more preserve them by his
matchless and marvelous power, as he has hitherto preserved our fathers.\fi
\bverse \iffalse For I say unto you, that if he had not extended his arm in the
preservation of our fathers they must have fallen into the hands of the
Lamanites, and become victims to their hatred.\fi
\bverse \iffalse And it came to pass that
after king Benjamin had made an end of these sayings to his son, that he gave
him charge concerning all the affairs of the kingdom.\fi
\bverse \iffalse And moreover, he
also gave him charge concerning the records which were engraven on the plates of
brass; and also the plates of Nephi; and also, the sword of Laban, and the ball
or director, which led our fathers through the wilderness, which was prepared by
the hand of the Lord that thereby they might be led, every one according to the
heed and diligence which they gave unto him.\fi
\bverse \iffalse Therefore, as they were
unfaithful they did not prosper nor progress in their journey, but were driven
back, and incurred the displeasure of God upon them; and therefore they were
smitten with famine and sore afflictions, to stir them up in remembrance of
their duty.\fi
\bverse \iffalse And now, it came to pass that Mosiah went and did as his
father had commanded him, and proclaimed unto all the people who were in the
land of Zarahemla that thereby they might gather themselves together, to go up
to the temple to hear the words which his father should speak unto them. \fi

% Chapitre 2
\bchapter
\bverse Les cieux donc et la terre furent parfaicts, et tout l’exercite d’iceux:
\bverse Car Dieu avoit accompli au septieme jour son œuvre qu’il avoit faicte, et se reposa au septieme jour de toute son œuvre qu’il avoit faicte.
\bverse Et Dieu benit le septieme iour, et le sanctifia: pour ce qu’en icelui il avoit cessé de toute son œuvre, qu’il avoit creée pour estre faicte.
\bverse Telles sont les generations du ciel et de la terre, quand ils furent creés, au jour que le Seigneur Dieu fit le ciel et la terre,
\bverse Et tout jetton du champ devant qu’il fust en la terre, et tout herbage du champ devant qu’il germast: car le Seigneur Dieu n’avoit point faict plouvoir sur la terre, et n’y avoit homme pour labourer la terre:
\bverse Mais une vapeur montoit de la terre, qui arroisoit tout le dessus de la terre.
\bverse Or le Seigneur Dieu avoit formé l’homme de la poudre de la terre, et avoit inspiré en la face d’icelui spiration de vie, et l’homme fut faict en ame vivante.

\end{document}


