\twocolumn[{%
 \centering
\bbook{THIRD BOOK OF NEPHI}{Alma}
\LARGE THE SON OF NEPHI, WHO WAS THE SON OF HELAMAN \\[1.5em]
 \normalsize
\iffalse And Helaman was the son of Helaman, who was the son of Alma, who was the son of Alma, being a descendant of Nephi who was the son of Lehi, who came out of Jerusalem in the first year of the reign of Zedekiah, the king of Judah. \fi
Helaman was the son of Helaman, who was the son of Alma, who was the son of Alma. And Alma was a descendant of Nephi, who was the son of Sariah and Lehi, who came out of Jerusalem during the first year that Zedekiah, the king of Judah, was the ruler.
\vspace{0.5cm} %Put space between opening paragraph and first chapter
}]
\renewcommand\curbook{3 Ne.}

\balance

%Chapter 1
\bchapter
%3 Nephi 1:1
\bverse \iffalse Now it came to pass that the ninety and first year had passed away and it was six hundred years from the time that Lehi left Jerusalem; and it was in the year that Lachoneus was the chief judge and the governor over the land. \fi
The ninety-first year had ended. Lachoneus was the chief judge and governor over the land, and it was six hundred years from the time that Sariah and Lehi left Jerusalem.
%3 Nephi 1:2
\bverse \iffalse And Nephi, the son of Helaman, had departed out of the land of Zarahemla, giving charge unto his son Nephi, who was his eldest son, concerning the plates of brass, and all the records which had been kept, and all those things which had been kept sacred from the departure of Lehi out of Jerusalem. \fi
And Nephi, the son of Helaman, left the land of Zarahemla and put his oldest son, Nephi, in charge of the brass plates, and all the records that had been kept, and all the things that had been kept sacred since Sariah and Lehi left Jerusalem.
%3 Nephi 1:3
\bverse \iffalse Then he departed out of the land, and whither he went, no man knoweth; and his son Nephi did keep the records in his stead, yea, the record of this people. \fi
Then he left the land of Zarahemla, and no one knew where he went. His son, Nephi, began to keep the records. He kept the record of their people.
%3 Nephi 1:4
\bverse \iffalse And it came to pass that in the commencement of the ninety and second year, behold, the prophecies of the prophets began to be fulfilled more fully; for there began to be greater signs and greater miracles wrought among the people. \fi
And in the beginnning of the ninety-second year many more of the prophecies that the prophets had made began to come true. And even greater signs and miracles began to be shown among the people.
%3 Nephi 1:5
\bverse \iffalse But there were some who began to say that the time was past for the words to be fulfilled, which were spoken by Samuel, the Lamanite. \fi
But some people began to say that the time for the words that Samuel, the Lamanite, had spoken to come true had already gone past.
%3 Nephi 1:6
\bverse \iffalse And they began to rejoice over their brethren, saying: Behold the time is past, and the words of Samuel are not fulfilled; therefore, your joy and your faith concerning this thing hath been vain. \fi
And they began to celebrate and make fun of their sisters and brothers who believed. They said: ``The time is past, and Samuel's words have not come true. You were foolish to believe the things that he said.''
%3 Nephi 1:7
\bverse \iffalse And it came to pass that they did make a great uproar throughout the land; and the people who believed began to be very sorrowful, lest by any means those things which had been spoken might not come to pass. \fi
And they caused chaos and arguments all over the land. And the people who believed began to be very sad that perhaps the prophecies that had been spoken might not come true.
%3 Nephi 1:8
\bverse \iffalse But behold, they did watch steadfastly for that day and that night and that day which should be as one day as if there were no night, that they might know that their faith had not been vain. \fi
But they watched carefully for the two days and one night that would be like one day with no night. They hoped it would come and show that it was not a mistake for them to have faith and believe.
%3 Nephi 1:9
\bverse \iffalse Now it came to pass that there was a day set apart by the unbelievers, that all those who believed in those traditions should be put to death except the sign should come to pass, which had been given by Samuel the prophet. \fi
And the people who did not believe had decided on a day, and on that day they decided they would put to death anyone who believed in those traditions, unless they saw the sign that Samuel the prophet had prophesied about.
%3 Nephi 1:10
\bverse \iffalse Now it came to pass that when Nephi, the son of Nephi, saw this wickedness of his people, his heart was exceedingly sorrowful. \fi
And when Nephi, the son of Nephi, saw how wicked his people were, his heart was filled with sadness.
%3 Nephi 1:11
\bverse \iffalse And it came to pass that he went out and bowed himself down upon the earth, and cried mightily to his God in behalf of his people, yea, those who were about to be destroyed because of their faith in the tradition of their fathers. \fi
He went out and bowed down on the ground, and cried out to God for his people that were about to be killed because they had faith in the traditions of their parents and ancestors.
%3 Nephi 1:12
\bverse \iffalse And it came to pass that he cried mightily unto the Lord, all that day; and behold, the voice of the Lord came unto him, saying: \fi
He cried out to the Lord the whole day. And the voice of the Lord came to him, and said:
%3 Nephi 1:13
\bverse \iffalse Lift up your head and be of good cheer; for behold, the time is at hand, and on this night shall the sign be given, and on the morrow come I into the world, to show unto the world that I will fulfill all that which I have caused to be spoken by the mouth of my holy prophets. \fi
``Lift up your head and be glad. The time is here, and tonight the sign will appear. Tomorrow I am coming into the world, to show the world that I will cause everything that I have told my holy prophets to speak to come true.
%3 Nephi 1:14
\bverse \iffalse Behold, I come unto my own, to fulfill all things which I have made known unto the children of men from the foundation of the world, and to do the will, both of the Father and of the Son--of the Father because of me, and of the Son because of my flesh. And behold, the time is at hand, and this night shall the sign be given. \fi
I am coming to my people, to cause all things that I have revealed to humankind since the beginning of the world to happen, and to do the will of the Father and the Son. I will do the will of the Father because I am His Son, and will I do the will of the Son because I will have a body of flesh. See---the time has come, and tonight the sign will appear.''
%3 Nephi 1:15
\bverse \iffalse And it came to pass that the words which came unto Nephi were fulfilled, according as they had been spoken; for behold, at the going down of the sun there was no darkness; and the people began to be astonished because there was no darkness when the night came. \fi
And the words that were told to Nephi all came true. And when the sun went down, there was no darkness. And the people began to be amazed, because there was no darkness when the night came.
%3 Nephi 1:16
\bverse \iffalse And there were many, who had not believed the words of the prophets, who fell to the earth and became as if they were dead, for they knew that the great plan of destruction which they had laid for those who believed in the words of the prophets had been frustrated; for the sign which had been given was already at hand. \fi
There were many people who had not believed the words of the prophets. They fell to the earth as if they were dead, because they knew that the plan that they had made to kill the people who believed the words of the prophets was destroyed.
%3 Nephi 1:17
\bverse \iffalse And they began to know that the Son of God must shortly appear; yea, in fine, all the people upon the face of the whole earth from the west to the east, both in the land north and in the land south, were so exceedingly astonished that they fell to the earth. \fi
They realized that the Son of God must appear soon. And all the people over the whole land, from the west to the east, and from the north to the south, were so amazed that they fell to the ground.
%3 Nephi 1:18
\bverse \iffalse For they knew that the prophets had testified of these things for many years, and that the sign which had been given was already at hand; and they began to fear because of their iniquity and their unbelief. \fi
They knew that the prophets had testified about these things for many years, and that the sign that the prophets said would come had already appeared. And the people began to be afraid because they knew that they had been wicked and had not believed.
%3 Nephi 1:19
\bverse \iffalse And it came to pass that there was no darkness in all that night, but it was as light as though it was mid-day. And it came to pass that the sun did rise in the morning again, according to its proper order; and they knew that it was the day that the Lord should be born, because of the sign which had been given. \fi
There was no darkness all that night; it was as light as though it were mid-day. And the sun rose in the morning, as usual. They knew that this was the day that the Lord would be born, because of the sign that had appeared.
%3 Nephi 1:20
\bverse \iffalse And it had come to pass, yea, all things, every whit, according to the words of the prophets. \fi
And it had all happened---every single thing---just as the prophets had spoken.
%3 Nephi 1:21
\bverse \iffalse And it came to pass also that a new star did appear, according to the word. \fi
A new star also appeared, just as the prophets had spoken.
%3 Nephi 1:22
\bverse \iffalse And it came to pass that from this time forth there began to be lyings sent forth among the people, by Satan, to harden their hearts, to the intent that they might not believe in those signs and wonders which they had seen; but notwithstanding these lyings and deceivings the more part of the people did believe, and were converted unto the Lord. \fi
And from this time onward lies began to be spread among the people by Satan, to harden their hearts and get them to not believe the signs and miracles they had seen. But even though Satan sent out these lies and tricks, most of the the people believed and were converted to the Lord.
%3 Nephi 1:23
\bverse \iffalse And it came to pass that Nephi went forth among the people, and also many others, baptizing unto repentance, in the which there was a great remission of sins. And thus the people began again to have peace in the land. \fi
And Nephi and many others went out among the people, baptizing people as a sign that they had repented. The sins of many people were forgiven, and there began to be peace in the land.
%3 Nephi 1:24
\bverse \iffalse And there were no contentions, save it were a few that began to preach, endeavoring to prove by the scriptures that it was no more expedient to observe the law of Moses. Now in this thing they did err, having not understood the scriptures. \fi
And there were no arguments among the people, except that a few people began to preach and try to prove using the scriptures that it was no longer necessary to keep the law of Moses. But they made a mistake; they had not understood the scriptures.
%3 Nephi 1:25
\bverse \iffalse But it came to pass that they soon became converted, and were convinced of the error which they were in, for it was made known unto them that the law was not yet fulfilled, and that it must be fulfilled in every whit; yea, the word came unto them that it must be fulfilled; yea, that one jot or tittle should not pass away till it should all be fulfilled; therefore in this same year were they brought to a knowledge of their error and did confess their faults. \fi
But as soon as they were converted, and became convinced that they had made a mistake, they confessed that they had been wrong. They found out that the time had not yet come to stop following the law of Moses. They found out that every single bit of the law and prophecy in the law of Moses had to be completed and come true.
%3 Nephi 1:26
\bverse \iffalse And thus the ninety and second year did pass away, bringing glad tidings unto the people because of the signs which did come to pass, according to the words of the prophecy of all the holy prophets. \fi
And the ninety-second year ended. During this year wonderful news had come to the people because of the signs that had come true, just as the holy prophets had prophesied.
%3 Nephi 1:27
\bverse \iffalse And it came to pass that the ninety and third year did also pass away in peace, save it were for the Gadianton robbers, who dwelt upon the mountains, who did infest the land; for so strong were their holds and their secret places that the people could not overpower them; therefore they did commit many murders, and did do much slaughter among the people. \fi
And the ninety-third year also ended peacefully, except that there were still Gadianton robbers who lived in the mountains and lived secretly among the people. The forts and defenses of the Gadianton robbers were so strong that the people could not overpower them. And they committed many murders and caused many deaths among the people.
%3 Nephi 1:28
\bverse \iffalse And it came to pass that in the ninety and fourth year they began to increase in great degree, because there were many dissenters of the Nephites who did flee unto them, which did cause much sorrow unto those Nephites who did remain in the land. \fi
And in the ninety-fourth year there began to be many more of them, because many people rebelled against the Nephites and left them, and ran to the Gadianton robbers. This made the Nephites who stayed in the land very sad.
%3 Nephi 1:29
\bverse \iffalse And there was also a cause of much sorrow among the Lamanites; for behold, they had many children who did grow up and began to wax strong in years, that they became for themselves, and were led away by some who were Zoramites, by their lyings and their flattering words, to join those Gadianton robbers. \fi
This also made the Lamanites very sad. They had many children who had grown up, and who had been led away by the lies and flattery of some Zoramites to join the Gadianton robbers.
%3 Nephi 1:30
\bverse \iffalse And thus were the Lamanites afflicted also, and began to decrease as to their faith and righteousness, because of the wickedness of the rising generation. \fi
So the Lamanites were also very sad. They started to lose their faith and become less righteous, because of the younger generations.
%Chapter 2
\bchapter
%3 Nephi 2:1
\bverse \iffalse And it came to pass that thus passed away the ninety and fifth year also, and the people began to forget those signs and wonders which they had heard, and began to be less and less astonished at a sign or a wonder from heaven, insomuch that they began to be hard in their hearts, and blind in their minds, and began to disbelieve all which they had heard and seen-- \fi
So the ninety-fifth year ended, and the people began to forget about the signs and miracles that they had heard. They began to be less and less surprised by signs and miracles from heaven, and their hearts became hard. Their minds became blind, and they stopped believing everything that they had heard and seen.
%3 Nephi 2:2
\bverse \iffalse Imagining up some vain thing in their hearts, that it was wrought by men and by the power of the devil, to lead away and deceive the hearts of the people; and thus did Satan get possession of the hearts of the people again, insomuch that he did blind their eyes and lead them away to believe that the doctrine of Christ was a foolish and a vain thing. \fi
They imagined foolish and empy things---that those miracles were worked by humans, or the power of the devil, to try to trick them into believing. This is how Satan got a hold of the hearts of the people again. He clouded their minds and led them to believe that the teachings and prophecies of Christ were foolish.
%3 Nephi 2:3
\bverse \iffalse And it came to pass that the people began to wax strong in wickedness and abominations; and they did not believe that there should be any more signs or wonders given; and Satan did go about, leading away the hearts of the people, tempting them and causing them that they should do great wickedness in the land. \fi
And the people began to be very wicked and do many horrible things. They did not believe that there would be any more signs or miracles. And Satan went around, leading away the hearts of the people. He tempted them and got them to do many very wicked things.
%3 Nephi 2:4
\bverse \iffalse And thus did pass away the ninety and sixth year; and also the ninety and seventh year; and also the ninety and eighth year; and also the ninety and ninth year; \fi
This is how the ninety-sixth year to the ninety-ninth year ended.
%3 Nephi 2:5
\bverse \iffalse And also an hundred years had passed away since the days of Mosiah, who was king over the people of the Nephites. \fi
And a hundred years had gone by since Mosiah, the king over the people of the Nephites, had lived.
%3 Nephi 2:6
\bverse \iffalse And six hundred and nine years had passed away since Lehi left Jerusalem. \fi
And six hundred and nine years had gone by since Sariah and Lehi left Jerusalem.
%3 Nephi 2:7
\bverse \iffalse And nine years had passed away from the time when the sign was given, which was spoken of by the prophets, that Christ should come into the world. \fi
And nine years had gone by since the sign that Christ would be born had appeared. The prophets had spoken about this sign.
%3 Nephi 2:8
\bverse \iffalse Now the Nephites began to reckon their time from this period when the sign was given, or from the coming of Christ; therefore, nine years had passed away. \fi
The Nephites began to count their years from the time that the sign was given that Christ was born. Nine years had gone by since that time.
%3 Nephi 2:9
\bverse \iffalse And Nephi, who was the father of Nephi, who had the charge of the records, did not return to the land of Zarahemla, and could nowhere be found in all the land. \fi
And Nephi, the father of that Nephi who was in charge of the records, did not come back to the land of Zarahemla. He could not be found anywhere in the land.
%3 Nephi 2:10
\bverse \iffalse And it came to pass that the people did still remain in wickedness, notwithstanding the much preaching and prophesying which was sent among them; and thus passed away the tenth year also; and the eleventh year also passed away in iniquity. \fi
And the people continued to live very wickedly, even though many people were sent to preach and prophesy to them. This is how the tenth year ended. The people were still doing very wicked things in the eleventh year.
%3 Nephi 2:11
\bverse \iffalse And it came to pass in the thirteenth year there began to be wars and contentions throughout all the land; for the Gadianton robbers had become so numerous, and did slay so many of the people, and did lay waste so many cities, and did spread so much death and carnage throughout the land, that it became expedient that all the people, both the Nephites and the Lamanites, should take up arms against them. \fi
And during the thirteenth year wars and fighting spread through the land. The gang of Gadianton robbers had grown very large, and killed very many people and destroyed very many cities. They caused so much death and fighting that all the people, both Nephites and Lamanites, saw that they had to go to war with them.
%3 Nephi 2:12
\bverse \iffalse Therefore, all the Lamanites who had become converted unto the Lord did unite with their brethren, the Nephites, and were compelled, for the safety of their lives and their women and their children, to take up arms against those Gadianton robbers, yea, and also to maintain their rights, and the privileges of their church and of their worship, and their freedom and their liberty. \fi
So all the Lamanites who were converted and believed in the Lord united with the Nephites. For the safety of their families and their own lives, and to keep their right to have a church and worship, and to keep their freedom, they realized that they had to go to war against the Gadianton robbers.
%3 Nephi 2:13
\bverse \iffalse And it came to pass that before this thirteenth year had passed away the Nephites were threatened with utter destruction because of this war, which had become exceedingly sore. \fi
And before this thirteenth year ended, the Nephites were about to be completely destroyed because of this war, which had become very bad.
%3 Nephi 2:14
\bverse \iffalse And it came to pass that those Lamanites who had united with the Nephites were numbered among the Nephites; \fi
And the Lamanites who had joined with the Nephites were also considered Nephites.
%3 Nephi 2:15
\bverse \iffalse And their curse was taken from them, and their skin became white like unto the Nephites; \fi
The curse that was on them was taken away.
%3 Nephi 2:16
\bverse \iffalse And their young men and their daughters became exceedingly fair, and they were numbered among the Nephites, and were called Nephites. And thus ended the thirteenth year. \fi
Their young men and their daughters became very handsome and beautiful. They were counted as Nephites, and they were called Nephites. And the thirteenth year ended.
%3 Nephi 2:17
\bverse \iffalse And it came to pass in the commencement of the fourteenth year, the war between the robbers and the people of Nephi did continue and did become exceedingly sore; nevertheless, the people of Nephi did gain some advantage of the robbers, insomuch that they did drive them back out of their lands into the mountains and into their secret places. \fi
And at the beginning of the fourteenth year, the war between the robbers and the Nephites continued, and got very bad. But the Nephites did have some success; they managed to drive the robbers out of their lands, back into the mountains and their hideouts.
%3 Nephi 2:18
\bverse \iffalse And thus ended the fourteenth year. And in the fifteenth year they did come forth against the people of Nephi; and because of the wickedness of the people of Nephi, and their many contentions and dissensions, the Gadianton robbers did gain many advantages over them. \fi
And the fourteenth year ended. And in the fifteenth year, the robbers came to battle against the Nephites. The Nephites were wicked, and they fought and argued very much among themselves, so the Gadianton robbers had many advantages over them.
%3 Nephi 2:19
\bverse \iffalse And thus ended the fifteenth year, and thus were the people in a state of many afflictions; and the sword of destruction did hang over them, insomuch that they were about to be smitten down by it, and this because of their iniquity. \fi
And the fifteenth year ended, and the Nephites were suffering in many ways. The sword of destruction was hanging over them, and it was about to fall down onto them, because of the many sins they committed.
%Chapter 3
\bchapter
%3 Nephi 3:1
\bverse \iffalse And now it came to pass that in the sixteenth year from the coming of Christ, Lachoneus, the governor of the land, received an epistle from the leader and the governor of this band of robbers; and these were the words which were written, saying: \fi
And sixteen years after Christ had come, Lachoneus, who was the governor of the land, got a letter from the leader and governor of the gang of robbers. This is what the letter said:
%3 Nephi 3:2
\bverse \iffalse Lachoneus, most noble and chief governor of the land, behold, I write this epistle unto you, and do give unto you exceedingly great praise because of your firmness, and also the firmness of your people, in maintaining that which ye suppose to be your right and liberty; yea, ye do stand well, as if ye were supported by the hand of a god, in the defence of your liberty, and your property, and your country, or that which ye do call so. \fi
``Lachoneus, great and chief governor of the land, I am writing to praise you greatly because you and your people have been so strong in defending what you call your rights and your freedom. You are very good fighters; it almost seems like there is some god who is helping you to defend what you call your freedom, your property, and your country.
%3 Nephi 3:3
\bverse \iffalse And it seemeth a pity unto me, most noble Lachoneus, that ye should be so foolish and vain as to suppose that ye can stand against so many brave men who are at my command, who do now at this time stand in their arms, and do await with great anxiety for the word--Go down upon the Nephites and destroy them. \fi
But I think it is a pity, most noble Lachoneus, that you are so foolish to think that you can defend yourselves against the many brave men that I command. Even right now they are standing with their weapons, and they are waiting impatiently for me to give the command---Go down and attack the Nephites, and destroy them.
%3 Nephi 3:4
\bverse \iffalse And I, knowing of their unconquerable spirit, having proved them in the field of battle, and knowing of their everlasting hatred towards you because of the many wrongs which ye have done unto them, therefore if they should come down against you they would visit you with utter destruction. \fi
I have tested them in battle, and I know that they cannot be conquered. I know that they hate you with a hate that will never go away, because of the many horrible things you have done to them. And if they came down to attack you, they would completely destroy you.
%3 Nephi 3:5
\bverse \iffalse Therefore I have written this epistle, sealing it with mine own hand, feeling for your welfare, because of your firmness in that which ye believe to be right, and your noble spirit in the field of battle. \fi
This is why I am writing this letter, and signing with my own hand. I feel sorry for you, because you fight so bravely to defend what you believe is right.
%3 Nephi 3:6
\bverse \iffalse Therefore I write unto you, desiring that ye would yield up unto this my people, your cities, your lands, and your possessions, rather than that they should visit you with the sword and that destruction should come upon you. \fi
I am writing to you to tell you to give up. I would rather have you give up yourselves, your cities, your lands, and the things you own to my people, than have them come and completely destroy you.
%3 Nephi 3:7
\bverse \iffalse Or in other words, yield yourselves up unto us, and unite with us and become acquainted with our secret works, and become our brethren that ye may be like unto us--not our slaves, but our brethren and partners of all our substance. \fi
In other words, stop fighting with us. Join us and come and learn about our secrets. Become our friends and family, and be like us---not our slaves, but our family and friends in all the things that we have and do.
%3 Nephi 3:8
\bverse \iffalse And behold, I swear unto you, if ye will do this, with an oath, ye shall not be destroyed; but if ye will not do this, I swear unto you with an oath, that on the morrow month I will command that my armies shall come down against you, and they shall not stay their hand and shall spare not, but shall slay you, and shall let fall the sword upon you even until ye shall become extinct. \fi
And I swear to you with a promise, that if you will do this you will not be destroyed. But if you will not do this, I swear to you with a promise that next month, I will command my armies to come and attack you, and they will not spare anyone. They will kill you, and will keep killing you until you are extinct.
%3 Nephi 3:9
\bverse \iffalse And behold, I am Giddianhi; and I am the governor of this the secret society of Gadianton; which society and the works thereof I know to be good; and they are of ancient date and they have been handed down unto us. \fi
I am Giddianhi, the governor of Gadianton's secret society. I know this society and the things that we do are good. They are from ancient times, and they have been handed down to us.
%3 Nephi 3:10
\bverse \iffalse And I write this epistle unto you, Lachoneus, and I hope that ye will deliver up your lands and your possessions, without the shedding of blood, that this my people may recover their rights and government, who have dissented away from you because of your wickedness in retaining from them their rights of government, and except ye do this, I will avenge their wrongs. I am Giddianhi. \fi
I write this letter to you, Lachoneus, and I hope that you will give up your lands and the things you own without a fight, so that my people can take back their rights and the government. My people have left your people because you have so wickedly taken their right to govern the people from them. And if you do not do this, I will take revenge. I am Giddianhi.''
%3 Nephi 3:11
\bverse \iffalse And now it came to pass when Lachoneus received this epistle he was exceedingly astonished, because of the boldness of Giddianhi demanding the possession of the land of the Nephites, and also of threatening the people and avenging the wrongs of those that had received no wrong, save it were they had wronged themselves by dissenting away unto those wicked and abominable robbers. \fi
When Lachoneus got this letter he was very surprised that Giddianhi had been so bold, and demanded that they give up the land of the Nephites. He was amazed that he threatened them and said that he would take revenge for his people, who Giddianhi said had been harmed and mistreated by the Nephites. Lachoneus knew that they had caused harm to themselves by leaving the Nephites and joining that gang of evil robbers.
%3 Nephi 3:12
\bverse \iffalse Now behold, this Lachoneus, the governor, was a just man, and could not be frightened by the demands and the threatenings of a robber; therefore he did not hearken to the epistle of Giddianhi, the governor of the robbers, but he did cause that his people should cry unto the Lord for strength against the time that the robbers should come down against them. \fi
Lachoneus, the governor, was a fair man, and he would not let himself be frightened by the demands and threats of a robber. He did not do as Giddianhi had said in his letter, but he told his people to cry out to the Lord for strength and protection when the robbers came to attack them.
%3 Nephi 3:13
\bverse \iffalse Yea, he sent a proclamation among all the people, that they should gather together their women, and their children, their flocks and their herds, and all their substance, save it were their land, unto one place. \fi
He sent a proclamation to all the people. He told them to gather their families together, and their flocks and their herds, and everything they owned except for their land, into one place.
%3 Nephi 3:14
\bverse \iffalse And he caused that fortifications should be built round about them, and the strength thereof should be exceedingly great. And he caused that armies, both of the Nephites and of the Lamanites, or of all them who were numbered among the Nephites, should be placed as guards round about to watch them, and to guard them from the robbers day and night. \fi
He had them build very, very strong forts and defenses around themselves. And he had the armies of the Nephites and Lamanites, who were all called Nephites, stand guard around them and protect them from the robbers, both day and night.
%3 Nephi 3:15
\bverse \iffalse Yea, he said unto them: As the Lord liveth, except ye repent of all your iniquities, and cry unto the Lord, ye will in no wise be delivered out of the hands of those Gadianton robbers. \fi
He said to them: ``Just as the Lord lives, if you do not repent of your sins and cry out to the Lord, you will not be protected from those Gadianton robbers.''
%3 Nephi 3:16
\bverse \iffalse And so great and marvelous were the words and prophecies of Lachoneus that they did cause fear to come upon all the people; and they did exert themselves in their might to do according to the words of Lachoneus. \fi
The things that Lachoneus said and prophesied were so great that they caused the people to be very afraid, and they worked very hard to do what Lachoneus had said.
%3 Nephi 3:17
\bverse \iffalse And it came to pass that Lachoneus did appoint chief captains over all the armies of the Nephites, to command them at the time that the robbers should come down out of the wilderness against them. \fi
And Lachoneus appointed chief captains over the armies of the Nephites to lead them when the robbers came out of the wilderness to attack them.
%3 Nephi 3:18
\bverse \iffalse Now the chiefest among all the chief captains and the great commander of the armies of the Nephites was appointed, and his name was Gidgiddoni. \fi
And the chief of all the chief captains, the great commander of the Nephite armies, was named Gidgiddoni.
%3 Nephi 3:19
\bverse \iffalse Now it was the custom among all the Nephites to appoint for their chief captains, (save it were in their times of wickedness) some one that had the spirit of revelation and also prophecy; therefore, this Gidgiddoni was a great prophet among them, as also was the chief judge. \fi
The custom of the Nephites was to appoint as their chief captain someone who had the spirit of revelation and prophecy. (They did this at all times except when they were living wickedly.) And Gidgiddoni was a great prophet among them. He was also the chief judge.
%3 Nephi 3:20
\bverse \iffalse Now the people said unto Gidgiddoni: Pray unto the Lord, and let us go up upon the mountains and into the wilderness, that we may fall upon the robbers and destroy them in their own lands. \fi
The people said to Gidgiddoni: ``Pray to the Lord, and let's go into the mountains and into the wilderness, and attack the robbers and destroy them in their own lands.''
%3 Nephi 3:21
\bverse \iffalse But Gidgiddoni saith unto them: The Lord forbid; for if we should go up against them the Lord would deliver us into their hands; therefore we will prepare ourselves in the center of our lands, and we will gather all our armies together, and we will not go against them, but we will wait till they shall come against us; therefore as the Lord liveth, if we do this he will deliver them into our hands. \fi
But Gidgiddoni said to them: ``No! If we go to attack them the Lord would let us be defeated by them. We will get ready in the center of our lands, and gather all our armies together. We will not go to attack them; instead, we will wait until they come to attack us. If we do this the Lord will help us to defeat them, as sure as the Lord lives.''
%3 Nephi 3:22
\bverse \iffalse And it came to pass in the seventeenth year, in the latter end of the year, the proclamation of Lachoneus had gone forth throughout all the face of the land, and they had taken their horses, and their chariots, and their cattle, and all their flocks, and their herds, and their grain, and all their substance, and did march forth by thousands and by tens of thousands, until they had all gone forth to the place which had been appointed that they should gather themselves together, to defend themselves against their enemies. \fi
And toward the end of the seventeenth year the proclamation of Lachoneus had gone through the whole land. The people took their horses and chariots, and their cattle and flocks, and their herds and grain, and everything they owned, to the place where they agreed that they should gather together to defend themselves. And there were thousands and thousands of people who came and gathered together.
%3 Nephi 3:23
\bverse \iffalse And the land which was appointed was the land of Zarahemla, and the land which was between the land Zarahemla and the land Bountiful, yea, to the line which was between the land Bountiful and the land Desolation. \fi
And the land where they decided to meet was the land of Zarahemla, and the land that was between the land Zarahemla and the land Bountiful, all the way to the line that was between the land Bountiful and The Ruins.
%3 Nephi 3:24
\bverse \iffalse And there were a great many thousand people who were called Nephites, who did gather themselves together in this land. Now Lachoneus did cause that they should gather themselves together in the land southward, because of the great curse which was upon the land northward. \fi
And there were thousands of Nephites who gathered together in this land. And Lachoneus had them gather themselves together in the land to the south, because of the great curse that was on the land in the north.
%3 Nephi 3:25
\bverse \iffalse And they did fortify themselves against their enemies; and they did dwell in one land, and in one body, and they did fear the words which had been spoken by Lachoneus, insomuch that they did repent of all their sins; and they did put up their prayers unto the Lord their God, that he would deliver them in the time that their enemies should come down against them to battle. \fi
And they prepared defenses against their enemies. They lived in one land, and in one group. They believed the things that Lachoneus had told them, and they repented of all their sins. They prayed to the Lord their God, asking him to save them from their enemies when they came to attack them.
%3 Nephi 3:26
\bverse \iffalse And they were exceedingly sorrowful because of their enemies.  And Gidgiddoni did cause that they should make weapons of war of every kind, and they should be strong with armor, and with shields, and with bucklers, after the manner of his instruction. \fi
They were full of sadness because of their enemies. And Gidgiddoni had them follow his instructions and make all kinds of weapons, armor, and shields.
%Chapter 4
\bchapter
%3 Nephi 4:1
\bverse \iffalse And it came to pass that in the latter end of the eighteenth year those armies of robbers had prepared for battle, and began to come down and to sally forth from the hills, and out of the mountains, and the wilderness, and their strongholds, and their secret places, and began to take possession of the lands, both which were in the land south and which were in the land north, and began to take possession of all the lands which had been deserted by the Nephites, and the cities which had been left desolate. \fi
And at the end of the eighteenth year those armies of robbers had prepared for battle. They began to come down out of the hills, and mountains, and the wilderness. They came down out of their forts and hideouts, and took over the lands in the south and in the north. They took control of the lands that the Nephites had left, and the cities that the Nephites had left empty. 
%3 Nephi 4:2
\bverse \iffalse But behold, there were no wild beasts nor game in those lands which had been deserted by the Nephites, and there was no game for the robbers save it were in the wilderness. \fi
But there were no wild beasts in the lands that the Nephites had left. The wilderness was the only place where the robbers could find animals to hunt.
%3 Nephi 4:3
\bverse \iffalse And the robbers could not exist save it were in the wilderness, for the want of food; for the Nephites had left their lands desolate, and had gathered their flocks and their herds and all their substance, and they were in one body. \fi
And because there was no food, the robbers could not survive anywhere but out in the wilderness. The Nephites had left their lands empty, and had gathered their flocks and herds and everything they owned into one group.
%3 Nephi 4:4
\bverse \iffalse Therefore, there was no chance for the robbers to plunder and to obtain food, save it were to come up in open battle against the Nephites; and the Nephites being in one body, and having so great a number, and having reserved for themselves provisions, and horses and cattle, and flocks of every kind, that they might subsist for the space of seven years, in the which time they did hope to destroy the robbers from off the face of the land; and thus the eighteenth year did pass away. \fi
The robbers had no chance to plunder and get food, unless they came out in open battle against the Nephites. And the Nephites were gathered into one very large group. They had saved food and supplies for themselves, and horses and cattle, and flocks of every kind. They had enough to survive for seven years. They hoped that this would be enough time to completely destroy the robbers. And the eighteenth year ended.
%3 Nephi 4:5
\bverse \iffalse And it came to pass that in the nineteenth year Giddianhi found that it was expedient that he should go up to battle against the Nephites, for there was no way that they could subsist save it were to plunder and rob and murder. \fi
And in the nineteenth year Giddianhi realized that he had to go to battle against the Nephites. The robbers had no way to survive except for plundering, robbing, and murdering.
%3 Nephi 4:6
\bverse \iffalse And they durst not spread themselves upon the face of the land insomuch that they could raise grain, lest the Nephites should come upon them and slay them; therefore Giddianhi gave commandment unto his armies that in this year they should go up to battle against the Nephites. \fi
They did not dare to spread out and plant grain and crops, because the Nephites would come and kill them. So this year Giddianhi commanded his armies to go to battle against the Nephites.
%3 Nephi 4:7
\bverse \iffalse And it came to pass that they did come up to battle; and it was in the sixth month; and behold, great and terrible was the day that they did come up to battle; and they were girded about after the manner of robbers; and they had a lamb-skin about their loins, and they were dyed in blood, and their heads were shorn, and they had head-plates upon them; and great and terrible was the appearance of the armies of Giddianhi, because of their armor, and because of their being dyed in blood. \fi
And in the sixth month, they came marching to battle. That day was terrible, and they were dressed and armored like robbers. They had lambskins around their waists that had been dyed in blood. Their heads were shaved, and they had helmets on. The armies of Giddianhi looked horrible and frightening because of the blood and armor that covered them.
%3 Nephi 4:8
\bverse \iffalse And it came to pass that the armies of the Nephites, when they saw the appearance of the army of Giddianhi, had all fallen to the earth, and did lift their cries to the Lord their God, that he would spare them and deliver them out of the hands of their enemies. \fi
And when the armies of Nephites saw Giddianhi's army, they fell to the ground and began to pray to the Lord their God, asking him to save them from their enemies.
%3 Nephi 4:9
\bverse \iffalse And it came to pass that when the armies of Giddianhi saw this they began to shout with a loud voice, because of their joy, for they had supposed that the Nephites had fallen with fear because of the terror of their armies. \fi
When Giddianhi's armies saw this they began to shout for joy, because they thought that the Nephites were terrified and had fallen with fear.
%3 Nephi 4:10
\bverse \iffalse But in this thing they were disappointed, for the Nephites did not fear them; but they did fear their God and did supplicate him for protection; therefore, when the armies of Giddianhi did rush upon them they were prepared to meet them; yea, in the strength of the Lord they did receive them. \fi
But this was a disappointment for them, because the Nephites were not afraid of them. They feared God, and begged God for protection. And when Giddianhi's armies rushed to attack them, they were prepared. They had the strength of the Lord with them.
%3 Nephi 4:11
\bverse \iffalse And the battle commenced in this the sixth month; and great and terrible was the battle thereof, yea, great and terrible was the slaughter thereof, insomuch that there never was known so great a slaughter among all the people of Lehi since he left Jerusalem. \fi
And the battle began in the sixth month. The battle and deaths were horrible. There had never been such a terrible battle or so much death since the time that Sariah and Lehi left Jerusalem.
%3 Nephi 4:12
\bverse \iffalse And notwithstanding the threatenings and the oaths which Giddianhi had made, behold, the Nephites did beat them, insomuch that they did fall back from before them. \fi
And even though Giddianhi had threatened them, the Nephites beat the robbers, and the robbers began to retreat.
%3 Nephi 4:13
\bverse \iffalse And it came to pass that Gidgiddoni commanded that his armies should pursue them as far as the borders of the wilderness, and that they should not spare any that should fall into their hands by the way; and thus they did pursue them and did slay them, to the borders of the wilderness, even until they had fulfilled the commandment of Gidgiddoni. \fi
And Gidgiddoni commanded his armies to chase them to the borders of the wilderness, and not spare anyone. They chased them to the borders of the wilderness and killed them, just as Gidgiddoni had commanded them.
%3 Nephi 4:14
\bverse \iffalse And it came to pass that Giddianhi, who had stood and fought with boldness, was pursued as he fled; and being weary because of his much fighting he was overtaken and slain. And thus was the end of Giddianhi the robber. \fi
And Giddianhi, who had fought boldly, was being chased as he ran away. He was exhausted from all the fighting, and Gidgiddoni's armies caught up to him and killed him. This was the end of Giddianhi the robber.
%3 Nephi 4:15
\bverse \iffalse And it came to pass that the armies of the Nephites did return again to their place of security. And it came to pass that this nineteenth year did pass away, and the robbers did not come again to battle; neither did they come again in the twentieth year. \fi
And the armies of the Nephites returned to their forts and defenses. And the nineteenth year ended, and the robbers did not come to battle. They did not come in the twentieth year either.
%3 Nephi 4:16
\bverse \iffalse And in the twenty and first year they did not come up to battle, but they came up on all sides to lay siege round about the people of Nephi; for they did suppose that if they should cut off the people of Nephi from their lands, and should hem them in on every side, and if they should cut them off from all their outward privileges, that they could cause them to yield themselves up according to their wishes. \fi
And in the twenty-first year they did not come to battle. But they came and surrounded the Nephites on every side. They thought that if they kept the Nephites from being able to go to their lands or leave the group, they could get the Nephites to give themselves up.
%3 Nephi 4:17
\bverse \iffalse Now they had appointed unto themselves another leader, whose name was Zemnarihah; therefore it was Zemnarihah that did cause that this siege should take place. \fi
The robbers had appointed another leader who was named Zemnarihah. It was Zemnarihah that had the robbers surround the Nephites.
%3 Nephi 4:18
\bverse \iffalse But behold, this was an advantage to the Nephites; for it was impossible for the robbers to lay siege sufficiently long to have any effect upon the Nephites, because of their much provision which they had laid up in store, \fi
But this was an advantage for the Nephites. It was impossible for the robbers to wait long enough for it to have any effect on the Nephites, because the Nephites had plenty of food and supplies that they had stored.
%3 Nephi 4:19
\bverse \iffalse And because of the scantiness of provisions among the robbers--for behold, they had nothing save it were meat for their subsistence, which meat they did obtain in the wilderness; \fi
And the robbers had very few supplies and food. They had nothing but meat to live on, and they got their meat by hunting in the wilderness.
%3 Nephi 4:20
\bverse \iffalse And it came to pass that the wild game became scarce in the wilderness--insomuch that the robbers were about to perish with hunger. \fi
There began to be very few animals that they could hunt in the wilderness, and the robbers were about to starve to death.
%3 Nephi 4:21
\bverse \iffalse And the Nephites were continually marching out by day and by night, and falling upon their armies, and cutting them off by thousands and by tens of thousands. \fi
And the Nephites marched out during the day and at night. They attacked the armies of the robbers, and killed thousands and tens of thousands of them.
%3 Nephi 4:22
\bverse \iffalse And thus it became the desire of the people of Zemnarihah to withdraw from their design, because of the great destruction which came upon them by night and by day. \fi
Zemnarihah's people wanted to abandon their plan, because they were being killed day and night.
%3 Nephi 4:23
\bverse \iffalse And it came to pass that Zemnarihah did give command unto his people that they should withdraw themselves from the siege, and march into the furthermost parts of the land northward. \fi
And Zemnarihah commanded his people to stop surrounding the Nephites and march into the farthest parts of the land in the north.
%3 Nephi 4:24
\bverse \iffalse And now, Gidgiddoni being aware of their design, and knowing of their weakness because of the want of food, and the great slaughter which had been made among them, therefore he did send out his armies in the night-time, and did cut off the way of their retreat, and did place his armies in the way of their retreat. \fi
Gidgiddoni realized that this was their plan, and he knew that they were weak because they had little food, and many of them had been killed. So he sent out his armies at night, and had them get ahead of the armies of robbers and cut them off before they could retreat to the north.
%3 Nephi 4:25
\bverse \iffalse And this did they do in the night-time, and got on their march beyond the robbers, so that on the morrow, when the robbers began their march, they were met by the armies of the Nephites both in their front and in their rear. \fi
They did this, and they got past the robbers. And the next day, when the robbers began their march, they ran into the armies of the Nephites both in front of them and from behind.
%3 Nephi 4:26
\bverse \iffalse And the robbers who were on the south were also cut off in their places of retreat. And all these things were done by command of Gidgiddoni. \fi
And the robbers who were to the south were also cut off before they could retreat. And it was Gidgiddoni who commanded all of this to happen.
%3 Nephi 4:27
\bverse \iffalse And there were many thousands who did yield themselves up prisoners unto the Nephites, and the remainder of them were slain. \fi
And there were many thousands who gave themselves up and became the prisoners of the Nephites. And the rest of them were killed.
%3 Nephi 4:28
\bverse \iffalse And their leader, Zemnarihah, was taken and hanged upon a tree, yea, even upon the top thereof until he was dead. And when they had hanged him until he was dead they did fell the tree to the earth, and did cry with a loud voice, saying: \fi
And their leader, Zemnarihah, was taken and hanged on a tree. And after they hanged him they cut the tree down and cried out with a loud voice, saying:
%3 Nephi 4:29
\bverse \iffalse May the Lord preserve his people in righteousness and in holiness of heart, that they may cause to be felled to the earth all who shall seek to slay them because of power and secret combinations, even as this man hath been felled to the earth. \fi
``May the Lord protect his people when they are righteous and their hearts are pure. May his people cut down to the ground anyone who tries to kill them for power using secret societies and gangs, just as this man has been cut down to the ground.''
%3 Nephi 4:30
\bverse \iffalse And they did rejoice and cry again with one voice, saying: May the God of Abraham, and the God of Isaac, and the God of Jacob, protect this people in righteousness, so long as they shall call on the name of their God for protection. \fi
And they rejoiced and cried out with one voice, saying: ``May the God of Sarah and Abraham, and Rebekah and Isaac, and Rachel and Jacob, protect this people, as long as they live righteously and call on God for protection.''
%3 Nephi 4:31
\bverse \iffalse And it came to pass that they did break forth, all as one, in singing, and praising their God for the great thing which he had done for them, in preserving them from falling into the hands of their enemies. \fi
And they all sang and praised their God for the great things that God had done for them, protecting them and saving them from their enemies.
%3 Nephi 4:32
\bverse \iffalse Yea, they did cry: Hosanna to the Most High God. And they did cry: Blessed be the name of the Lord God Almighty, the Most High God. \fi
They cried out: ``Praise to the Most High God! Blessed be the name of the Lord God Almighty, the Most High God!''
%3 Nephi 4:33
\bverse \iffalse And their hearts were swollen with joy, unto the gushing out of many tears, because of the great goodness of God in delivering them out of the hands of their enemies; and they knew it was because of their repentance and their humility that they had been delivered from an everlasting destruction. \fi
And their hearts were filled with joy. They cried because of God's great goodness in saving them from their enemies. And they knew God had saved them from neverending destruction because they had repented and become humble.
%Chapter 5
\bchapter
%3 Nephi 5:1
\bverse \iffalse And now behold, there was not a living soul among all the people of the Nephites who did doubt in the least the words of all the holy prophets who had spoken; for they knew that it must needs be that they must be fulfilled. \fi
And there was not a single Nephite who doubted the words that all of the holy prophets had spoken. They knew that all of those words would come true.
%3 Nephi 5:2
\bverse \iffalse And they knew that it must be expedient that Christ had come, because of the many signs which had been given, according to the words of the prophets; and because of the things which had come to pass already they knew that it must needs be that all things should come to pass according to that which had been spoken. \fi
They knew that Christ must come because of the many signs that had been given, just as the prophets had said. And because of the things that had already happened, they knew that all the things that the scriptures said would happen would come true.
%3 Nephi 5:3
\bverse \iffalse Therefore they did forsake all their sins, and their abominations, and their whoredoms, and did serve God with all diligence day and night. \fi
So they left their lives of sin and evil, and they worked very hard to serve God day and night.
%3 Nephi 5:4
\bverse \iffalse And now it came to pass that when they had taken all the robbers prisoners, insomuch that none did escape who were not slain, they did cast their prisoners into prison, and did cause the word of God to be preached unto them; and as many as would repent of their sins and enter into a covenant that they would murder no more were set at liberty. \fi
Any of the robbers whom they had taken prisoner, who tried to escape, were killed. And they took their prisoners and put them in prison, and had the word of God preached to them. And all of them who repented of their sins and made a promise to never murder again were released and made free.
%3 Nephi 5:5
\bverse \iffalse But as many as there were who did not enter into a covenant, and who did still continue to have those secret murders in their hearts, yea, as many as were found breathing out threatenings against their brethren were condemned and punished according to the law. \fi
But any of them who would not make this promise, who kept wanting to commit murder in secret and were heard making threats, were punished according to the law.
%3 Nephi 5:6
\bverse \iffalse And thus they did put an end to all those wicked, and secret, and abominable combinations, in the which there was so much wickedness, and so many murders committed. \fi
So they put an end to all those wicked, secret, horrible societies that had caused so much evil and so many murders.
%3 Nephi 5:7
\bverse \iffalse And thus had the twenty and second year passed away, and the twenty and third year also, and the twenty and fourth, and the twenty and fifth; and thus had twenty and five years passed away. \fi
And the twenty-second year ended; so did the twenty-third, twenty-fourth, and twenty-fifth years. And now twenty-five years had gone by.
%3 Nephi 5:8
\bverse \iffalse And there had many things transpired which, in the eyes of some, would be great and marvelous; nevertheless, they cannot all be written in this book; yea, this book cannot contain even a hundredth part of what was done among so many people in the space of twenty and five years; \fi
And many things had happened that some people would say were great and remarkable. But they cannot all be written in this book; this book cannot even contain a small part of the things that had happened among so many people in twenty-five years.
%3 Nephi 5:9
\bverse \iffalse But behold there are records which do contain all the proceedings of this people; and a shorter but true account was given by Nephi. \fi
But there are records that tell all the things that happened among this people. A shorter version of all of these things that happened was made by Nephi.
%3 Nephi 5:10
\bverse \iffalse Therefore I have made my record of these things according to the record of Nephi, which was engraven on the plates which were called the plates of Nephi. \fi
So I have used the record that Nephi kept to make my record of these things. His record was engraved on the plates that are called the plates of Nephi.
%3 Nephi 5:11
\bverse \iffalse And behold, I do make the record on plates which I have made with mine own hands. \fi
And I am making my record on plates that I have made with my own hands.
%3 Nephi 5:12
\bverse \iffalse And behold, I am called Mormon, being called after the land of Mormon, the land in which Alma did establish the church among the people, yea, the first church which was established among them after their transgression. \fi
My name is Mormon. I am named after the land of Mormon---the land where Alma set up the church among the people. It was the first church set up among them after they sinned.
%3 Nephi 5:13
\bverse \iffalse Behold, I am a disciple of Jesus Christ, the Son of God. I have been called of him to declare his word among his people, that they might have everlasting life. \fi
I am a disciple of Jesus Christ, the Son of God. I have been called by him to declared his words among his people, so that they can have eternal life.
%3 Nephi 5:14
\bverse \iffalse And it hath become expedient that I, according to the will of God, that the prayers of those who have gone hence, who were the holy ones, should be fulfilled according to their faith, should make a record of these things which have been done-- \fi
And I must make this record according to God's will, so that the prayers of the people who have died, who were holy, can be answered according to their faith.
%3 Nephi 5:15
\bverse \iffalse Yea, a small record of that which hath taken place from the time that Lehi left Jerusalem, even down until the present time. \fi
It is a small record of the things that have happened since Sariah and Lehi left Jerusalem until now.
%3 Nephi 5:16
\bverse \iffalse Therefore I do make my record from the accounts which have been given by those who were before me, until the commencement of my day; \fi
And I make my record based on the records that were kept by the people who came before me, until the time that I began to keep the record myself.
%3 Nephi 5:17
\bverse \iffalse And then I do make a record of the things which I have seen with mine own eyes. \fi
And then I keep a record of the things that I have seen with my own eyes.
%3 Nephi 5:18
\bverse \iffalse And I know the record which I make to be a just and a true record; nevertheless there are many things which, according to our language, we are not able to write. \fi
I know that the record that I am keeping is fair and true. But there are many things that we cannot write in our language.
%3 Nephi 5:19
\bverse \iffalse And now I make an end of my saying, which is of myself, and proceed to give my account of the things which have been before me. \fi
And now I will continue to tell the story of the things that have happened before.
%3 Nephi 5:20
\bverse \iffalse I am Mormon, and a pure descendant of Lehi. I have reason to bless my God and my Savior Jesus Christ, that he brought our fathers out of the land of Jerusalem, (and no one knew it save it were himself and those whom he brought out of that land) and that he hath given me and my people so much knowledge unto the salvation of our souls. \fi
I am Mormon, a descendant of Sariah and Lehi. I have good reason to praise my God and my Savior Jesus Christ, who brought our ancestors out of the land of Jerusalem. No one knew about this besides God and the people that God brought out of that land. And God has given me and my people knowledge and understanding so that our souls can be saved.
%3 Nephi 5:21
\bverse \iffalse Surely he hath blessed the house of Jacob, and hath been merciful unto the seed of Joseph. \fi
God has blessed the descendants of Jacob, and has shown mercy to the descendants of Asenath and Joseph.
%3 Nephi 5:22
\bverse \iffalse And insomuch as the children of Lehi have kept his commandments he hath blessed them and prospered them according to his word. \fi
And when the children of Sariah and Lehi have kept the Lord's commandments, the Lord has blessed them and given them all that they need, just as he promised.
%3 Nephi 5:23
\bverse \iffalse Yea, and surely shall he again bring a remnant of the seed of Joseph to the knowledge of the Lord their God. \fi
And he will surely bring a part of the descendants of Asenath and Joseph to know about the Lord their God.
%3 Nephi 5:24
\bverse \iffalse And as surely as the Lord liveth, will he gather in from the four quarters of the earth all the remnant of the seed of Jacob, who are scattered abroad upon all the face of the earth. \fi
And as surely as the Lord lives he will gather all of the descendants of Jacob, who are scattered over the whole earth, from every corner of the earth.
%3 Nephi 5:25
\bverse \iffalse And as he hath covenanted with all the house of Jacob, even so shall the covenant wherewith he hath covenanted with the house of Jacob be fulfilled in his own due time, unto the restoring all the house of Jacob unto the knowledge of the covenant that he hath covenanted with them. \fi
He has made promises to the descendants of Jacob. And when he decides the time is right he will keep his promises to them, and all the descendants of Jacob will know about the promises that he has made to them.
%3 Nephi 5:26
\bverse \iffalse And then shall they know their Redeemer, who is Jesus Christ, the Son of God; and then shall they be gathered in from the four quarters of the earth unto their own lands, from whence they have been dispersed; yea, as the Lord liveth so shall it be. Amen. \fi
Then they will know that their Redeemer, who is Jesus Christ, the Son of God. Then they will be gathered in from every corner of the earth back to their own lands. And this will happen as surely as the Lord lives. Amen.
%Chapter 6
\bchapter
%3 Nephi 6:1
\bverse \iffalse And now it came to pass that the people of the Nephites did all return to their own lands in the twenty and sixth year, every man, with his family, his flocks and his herds, his horses and his cattle, and all things whatsoever did belong unto them. \fi
And now, the Nephites all went back to their own lands in the twenty-sixth year. Every family took their flocks and herds, and horses and cattle, and all the things that belonged to them.
%3 Nephi 6:2
\bverse \iffalse And it came to pass that they had not eaten up all their provisions; therefore they did take with them all that they had not devoured, of all their grain of every kind, and their gold, and their silver, and all their precious things, and they did return to their own lands and their possessions, both on the north and on the south, both on the land northward and on the land southward. \fi
They had not eaten up all of their supplies, so they took with them all the supplies that they had not used---their grains, gold, silver, and precious things---and went back to their own lands, both in the north and in the south.
%3 Nephi 6:3
\bverse \iffalse And they granted unto those robbers who had entered into a covenant to keep the peace of the land, who were desirous to remain Lamanites, lands, according to their numbers, that they might have, with their labors, wherewith to subsist upon; and thus they did establish peace in all the land. \fi
And they let those robbers who had a made a promise to keep peace, who wanted to be Lamanites, have their own lands so that they could work and live. This is how they established peace in the land.
%3 Nephi 6:4
\bverse \iffalse And they began again to prosper and to wax great; and the twenty and sixth and seventh years passed away, and there was great order in the land; and they had formed their laws according to equity and justice. \fi
And they once again began to have plenty, and there began to be more and more of them. And the twenty-sixth and twenty-seventh years ended, and there was great peace in the land. And their laws were equal and fair.
%3 Nephi 6:5
\bverse \iffalse And now there was nothing in all the land to hinder the people from prospering continually, except they should fall into transgression. \fi
And there was nothing in all the land to keep the people from growing and living happily, unless they started to live sinfully.
%3 Nephi 6:6
\bverse \iffalse And now it was Gidgiddoni, and the judge, Lachoneus, and those who had been appointed leaders, who had established this great peace in the land. \fi
And it was Gidgiddoni, and the judge, Lachoneus, and those who had been made leaders, who had created peace in the land.
%3 Nephi 6:7
\bverse \iffalse And it came to pass that there were many cities built anew, and there were many old cities repaired. \fi
Many new cities were built, and many old cities were repaired.
%3 Nephi 6:8
\bverse \iffalse And there were many highways cast up, and many roads made, which led from city to city, and from land to land, and from place to place. \fi
Many highways and roads were made. They led from city to city, and from land to land, and from place to place.
%3 Nephi 6:9
\bverse \iffalse And thus passed away the twenty and eighth year, and the people had continual peace. \fi
And the twenty-eighth year ended, and the people lived in peace.
%3 Nephi 6:10
\bverse \iffalse But it came to pass in the twenty and ninth year there began to be some disputings among the people; and some were lifted up unto pride and boastings because of their exceedingly great riches, yea, even unto great persecutions; \fi
But in the twenty-ninth year there began to be some problems among the people. Some of them began to be proud and to brag, and even bully others, because they had very much money and many possessions.
%3 Nephi 6:11
\bverse \iffalse For there were many merchants in the land, and also many lawyers, and many officers. \fi
And there were many sellers, traders, lawyers, and government officials in the land.
%3 Nephi 6:12
\bverse \iffalse And the people began to be distinguished by ranks, according to their riches and their chances for learning, yea, some were ignorant because of their poverty, and others did receive great learning because of their riches. \fi
And the people began to separate into ranks, according to how much money and education they had or could get. Some people were uneducated, because they were poor. And other people were very educated because of their money.
%3 Nephi 6:13
\bverse \iffalse Some were lifted up in pride, and others were exceedingly humble; some did return railing for railing, while others would receive railing and persecution and all manner of afflictions, and would not turn and revile again, but were humble and penitent before God. \fi
Some were very proud, while others were very humble. Some would fight back and bully when they were bullied, but others took the bullying and unkindness, and would not hit or bully back. They were humble and remembered God.
%3 Nephi 6:14
\bverse \iffalse And thus there became a great inequality in all the land, insomuch that the church began to be broken up; yea, insomuch that in the thirtieth year the church was broken up in all the land save it were among a few of the Lamanites who were converted unto the true faith; and they would not depart from it, for they were firm, and steadfast, and immovable, willing with all diligence to keep the commandments of the Lord. \fi
So there began to be great inequality in the land, and the church began to be divided. And in the thirtieth year the church members were no longer united, except for among a few Lamanites who were converted to the true faith. They would not leave the true faith; they were firm and true, and they were willing to do everything they could to keep the Lord's commandments.
%3 Nephi 6:15
\bverse \iffalse Now the cause of this iniquity of the people was this--Satan had great power, unto the stirring up of the people to do all manner of iniquity, and to the puffing them up with pride, tempting them to seek for power, and authority, and riches, and the vain things of the world. \fi
And the cause of this evil was this: Satan had gotten great power, and had gotten the people to do all sorts of evil things. They were puffed up with pride, and Satan tempted them to try to get power and authority, and money, and the empty things of the world.
%3 Nephi 6:16
\bverse \iffalse And thus Satan did lead away the hearts of the people to do all manner of iniquity; therefore they had enjoyed peace but a few years. \fi
Satan led the hearts of the people to commit all kinds of sins. And they had only had peace for a few years when this happened.
%3 Nephi 6:17
\bverse \iffalse And thus, in the commencement of the thirtieth year--the people having been delivered up for the space of a long time to be carried about by the temptations of the devil whithersoever he desired to carry them, and to do whatsoever iniquity he desired they should--and thus in the commencement of this, the thirtieth year, they were in a state of awful wickedness. \fi
The people had been doing evil things and listening to the temptations of the devil for a long time now. He got them to do whatever evil thing he wanted them to do. And in the beginning of the thirtieth year they were living very sinfully.
%3 Nephi 6:18
\bverse \iffalse Now they did not sin ignorantly, for they knew the will of God concerning them, for it had been taught unto them; therefore they did wilfully rebel against God. \fi
They sinned even though they knew better. They knew what God wanted them to do, because it had been taught to them. But they chose themselves to rebel against God.
%3 Nephi 6:19
\bverse \iffalse And now it was in the days of Lachoneus, the son of Lachoneus, for Lachoneus did fill the seat of his father and did govern the people that year. \fi
And this happened while Lachoneus, the son of Lachoneus, was the chief judge. He became the chief judge after his father, Lachoneus.
%3 Nephi 6:20
\bverse \iffalse And there began to be men inspired from heaven and sent forth, standing among the people in all the land, preaching and testifying boldly of the sins and iniquities of the people, and testifying unto them concerning the redemption which the Lord would make for his people, or in other words, the resurrection of Christ; and they did testify boldly of his death and sufferings. \fi
And there began to be people who were inspired from heaven to come out among the people, and preach and testify boldly about the sins that the people were committing. They testified to them about how Christ would redeem his people, and be resurrected. They testified boldly about how Christ was going to die and suffer.
%3 Nephi 6:21
\bverse \iffalse Now there were many of the people who were exceedingly angry because of those who testified of these things; and those who were angry were chiefly the chief judges, and they who had been high priests and lawyers; yea, all those who were lawyers were angry with those who testified of these things. \fi
Many of the people were furious at the people who testified about these things. The ones who were angry were mostly the chief judges, high priests, and lawyers. In fact, every lawyer was angry at the people who testified about these things.
%3 Nephi 6:22
\bverse \iffalse Now there was no lawyer nor judge nor high priest that could have power to condemn any one to death save their condemnation was signed by the governor of the land. \fi
Now, the lawyers, judges, and high priests did not have power to put anyone to death, unless they had gotten a signature from the governor of the land.
%3 Nephi 6:23
\bverse \iffalse Now there were many of those who testified of the things pertaining to Christ who testified boldly, who were taken and put to death secretly by the judges, that the knowledge of their death came not unto the governor of the land until after their death. \fi
But many of the people who testified boldly about Christ were secretly taken and put to death; and the governor of the land did not find out about their death until after they had died.
%3 Nephi 6:24
\bverse \iffalse Now behold, this was contrary to the laws of the land, that any man should be put to death except they had power from the governor of the land-- \fi
But it was against the laws of the land to put a person to death without permission from the governor.
%3 Nephi 6:25
\bverse \iffalse Therefore a complaint came up unto the land of Zarahemla, to the governor of the land, against these judges who had condemned the prophets of the Lord unto death, not according to the law. \fi
So a complaint against these judges who had broken the law and put the prophets of the Lord to death was sent to the governor in the land of Zarahemla.
%3 Nephi 6:26
\bverse \iffalse Now it came to pass that they were taken and brought up before the judge to be judged of the crime which they had done, according to the law which had been given by the people. \fi
And those judges were taken and brought before the judge, to be judged according to the laws of the people for the crimes they committed.
%3 Nephi 6:27
\bverse \iffalse Now it came to pass that those judges had many friends and kindreds; and the remainder, yea, even almost all the lawyers and the high priests, did gather themselves together, and unite with the kindreds of those judges who were to be tried according to the law. \fi
Those judges had many friends. And the rest---almost all of the lawyers and high priests---gathered together and joined with the friends of those judges who were going to be punished according to the law.
%3 Nephi 6:28
\bverse \iffalse And they did enter into a covenant one with another, yea, even into that covenant which was given by them of old, which covenant was given and administered by the devil, to combine against all righteousness. \fi
And they made a secret promise with each other. It was the secret promise from ancient times that came from the devil, and was spread by the devil. It was a promise to fight against everything and everyone that is righteous.
%3 Nephi 6:29
\bverse \iffalse Therefore they did combine against the people of the Lord, and enter into a covenant to destroy them, and to deliver those who were guilty of murder from the grasp of justice, which was about to be administered according to the law. \fi
So they combined against the people of the Lord, and made a secret promise to destroy them, and to save the judges who were guilty, who were about to be punished.
%3 Nephi 6:30
\bverse \iffalse And they did set at defiance the law and the rights of their country; and they did covenant one with another to destroy the governor, and to establish a king over the land, that the land should no more be at liberty but should be subject unto kings. \fi
And they rejected the law and the rights of their country. They made a promise and a plan with each other to kill the governor and put a king over the land, so that instead of being a free land, it would be a land where kings ruled.
%Chapter 7
\bchapter
%3 Nephi 7:1
\bverse \iffalse Now behold, I will show unto you that they did not establish a king over the land; but in this same year, yea, the thirtieth year, they did destroy upon the judgment seat, yea, did murder the chief judge of the land. \fi
I will show you that they did not manage to put a king over the land, but in this same year, the thirtieth year, they murdered the chief judge.
%3 Nephi 7:2
\bverse \iffalse And the people were divided one against another; and they did separate one from another into tribes, every man according to his family and his kindred and friends; and thus they did destroy the government of the land. \fi
The people were divided against each other, and they separated into tribes. Every family separated into groups with their friends and the people they knew. This destroyed the government of the land.
%3 Nephi 7:3
\bverse \iffalse And every tribe did appoint a chief or a leader over them; and thus they became tribes and leaders of tribes. \fi
Every tribe chose a chief or a leader over them. And now the people were divided into tribes and leaders of tribes.
%3 Nephi 7:4
\bverse \iffalse Now behold, there was no man among them save he had much family and many kindreds and friends; therefore their tribes became exceedingly great. \fi
And all of the people had many family and friends, so their tribes became very large.
%3 Nephi 7:5
\bverse \iffalse Now all this was done, and there were no wars as yet among them; and all this iniquity had come upon the people because they did yield themselves unto the power of Satan. \fi
All of this happened, but they had not yet had any wars. And all of this evil happened because the people allowed themselves to be tempted by Satan and let him take control.
%3 Nephi 7:6
\bverse \iffalse And the regulations of the government were destroyed, because of the secret combination of the friends and kindreds of those who murdered the prophets. \fi
And the laws and order of the government were destroyed, because of the secret society of the friends of the people who murdered the prophets.
%3 Nephi 7:7
\bverse \iffalse And they did cause a great contention in the land, insomuch that the more righteous part of the people had nearly all become wicked; yea, there were but few righteous men among them. \fi
They caused great arguments and problems in the land, so that most of the righteous people had also become wicked. And there were only a few righteous people left among them.
%3 Nephi 7:8
\bverse \iffalse And thus six years had not passed away since the more part of the people had turned from their righteousness, like the dog to his vomit, or like the sow to her wallowing in the mire. \fi
And in less than six years, most of the people had turned away from doing what was right, like a dog going back to its vomit, or a sheep going back to the swamp.
%3 Nephi 7:9
\bverse \iffalse Now this secret combination, which had brought so great iniquity upon the people, did gather themselves together, and did place at their head a man whom they did call Jacob; \fi
And this secret society, which had brought so much evil among the people, gathered together and made a man called Jacob their leader.
%3 Nephi 7:10
\bverse \iffalse And they did call him their king; therefore he became a king over this wicked band; and he was one of the chiefest who had given his voice against the prophets who testified of Jesus. \fi
They called him their king. So he became the king over this wicked gang. And he was one of the people who spoke loudly against the prophets who testified of Jesus.
%3 Nephi 7:11
\bverse \iffalse And it came to pass that they were not so strong in number as the tribes of the people, who were united together save it were their leaders did establish their laws, every one according to his tribe; nevertheless they were enemies; notwithstanding they were not a righteous people, yet they were united in the hatred of those who had entered into a covenant to destroy the government. \fi
There were not as many of them as there were people who belonged to the tribes of the people. The people who separated into tribes were united, except that each tribe had its own laws that were made by its leaders. And even though these people were not righteous, they were the enemies of the secret society. All of them hated the people who had made a promise to destroy the government.
%3 Nephi 7:12
\bverse \iffalse Therefore, Jacob seeing that their enemies were more numerous than they, he being the king of the band, therefore he commanded his people that they should take their flight into the northernmost part of the land, and there build up unto themselves a kingdom, until they were joined by dissenters, (for he flattered them that there would be many dissenters) and they become sufficiently strong to contend with the tribes of the people; and they did so. \fi
And Jacob, the king of the robbers, saw that they were outnumbered by their enemies. So he commanded his people to run away into the most northern part of the land and build a kingdom for themselves, until they were joined by others who wanted to leave the Nephites. He told his people that many people would leave the Nephites, and they would become strong enough to battle the tribes of the people. And this is what they did.
%3 Nephi 7:13
\bverse \iffalse And so speedy was their march that it could not be impeded until they had gone forth out of the reach of the people. And thus ended the thirtieth year; and thus were the affairs of the people of Nephi. \fi
And they marched so quickly that they could not be stopped. They marched so far that they could not be caught. And the thirtieth year of the judges ended.
%3 Nephi 7:14
\bverse \iffalse And it came to pass in the thirty and first year that they were divided into tribes, every man according to his family, kindred and friends; nevertheless they had come to an agreement that they would not go to war one with another; but they were not united as to their laws, and their manner of government, for they were established according to the minds of those who were their chiefs and their leaders. But they did establish very strict laws that one tribe should not trespass against another, insomuch that in some degree they had peace in the land; nevertheless, their hearts were turned from the Lord their God, and they did stone the prophets and did cast them out from among them. \fi
And in the thirty-first year the people were divided into tribes with their family and friends. The tribes made an agreement that they would not go to war with each other. But they all had different laws and governments, because their chiefs and leaders decided the laws. But they all had strict laws that one tribe should not cause problems with another tribe. These laws made some peace among the tribes, but their hearts were turned away from the Lord their God, and they stoned the prophets and threw them out.
%3 Nephi 7:15
\bverse \iffalse And it came to pass that Nephi--having been visited by angels and also the voice of the Lord, therefore having seen angels, and being eye-witness, and having had power given unto him that he might know concerning the ministry of Christ, and also being eye-witness to their quick return from righteousness unto their wickedness and abominations; \fi
And Nephi was visited by angels, and the voice of the Lord spoke to him. He had seen angels with his own eyes, and he had power given to him from God, so that he knew about Christ's mission and his preaching. He saw for himself how the people had quickly turned from living righteously to living in sin and great evil.
%3 Nephi 7:16
\bverse \iffalse Therefore, being grieved for the hardness of their hearts and the blindness of their minds--went forth among them in that same year, and began to testify, boldly, repentance and remission of sins through faith on the Lord Jesus Christ. \fi
He was very sad that their hearts were so hard and their minds were so blind. So he went out among the people that year, and began to teach and testify boldly about repentance and the forgiveness of sins through faith in the Lord Jesus Christ.
%3 Nephi 7:17
\bverse \iffalse And he did minister many things unto them; and all of them cannot be written, and a part of them would not suffice, therefore they are not written in this book. And Nephi did minister with power and with great authority. \fi
He ministered and taught many things to them. And it is impossible to write everything he did, and it would not be enough to write only a part of what he did. So these things are not written in this book. And Nephi ministered and taught with power and great authority.
%3 Nephi 7:18
\bverse \iffalse And it came to pass that they were angry with him, even because he had greater power than they, for it were not possible that they could disbelieve his words, for so great was his faith on the Lord Jesus Christ that angels did minister unto him daily. \fi
And they were angry with him, because he had even more power than they did. The things he said were so powerful that it was impossible for them to disbelieve. His faith in the Lord Jesus Christ was so strong that angels visited him and spoke to him every day.
%3 Nephi 7:19
\bverse \iffalse And in the name of Jesus did he cast out devils and unclean spirits; and even his brother did he raise from the dead, after he had been stoned and suffered death by the people. \fi
And in the name of Jesus he forced demons and evil spirits out of people. He even raised his brother from the dead, after he had been stoned to death by the people.
%3 Nephi 7:20
\bverse \iffalse And the people saw it, and did witness of it, and were angry with him because of his power; and he did also do many more miracles, in the sight of the people, in the name of Jesus. \fi
And the people saw all of this and knew that it had happened. They were angry with him because of his power. He did many miracles in the name of Jesus; he did these miracles in places where the people saw that he was doing them.
%3 Nephi 7:21
\bverse \iffalse And it came to pass that the thirty and first year did pass away, and there were but few who were converted unto the Lord; but as many as were converted did truly signify unto the people that they had been visited by the power and Spirit of God, which was in Jesus Christ, in whom they believed. \fi
And the thirty-first year ended, and only a few people were converted to the Lord. But anyone who was converted showed the people by their actions that they had been visited by God's Spirit and power, which was in Jesus Christ. Jesus Christ was the one they believed in.
%3 Nephi 7:22
\bverse \iffalse And as many as had devils cast out from them, and were healed of their sicknesses and their infirmities, did truly manifest unto the people that they had been wrought upon by the Spirit of God, and had been healed; and they did show forth signs also and did do some miracles among the people. \fi
And anyone who had devils forced out of them, and were healed of their sickness and health problems, truly showed the people that it was the Spirit of God that had healed them. They showed signs and did some miracles among the people.
%3 Nephi 7:23
\bverse \iffalse Thus passed away the thirty and second year also. And Nephi did cry unto the people in the commencement of the thirty and third year; and he did preach unto them repentance and remission of sins. \fi
And the thirty-second year also ended. And in the beginning of the thirty-third year Nephi cried out to the people, and preached to them about repentance and having their sins forgiven.
%3 Nephi 7:24
\bverse \iffalse Now I would have you to remember also, that there were none who were brought unto repentance who were not baptized with water. \fi
Now I want you to remember that all of the people who repented were baptized with water.
%3 Nephi 7:25
\bverse \iffalse Therefore, there were ordained of Nephi, men unto this ministry, that all such as should come unto them should be baptized with water, and this as a witness and a testimony before God, and unto the people, that they had repented and received a remission of their sins. \fi
And Nephi ordained others to help with the work of God, so that anyone who came to them could be baptized with water, and show a witness and testimony to God, and to the people, that they had repented and been forgiven for their sins.
%3 Nephi 7:26
\bverse \iffalse And there were many in the commencement of this year that were baptized unto repentance; and thus the more part of the year did pass away. \fi
And there were many people in the beginning of this year, and throughout most of this year, who repented and were baptized.
%Chapter 8
\bchapter
%3 Nephi 8:1
\bverse \iffalse And now it came to pass that according to our record, and we know our record to be true, for behold, it was a just man who did keep the record--for he truly did many miracles in the name of Jesus; and there was not any man who could do a miracle in the name of Jesus save he were cleansed every whit from his iniquity-- \fi
We know that the record we have is true, because it was kept by a righteous man. We know that he was a righteous man, because he did many miracles in the name of Jesus. And no one could do a miracle in the name of Jesus unless they were completely clean from sin and evil.
%3 Nephi 8:2
\bverse \iffalse And now it came to pass, if there was no mistake made by this man in the reckoning of our time, the thirty and third year had passed away; \fi
And if this man did not make a mistake in his record, thirty-three years had gone by.
%3 Nephi 8:3
\bverse \iffalse And the people began to look with great earnestness for the sign which had been given by the prophet Samuel, the Lamanite, yea, for the time that there should be darkness for the space of three days over the face of the land. \fi
And the people began to look for the sign that the prophet Samuel, a Lamanite, had prophesied---the time when there would be darkness for three days over the land. 
%3 Nephi 8:4
\bverse \iffalse And there began to be great doubtings and disputations among the people, notwithstanding so many signs had been given. \fi
And the people started to doubt and argue with each other, even though they had already seen so many other signs.
%3 Nephi 8:5
\bverse \iffalse And it came to pass in the thirty and fourth year, in the first month, on the fourth day of the month, there arose a great storm, such an one as never had been known in all the land. \fi
And on the fourth day of the first month in the thirty-fourth year, a horrible storm came. The people had never seen such a terrible storm before.
%3 Nephi 8:6
\bverse \iffalse And there was also a great and terrible tempest; and there was terrible thunder, insomuch that it did shake the whole earth as if it was about to divide asunder. \fi
There was also a terrible windstorm, and thunder that shook the ground as if it were about to split in half.
%3 Nephi 8:7
\bverse \iffalse And there were exceedingly sharp lightnings, such as never had been known in all the land. \fi
And there was very sharp lightning, worse than any that the people had ever been seen before.
%3 Nephi 8:8
\bverse \iffalse And the city of Zarahemla did take fire. \fi
And the city of Zarahemla caught fire.
%3 Nephi 8:9
\bverse \iffalse And the city of Moroni did sink into the depths of the sea, and the inhabitants thereof were drowned. \fi
And city of Moroni sunk into the sea, and the people who lived there drowned.
%3 Nephi 8:10
\bverse \iffalse And the earth was carried up upon the city of Moronihah that in the place of the city there became a great mountain. \fi
And the earth covered the city of Moronihah, so that there was a great mountain where the city had been.
%3 Nephi 8:11
\bverse \iffalse And there was a great and terrible destruction in the land southward. \fi
And there was terrible destruction in the land to the south.
%3 Nephi 8:12
\bverse \iffalse But behold, there was a more great and terrible destruction in the land northward; for behold, the whole face of the land was changed, because of the tempest and the whirlwinds and the thunderings and the lightnings, and the exceedingly great quaking of the whole earth; \fi
But the destruction was even worse in the land to the north. The whole face of the land was changed, because of the windstorm and whirlwinds, and the thunder and lightning, and the earthquakes.
%3 Nephi 8:13
\bverse \iffalse And the highways were broken up, and the level roads were spoiled, and many smooth places became rough. \fi
The highways were broken up, and flat roads became jagged, and many smooth places became rough.
%3 Nephi 8:14
\bverse \iffalse And many great and notable cities were sunk, and many were burned, and many were shaken till the buildings thereof had fallen to the earth, and the inhabitants thereof were slain, and the places were left desolate. \fi
And many great cities were sunk, and burned, and shaken until the buildings in them had fallen to the ground, and the people who lived there were killed, and the places were left empty and destroyed.
%3 Nephi 8:15
\bverse \iffalse And there were some cities which remained; but the damage thereof was exceedingly great, and there were many of them who were slain. \fi
There were some cities that survived, but all of these were very damaged, and many people who lived in them died.
%3 Nephi 8:16
\bverse \iffalse And there were some who were carried away in the whirlwind; and whither they went no man knoweth, save they know that they were carried away. \fi
Some people were carried away in the whirlwind. No one knew where they were carried to.
%3 Nephi 8:17
\bverse \iffalse And thus the face of the whole earth became deformed, because of the tempests, and the thunderings, and the lightnings, and the quaking of the earth. \fi
The whole face of the earth was changed, because of the windstorms, and thunder, and lightning, and earthquakes.
%3 Nephi 8:18
\bverse \iffalse And behold, the rocks were rent in twain; they were broken up upon the face of the whole earth, insomuch that they were found in broken fragments, and in seams and in cracks, upon all the face of the land. \fi
The rocks were broken up throughout the whole land; they were found cracked, split, and in pieces over the whole land.
%3 Nephi 8:19
\bverse \iffalse And it came to pass that when the thunderings, and the lightnings, and the storm, and the tempest, and the quakings of the earth did cease--for behold, they did last for about the space of three hours; and it was said by some that the time was greater; nevertheless, all these great and terrible things were done in about the space of three hours--and then behold, there was darkness upon the face of the land. \fi
And after about three hours, the thunder, and lightning, and storm, and windstorms, and earthquakes stopped. Some people said that they went on for longer than three hours; but it all actually happened in about three hours. And then darkness covered the land.
%3 Nephi 8:20
\bverse \iffalse And it came to pass that there was thick darkness upon all the face of the land, insomuch that the inhabitants thereof who had not fallen could feel the vapor of darkness; \fi
Thick darkness covered the entire land. The darkness was so thick that the people who had not died could feel the darkness.
%3 Nephi 8:21
\bverse \iffalse And there could be no light, because of the darkness, neither candles, neither torches; neither could there be fire kindled with their fine and exceedingly dry wood, so that there could not be any light at all; \fi
And because of the darkness, it was impossible to see any light. They could not light candles or torches, or start a fire---not even with their best and dryest wood.
%3 Nephi 8:22
\bverse \iffalse And there was not any light seen, neither fire, nor glimmer, neither the sun, nor the moon, nor the stars, for so great were the mists of darkness which were upon the face of the land. \fi
And the mists of darkness that covered the land were so great that no one could see any light, or fire, or sun, or moon, or stars.
%3 Nephi 8:23
\bverse \iffalse And it came to pass that it did last for the space of three days that there was no light seen; and there was great mourning and howling and weeping among all the people continually; yea, great were the groanings of the people, because of the darkness and the great destruction which had come upon them. \fi
For three days no one saw any light. The people cried, and groaned, and mourned without stopping, because of the darkness and terrible destruction that had come to them.
%3 Nephi 8:24
\bverse \iffalse And in one place they were heard to cry, saying: O that we had repented before this great and terrible day, and then would our brethren have been spared, and they would not have been burned in that great city Zarahemla. \fi
In one place the people said: ``O how we wish we had repented before this terrible day. Then our sisters and brothers would not have been killed; they would not have been burned in that great city Zarahemla!''
%3 Nephi 8:25
\bverse \iffalse And in another place they were heard to cry and mourn, saying: O that we had repented before this great and terrible day, and had not killed and stoned the prophets, and cast them out; then would our mothers and our fair daughters, and our children have been spared, and not have been buried up in that great city Moronihah. And thus were the howlings of the people great and terrible. \fi
In another place the people cried and mourned, and said: ``O how we wish we had repented before this terrible day! We wish we had not killed and stoned the prophets, and thrown them out. Then our mothers, and our wives, and our children would not have died. They would not have been buried in that great city Moronihah!'' These are the sorts of things that the people said as they howled and cried.
%Chapter 9
\bchapter
\bverse \iffalse And it came to pass that there was a voice heard among all the inhabitants of the earth, upon all the face of this land, crying: \fi
And then the people throughout the entire land heard a voice that cried out:
%3 Nephi 9:2
\bverse \iffalse Wo, wo, wo unto this people; wo unto the inhabitants of the whole earth except they shall repent; for the devil laugheth, and his angels rejoice, because of the slain of the fair sons and daughters of my people; and it is because of their iniquity and abominations that they are fallen! \fi
``How very terrible it will be for this people! How terrible it will be for everyone who lives on earth, unless they repent! The devil is laughing and his angels are celebrating, because the beautiful children of my people have been killed. And they have died because of the very evil things they did and the horrible sins they committed!
%3 Nephi 9:3
\bverse \iffalse Behold, that great city Zarahemla have I burned with fire, and the inhabitants thereof. \fi
I have burned up that great city, Zarahemla, and all the people living there, with fire.
%3 Nephi 9:4
\bverse \iffalse And behold, that great city Moroni have I caused to be sunk in the depths of the sea, and the inhabitants thereof to be drowned. \fi
I have caused that great city Moroni to sink into the depths of the sea, and the people who lived there to drown.
%3 Nephi 9:5
\bverse \iffalse And behold, that great city Moronihah have I covered with earth, and the inhabitants thereof, to hide their iniquities and their abominations from before my face, that the blood of the prophets and the saints shall not come any more unto me against them. \fi
I have covered that great city Moronihah and the people who lived there with earth, to hide their sins and evil actions from my sight, so that the blood of the prophets and saints that they killed will stop crying out to me for revenge.
%3 Nephi 9:6
\bverse \iffalse And behold, the city of Gilgal have I caused to be sunk, and the inhabitants thereof to be buried up in the depths of the earth; \fi
I have caused the city of Gilgal to sink, and the people who lived there to be swallowed by the earth.
%3 Nephi 9:7
\bverse \iffalse Yea, and the city of Onihah and the inhabitants thereof, and the city of Mocum and the inhabitants thereof, and the city of Jerusalem and the inhabitants thereof; and waters have I caused to come up in the stead thereof, to hide their wickedness and abominations from before my face, that the blood of the prophets and the saints shall not come up any more unto me against them. \fi
I have also done this to the cities of Onihah,  Mocum, and Jerusalem, and to all of the people living in them. I have made the waters come to cover them to hide their wickedness and evil from my sight, so that the blood of the prophets and the saints will stop crying out to me for revenge.
%3 Nephi 9:8
\bverse \iffalse And behold, the city of Gadiandi, and the city of Gadiomnah, and the city of Jacob, and the city of Gimgimno, all these have I caused to be sunk, and made hills and valleys in the places thereof; and the inhabitants thereof have I buried up in the depths of the earth, to hide their wickedness and abominations from before my face, that the blood of the prophets and the saints should not come up any more unto me against them. \fi
And I have made the cities of Gadiandi, and Gadiomnah, and Jacob, and Gimgimno, and all the people who lived in them, sink into the earth. I have put hills and valleys on top of them, to hide their wickedness and evil from my sight, so that the blood of the prophets and the saints will stop crying out to me for revenge.
%3 Nephi 9:9
\bverse \iffalse And behold, that great city Jacobugath, which was inhabited by the people of king Jacob, have I caused to be burned with fire because of their sins and their wickedness, which was above all the wickedness of the whole earth, because of their secret murders and combinations; for it was they that did destroy the peace of my people and the government of the land; therefore I did cause them to be burned, to destroy them from before my face, that the blood of the prophets and the saints should not come up unto me any more against them. \fi
And I have burned that great city Jacobugath, where the people of king Jacob lived, because of their sins and wickedness. They were the most wicked people on earth, because of their secret murders and plans. They were the ones who destroyed the peace of my people and the government. I caused them to be burned and destroyed, so that the blood of the prophets and the saints will stop crying out to me for revenge.
%3 Nephi 9:10
\bverse \iffalse And behold, the city of Laman, and the city of Josh, and the city of Gad, and the city of Kishkumen, have I caused to be burned with fire, and the inhabitants thereof, because of their wickedness in casting out the prophets, and stoning those whom I did send to declare unto them concerning their wickedness and their abominations. \fi
And I have burned the cities of Laman, and Josh, and Gad, and Kishkumen, and everyone that lived in them, because they did evil when they threw out the prophets, and stoned the people that I sent to them to tell them that they were living in horrible wickedness.
%3 Nephi 9:11
\bverse \iffalse And because they did cast them all out, that there were none righteous among them, I did send down fire and destroy them, that their wickedness and abominations might be hid from before my face, that the blood of the prophets and the saints whom I sent among them might not cry unto me from the ground against them. \fi
They threw them all out, so that there were no more righteous people living among them. And because of this I sent fire down to destroy them, so that their wickedness and evil would be hidden from my sight, so that the blood of the prophets and the saints would stop crying out to me for revenge.
%3 Nephi 9:12
\bverse \iffalse And many great destructions have I caused to come upon this land, and upon this people, because of their wickedness and their abominations. \fi
I have caused many parts of this land and this people to be completely destroyed, because they were wicked and evil.
%3 Nephi 9:13
\bverse \iffalse O all ye that are spared because ye were more righteous than they, will ye not now return unto me, and repent of your sins, and be converted, that I may heal you? \fi
All of you that are spared because you were more righteous than they were---won't you come back to me and repent of your sins, and be converted, so that I can heal you?
%3 Nephi 9:14
\bverse \iffalse Yea, verily I say unto you, if ye will come unto me ye shall have eternal life. Behold, mine arm of mercy is extended towards you, and whosoever will come, him will I receive; and blessed are those who come unto me. \fi
I tell you the truth. If you come to me you will have eternal life. My merciful arm is stretched out to you, and I will welcome anyone who comes to me. And the people who come to me are blessed.
%3 Nephi 9:15
\bverse \iffalse Behold, I am Jesus Christ the Son of God. I created the heavens and the earth, and all things that in them are. I was with the Father from the beginning. I am in the Father, and the Father in me; and in me hath the Father glorified his name. \fi
I am Jesus Christ, the Son of God. I created the heavens and the earth, and everything that is in them. I was with the Father from the beginning. I live in the Father, and the Father lives in me. And the Father has brought glory to his name through me.
%3 Nephi 9:16
\bverse \iffalse I came unto my own, and my own received me not. And the scriptures concerning my coming are fulfilled. \fi
I came to my own people, and my own people did not welcome me. And the scriptures that tell about my coming have come true.
%3 Nephi 9:17
\bverse \iffalse And as many as have received me, to them have I given to become the sons of God; and even so will I to as many as shall believe on my name, for behold, by me redemption cometh, and in me is the law of Moses fulfilled. \fi
And I have made everyone who has welcomed me the children of God. I will make anyone who believes in me a child of God, because I am the one who saves them, and I am the one who makes the law of Moses complete and true.
%3 Nephi 9:18
\bverse \iffalse I am the light and the life of the world. I am Alpha and Omega, the beginning and the end. \fi
I am the light and the life of the world. I am the Beginning and the End.
%3 Nephi 9:19
\bverse \iffalse And ye shall offer up unto me no more the shedding of blood; yea, your sacrifices and your burnt offerings shall be done away, for I will accept none of your sacrifices and your burnt offerings. \fi
You shall no longer sacrifice the blood of animals to me. You will no longer do sacrifices and burnt offerings, because I will no longer accept them.
%3 Nephi 9:20
\bverse \iffalse And ye shall offer for a sacrifice unto me a broken heart and a contrite spirit. And whoso cometh unto me with a broken heart and a contrite spirit, him will I baptize with fire and with the Holy Ghost, even as the Lamanites, because of their faith in me at the time of their conversion, were baptized with fire and with the Holy Ghost, and they knew it not. \fi
Your sacrifice to me will be a heart that is broken and sorry for your sins. And I will baptize with fire and the Holy Ghost anyone who comes to me with a broken heart, who is sorry for their sins. This is what I did for the Lamanites, because they had faith in me when they were converted. They were baptized with fire and with the Holy Ghost, and they did not know it.
%3 Nephi 9:21
\bverse \iffalse Behold, I have come unto the world to bring redemption unto the world, to save the world from sin. \fi
I have come to the world to redeem the world, and save the world from sin.
%3 Nephi 9:22
\bverse \iffalse Therefore, whoso repenteth and cometh unto me as a little child, him will I receive, for of such is the kingdom of God.  Behold, for such I have laid down my life, and have taken it up again; therefore repent, and come unto me ye ends of the earth, and be saved. \fi
And I will welcome anyone who repents and comes to me as a little child; the kingdom of God belongs to people who are like little children. I have given my life and been resurrected for such people. So come to me, all people everywhere, and repent and be saved!''
%Chapter 10
\bchapter
%3 Nephi 10:1
\bverse \iffalse And now behold, it came to pass that all the people of the land did hear these sayings, and did witness of it. And after these sayings there was silence in the land for the space of many hours; \fi
All the people all over the land heard the voice that said these things. And after they heard this voice there was silence in the land for many hours.
%3 Nephi 10:2
\bverse \iffalse For so great was the astonishment of the people that they did cease lamenting and howling for the loss of their kindred which had been slain; therefore there was silence in all the land for the space of many hours. \fi
The people were so amazed that they stopped howling and crying for their family and friends who had been killed. So there was silence throughout the whole land for many hours.
%3 Nephi 10:3
\bverse \iffalse And it came to pass that there came a voice again unto the people, and all the people did hear, and did witness of it, saying: \fi
And a voice came to the people again. All of the people heard it, and gave their testimony that they heard it. It said:
%3 Nephi 10:4
\bverse \iffalse O ye people of these great cities which have fallen, who are descendants of Jacob, yea, who are of the house of Israel, how oft have I gathered you as a hen gathereth her chickens under her wings, and have nourished you. \fi
``O you people of these great cities that have been destroyed, you descendants of Jacob, who are Israelites---how often I have gathered you like a hen gathers her chickens under her wings, and have cared for you!''
%3 Nephi 10:5
\bverse \iffalse And again, how oft would I have gathered you as a hen gathereth her chickens under her wings, yea, O ye people of the house of Israel, who have fallen; yea, O ye people of the house of Israel, ye that dwell at Jerusalem, as ye that have fallen; yea, how oft would I have gathered you as a hen gathereth her chickens, and ye would not. \fi
``How often I have wanted to gather you as a hen gathers her chickens under her wings, you descendants of Israel, who have died! O you descendants of Israel, both you who live near Jerusalem and you who have been killed---how often I have wanted to gather you as a hen gathers her chickens, but you refused!
%3 Nephi 10:6
\bverse \iffalse O ye house of Israel whom I have spared, how oft will I gather you as a hen gathereth her chickens under her wings, if ye will repent and return unto me with full purpose of heart. \fi
O you descendants of Israel whom I have spared, how often I will gather you as a hen gathers her chickens under her wings, if you repent and come back to me with all of your heart!
%3 Nephi 10:7
\bverse \iffalse But if not, O house of Israel, the places of your dwellings shall become desolate until the time of the fulfilling of the covenant to your fathers. \fi
But if you do not, O family of Israel, the places where you live will be empty until the time that the promises of the covenant I made to your ancestors come true.''
%3 Nephi 10:8
\bverse \iffalse And now it came to pass that after the people had heard these words, behold, they began to weep and howl again because of the loss of their kindred and friends. \fi
After the people heard these words, they began to howl and cry again for their family and friends who had died.
%3 Nephi 10:9
\bverse \iffalse And it came to pass that thus did the three days pass away.  And it was in the morning, and the darkness dispersed from off the face of the land, and the earth did cease to tremble, and the rocks did cease to rend, and the dreadful groanings did cease, and all the tumultuous noises did pass away. \fi
This is how those three days went. And in the morning, the darkness left the land and the earth stopped shaking; the rocks stopped breaking into pieces, and all of the dreadful noises stopped.
%3 Nephi 10:10
\bverse \iffalse And the earth did cleave together again, that it stood; and the mourning, and the weeping, and the wailing of the people who were spared alive did cease; and their mourning was turned into joy, and their lamentations into the praise and thanksgiving unto the Lord Jesus Christ, their Redeemer. \fi
And the places where the earth had split came together again, and stopped rumbling. The mourning and crying of the people who were spared alive stopped. And their sadness was turned into joy. Their crying was turned into praise and thanks to the Lord Jesus Christ, their Redeemer.
%3 Nephi 10:11
\bverse \iffalse And thus far were the scriptures fulfilled which had been spoken by the prophets. \fi
And so far the scriptures that the prophets had spoken had come true.
%3 Nephi 10:12
\bverse \iffalse And it was the more righteous part of the people who were saved, and it was they who received the prophets and stoned them not; and it was they who had not shed the blood of the saints, who were spared-- \fi
And it was the more righteous part of the people who were saved. They were the ones who accepted the prophets and did not stone them. They were the ones who had not killed the saints.
%3 Nephi 10:13
\bverse \iffalse And they were spared and were not sunk and buried up in the earth; and they were not drowned in the depths of the sea; and they were not burned by fire, neither were they fallen upon and crushed to death; and they were not carried away in the whirlwind; neither were they overpowered by the vapor of smoke and of darkness. \fi
They were spared; they were not sunk and buried in the earth; they were not drowned in the sea; they were not burned up by fire; they were not crushed to death; they were not carried away in a whirlwind; they did not suffocate to death from the smoke and darkness.
%3 Nephi 10:14
\bverse \iffalse And now, whoso readeth, let him understand; he that hath the scriptures, let him search them, and see and behold if all these deaths and destructions by fire, and by smoke, and by tempests, and by whirlwinds, and by the opening of the earth to receive them, and all these things are not unto the fulfilling of the prophecies of many of the holy prophets. \fi
And anyone who is reading this, you should understand: if you have the scriptures, look through them. See if all these deaths and all this destruction by fire, smoke, windstorms, whirlwinds, earthquakes, and all of these things are not what the holy prophets said would happen.
%3 Nephi 10:15
\bverse \iffalse Behold, I say unto you, Yea, many have testified of these things at the coming of Christ, and were slain because they testified of these things. \fi
I tell you, yes, many have testified that these things would happen when Christ came, and they were killed because they testified about these things.
%3 Nephi 10:16
\bverse \iffalse Yea, the prophet Zenos did testify of these things, and also Zenock spake concerning these things, because they testified particularly concerning us, who are the remnant of their seed. \fi
The prophet Zenos testified about these things, and so did Zenock. They testified about us because we are their descendants.
%3 Nephi 10:17
\bverse \iffalse Behold, our father Jacob also testified concerning a remnant of the seed of Joseph. And behold, are not we a remnant of the seed of Joseph? And these things which testify of us, are they not written upon the plates of brass which our father Lehi brought out of Jerusalem? \fi
And our father Jacob also testified about the descendants of Asenath and Joseph. And we are the descendants of Asenath and Joseph. And these things that testify about us are written on the plates of brass that our mother Sariah and father Lehi brought out of Jerusalem.
%3 Nephi 10:18
\bverse \iffalse And it came to pass that in the ending of the thirty and fourth year, behold, I will show unto you that the people of Nephi who were spared, and also those who had been called Lamanites, who had been spared, did have great favors shown unto them, and great blessings poured out upon their heads, insomuch that soon after the ascension of Christ into heaven he did truly manifest himself unto them-- \fi
And at the end of the thirty-fourth year, I will show you that the people of Nephi who were spared, and the Lamanites who had been spared, were given very many great blessings. Their blessings were so great that soon after Christ had gone up into heaven, he showed himself to them.
%3 Nephi 10:19
\bverse \iffalse Showing his body unto them, and ministering unto them; and an account of his ministry shall be given hereafter. Therefore for this time I make an end of my sayings. \fi
He showed his body to them, and blessed and took care of them. And now I will stop writing my own words, and I will tell the story of the things he did.
%Chapter 11
\bchapter
%3 Nephi 11:1
\bverse \iffalse And now it came to pass that there were a great multitude gathered together, of the people of Nephi, round about the temple which was in the land Bountiful; and they were marveling and wondering one with another, and were showing one to another the great and marvelous change which had taken place. \fi
There was an enormous crowd of Nephites gathered around the temple in the land Bountiful. They were telling each other how surprised and amazed they were by how much things had changed around them, and showing these great changes to each other.
%3 Nephi 11:2
\bverse \iffalse And they were also conversing about this Jesus Christ, of whom the sign had been given concerning his death. \fi
They were also talking about Jesus Christ; the sign that he died had happened.
%3 Nephi 11:3
\bverse \iffalse And it came to pass that while they were thus conversing one with another, they heard a voice as if it came out of heaven; and they cast their eyes round about, for they understood not the voice which they heard; and it was not a harsh voice, neither was it a loud voice; nevertheless, and notwithstanding it being a small voice it did pierce them that did hear to the center, insomuch that there was no part of their frame that it did not cause to quake; yea, it did pierce them to the very soul, and did cause their hearts to burn. \fi
And while they were talking with each other, they heard a voice that sounded like it came out of heaven. They looked around, because they did not understand the voice that they heard. It was not a harsh voice, and it was not a loud voice. And even though it was a small voice, it spoke to the very center of their hearts and souls, and made their entire bodies shake and tremble. It went to the center of their souls, and made their hearts burn.
%3 Nephi 11:4
\bverse \iffalse And it came to pass that again they heard the voice, and they understood it not. \fi
And they heard the voice again, but they did not understand it.
%3 Nephi 11:5
\bverse \iffalse And again the third time they did hear the voice, and did open their ears to hear it; and their eyes were towards the sound thereof; and they did look steadfastly towards heaven, from whence the sound came. \fi
When they heard the voice a third time, they turned their ears to hear it, and they looked toward the sound of the voice. They looked up toward heaven, where the voice came from.
%3 Nephi 11:6
\bverse \iffalse And behold, the third time they did understand the voice which they heard; and it said unto them: \fi
And the third time, they understood the voice that they heard. It said to them:
%3 Nephi 11:7
\bverse \iffalse Behold my Beloved Son, in whom I am well pleased, in whom I have glorified my name--hear ye him. \fi
Look and see my Beloved Son, whom I am very pleased with. I have glorified my name through him. Listen to him.
%3 Nephi 11:8
\bverse \iffalse And it came to pass, as they understood they cast their eyes up again towards heaven; and behold, they saw a Man descending out of heaven; and he was clothed in a white robe; and he came down and stood in the midst of them; and the eyes of the whole multitude were turned upon him, and they durst not open their mouths, even one to another, and wist not what it meant, for they thought it was an angel that had appeared unto them. \fi
And when they understood they looked up toward heaven. They saw a Man coming out of heaven. He had on a white robe, and he came down and stood in the middle of the crowds. They were all looking at him. They did not dare to open their mouths---not even to talk to each other. They did not know what this meant; they thought it was an angel that had appeared to them.
%3 Nephi 11:9
\bverse \iffalse And it came to pass that he stretched forth his hand and spake unto the people, saying: \fi
And he stretched out his hand and spoke to the people. He said:
%3 Nephi 11:10
\bverse \iffalse Behold, I am Jesus Christ, whom the prophets testified shall come into the world. \fi
``I am Jesus Christ, the one whom the prophets testified would come into the world.
%3 Nephi 11:11
\bverse \iffalse And behold, I am the light and the life of the world; and I have drunk out of that bitter cup which the Father hath given me, and have glorified the Father in taking upon me the sins of the world, in the which I have suffered the will of the Father in all things from the beginning. \fi
I am the light of the world, and the life of the world. I have drunk out of the bitter cup\footnote{When Jesus says that he drank out of the bitter cup, he means that he has done the very hard things that the Father wanted him to do.} that the Father gave me. I have brought glory to the Father by taking all the sins of the world on myself; I have done everything that the Father wanted me to do since the beginning.''
%3 Nephi 11:12
\bverse \iffalse And it came to pass that when Jesus had spoken these words the whole multitude fell to the earth; for they remembered that it had been prophesied among them that Christ should show himself unto them after his ascension into heaven. \fi
When Jesus had spoken these words the entire crowd fell to the ground, because they remembered the prophecies that Christ would show himself to them after he went up into heaven.
%3 Nephi 11:13
\bverse \iffalse And it came to pass that the Lord spake unto them saying: \fi
The Lord continued to speak to them. He said:
%3 Nephi 11:14
\bverse \iffalse Arise and come forth unto me, that ye may thrust your hands into my side, and also that ye may feel the prints of the nails in my hands and in my feet, that ye may know that I am the God of Israel, and the God of the whole earth, and have been slain for the sins of the world. \fi
``Stand up, and come to me. Put your hands into my side, and feel the prints of the nails in my hands and feet, so that you can know that I am the God of Israel, and the God of the whole earth. I have been killed for the sins of the world.''
%3 Nephi 11:15
\bverse \iffalse And it came to pass that the multitude went forth, and thrust their hands into his side, and did feel the prints of the nails in his hands and in his feet; and this they did do, going forth one by one until they had all gone forth, and did see with their eyes and did feel with their hands, and did know of a surety and did bear record, that it was he, of whom it was written by the prophets, that should come. \fi
And the crowds came forward, and put their hands into his side, and felt the prints of the nails in his hands and feet. They came forward one by one, until all of them had come, and seen with their eyes and felt with their hands. They knew for themselves, and bore testimony, that he was the one that the prophets had written about, and said would come.
%3 Nephi 11:16
\bverse \iffalse And when they had all gone forth and had witnessed for themselves, they did cry out with one accord, saying: \fi
And after they had all come forward and seen for themselves, they cried out together and said:
%3 Nephi 11:17
\bverse \iffalse Hosanna! Blessed be the name of the Most High God! And they did fall down at the feet of Jesus, and did worship him. \fi
``Hosanna! Bless the name of the Most High God!'' And they fell down at Jesus's feet and worshipped him. 
%3 Nephi 11:18
\bverse \iffalse And it came to pass that he spake unto Nephi (for Nephi was among the multitude) and he commanded him that he should come forth. \fi
Nephi was also in the crowd. Jesus spoke to Nephi, and commanded him to come forward.
%3 Nephi 11:19
\bverse \iffalse And Nephi arose and went forth, and bowed himself before the Lord and did kiss his feet. \fi
Nephi stood up and came forward, and bowed down before the Lord and kissed his feet.
%3 Nephi 11:20
\bverse \iffalse And the Lord commanded him that he should arise. And he arose and stood before him. \fi
The Lord commanded him to stand up, so Nephi stood up, in front of Jesus.
%3 Nephi 11:21
\bverse \iffalse And the Lord said unto him: I give unto you power that ye shall baptize this people when I am again ascended into heaven. \fi
And the Lord said to him: ``I give you power to baptize this people when I go back up to heaven.''
%3 Nephi 11:22
\bverse \iffalse And again the Lord called others, and said unto them likewise; and he gave unto them power to baptize. And he said unto them: On this wise shall ye baptize; and there shall be no disputations among you. \fi
And the Lord called others, and said the same things to them. He gave them power to baptize. And he said to them: ``This is how you should baptize. Do not let there be any arguments about this among you.
%3 Nephi 11:23
\bverse \iffalse Verily I say unto you, that whoso repenteth of his sins through your words and desireth to be baptized in my name, on this wise shall ye baptize them--Behold, ye shall go down and stand in the water, and in my name shall ye baptize them. \fi
I tell you the truth. Everyone who repents of their sins because of your words, and wants to be baptized in my name, will be baptized like this: You will go down and stand in the water, and you will baptize them in my name.
%3 Nephi 11:24
\bverse \iffalse And now behold, these are the words which ye shall say, calling them by name, saying: \fi
And these are the words that you will say. You will call them by name, and say:
%3 Nephi 11:25
\bverse \iffalse Having authority given me of Jesus Christ, I baptize you in the name of the Father, and of the Son, and of the Holy Ghost.  Amen. \fi
Having authority given to me by Jesus Christ, I baptize you in the name of the Father, and of the Son, and of the Holy Ghost. Amen.
%3 Nephi 11:26
\bverse \iffalse And then shall ye immerse them in the water, and come forth again out of the water. \fi
And then you will bury them in the water, and come back out of the water.
%3 Nephi 11:27
\bverse \iffalse And after this manner shall ye baptize in my name; for behold, verily I say unto you, that the Father, and the Son, and the Holy Ghost are one; and I am in the Father, and the Father in me, and the Father and I are one. \fi
This is how you should baptize in my name, because I tell you the truth: The Father, and the Son, and the Holy Ghost, are one. I am in the Father, and the Father is in me, and the Father and I are one.
%3 Nephi 11:28
\bverse \iffalse And according as I have commanded you thus shall ye baptize.  And there shall be no disputations among you, as there have hitherto been; neither shall there be disputations among you concerning the points of my doctrine, as there have hitherto been. \fi
And you must baptize in the way that I have commanded you. There should not be any more arguments among you about this. There should not be any more of the arguments that you have had about what my teachings mean, either. 
%3 Nephi 11:29
\bverse \iffalse For verily, verily I say unto you, he that hath the spirit of contention is not of me, but is of the devil, who is the father of contention, and he stirreth up the hearts of men to contend with anger, one with another. \fi
I tell you the truth: Anyone who causes people to fight and argue does not belong to me. They belong to the devil, who is the father of all arguments and fights. He stirs up people's hearts, and gets them to fight and argue, and be angry with each other.
%3 Nephi 11:30
\bverse \iffalse Behold, this is not my doctrine, to stir up the hearts of men with anger, one against another; but this is my doctrine, that such things should be done away. \fi
I do not teach you to stir up people's hearts and make them angry at each other. But this is my teaching: you should stop being angry with each other.
%3 Nephi 11:31
\bverse \iffalse Behold, verily, verily, I say unto you, I will declare unto you my doctrine. \fi
I tell you the truth, and I will give you my teachings.
%3 Nephi 11:32
\bverse \iffalse And this is my doctrine, and it is the doctrine which the Father hath given unto me; and I bear record of the Father, and the Father beareth record of me, and the Holy Ghost beareth record of the Father and me; and I bear record that the Father commandeth all men, everywhere, to repent and believe in me. \fi
This is my teaching, and it is the teaching that the Father has given to me. And I bear testimony of the Father, and the Father bears testimony of me, and the Holy Ghost bears testimony of the Father and me. And I bear testimony that the Father commands all people everywhere to repent and believe in me.
%3 Nephi 11:33
\bverse \iffalse And whoso believeth in me, and is baptized, the same shall be saved; and they are they who shall inherit the kingdom of God. \fi
And anyone who believes in me, and is baptized, will be saved. These are the people who will receive the kingdom of God.
%3 Nephi 11:34
\bverse \iffalse And whoso believeth not in me, and is not baptized, shall be damned. \fi
And anyone who does not believe in me, and is not baptized, will be judged guilty.
%3 Nephi 11:35
\bverse \iffalse Verily, verily, I say unto you, that this is my doctrine, and I bear record of it from the Father; and whoso believeth in me believeth in the Father also; and unto him will the Father bear record of me, for he will visit him with fire and with the Holy Ghost. \fi
I tell you the truth. This is my teaching, and I come from the Father to bear testimony of it. And anyone who believes in me also believes in the Father. The Father will bear testimony of me to that person, for he will visit that person with fire and with the Holy Ghost.
%3 Nephi 11:36
\bverse \iffalse And thus will the Father bear record of me, and the Holy Ghost will bear record unto him of the Father and me; for the Father, and I, and the Holy Ghost are one. \fi
So the Father will bear testimony of me, and the Holy Ghost will bear testimony of the Father and me, because the Father, and I, and the Holy Ghost, are one.
%3 Nephi 11:37
\bverse \iffalse And again I say unto you, ye must repent, and become as a little child, and be baptized in my name, or ye can in nowise receive these things. \fi
And I say to you again that you must repent and become like a little child, and be baptized in my name, or you cannot receive these things. 
%3 Nephi 11:38
\bverse \iffalse And again I say unto you, ye must repent, and be baptized in my name, and become as a little child, or ye can in nowise inherit the kingdom of God. \fi
I say to you again that you must repent and be baptized in my name, and become like a little child, or you cannot inherit the kingdom of God.
%3 Nephi 11:39
\bverse \iffalse Verily, verily, I say unto you, that this is my doctrine, and whoso buildeth upon this buildeth upon my rock, and the gates of hell shall not prevail against them. \fi
I tell you the truth. This is my teaching, and the person who builds on this teaching, builds on my rock. And all the powers of hell will not be able to conquer them. 
%3 Nephi 11:40
\bverse \iffalse And whoso shall declare more or less than this, and establish it for my doctrine, the same cometh of evil, and is not built upon my rock; but he buildeth upon a sandy foundation, and the gates of hell stand open to receive such when the floods come and the winds beat upon them. \fi
And anyone who says anything different from this, either more or less, and claims that it is my teaching, is speaking evil. They are not built on my rock; they are building on sandy ground. The gates of hell are open, waiting to receive them, when the floods come and the winds beat down on them.
%3 Nephi 11:41
\bverse \iffalse Therefore, go forth unto this people, and declare the words which I have spoken, unto the ends of the earth. \fi
So go out among this people, and declare the words that I have said to the ends of the earth.''
%Chapter 12
\bchapter
%3 Nephi 12:1
\bverse \iffalse And it came to pass that when Jesus had spoken these words unto Nephi, and to those who had been called, (now the number of them who had been called, and received power and authority to baptize, was twelve) and behold, he stretched forth his hand unto the multitude, and cried unto them, saying: Blessed are ye if ye shall give heed unto the words of these twelve whom I have chosen from among you to minister unto you, and to be your servants; and unto them I have given power that they may baptize you with water; and after that ye are baptized with water, behold, I will baptize you with fire and with the Holy Ghost; therefore blessed are ye if ye shall believe in me and be baptized, after that ye have seen me and know that I am. \fi
There were twelve who were called by Jesus, and who were given power and authority to baptize. And when Jesus finished saying these things to Nephi and to the others who had been called, he stretched out his hand to the multitude and cried out to them, saying: ``I have chosen these twelve people to take care of you, and teach you, and be your servants. I have given them power to baptize you with water. Any of you who obeys the words that they speak will find happiness. And after you are baptized with water, I will baptize you with fire and with the Holy Ghost. Now that you have seen me and you know that I exist, if you will believe in me and be baptized, you will be blessed.
%3 Nephi 12:2
\bverse \iffalse And again, more blessed are they who shall believe in your words because that ye shall testify that ye have seen me, and that ye know that I am. Yea, blessed are they who shall believe in your words, and come down into the depths of humility and be baptized, for they shall be visited with fire and with the Holy Ghost, and shall receive a remission of their sins. \fi
And the people who believe you, when you testify that you have seen me and that you know that I am real, are even more blessed. The people who believe your words, and who are humble and choose to be baptized, are blessed. They will be visited with fire and the Holy Ghost, and their sins will be forgiven.
%3 Nephi 12:3
\bverse \iffalse Yea, blessed are the poor in spirit who come unto me, for theirs is the kingdom of heaven. \fi
People who understand that they are poor without God are blessed, and the kingdom of heaven is theirs.
%3 Nephi 12:4
\bverse \iffalse And again, blessed are all they that mourn, for they shall be comforted. \fi
People who are very sad are blessed, and they will be comforted.
%3 Nephi 12:5
\bverse \iffalse And blessed are the meek, for they shall inherit the earth. \fi
People who are humble are blessed; the earth will belong to them.
%3 Nephi 12:6
\bverse \iffalse And blessed are all they who do hunger and thirst after righteousness, for they shall be filled with the Holy Ghost. \fi
People who want more than anything else to do what is right are blessed; they will be filled with the Holy Ghost. 
%3 Nephi 12:7
\bverse \iffalse And blessed are the merciful, for they shall obtain mercy. \fi
People who give mercy are blessed; they will be given mercy.
%3 Nephi 12:8
\bverse \iffalse And blessed are all the pure in heart, for they shall see God. \fi
People who have pure hearts and minds are blessed; they will see God.
%3 Nephi 12:9
\bverse \iffalse And blessed are all the peacemakers, for they shall be called the children of God. \fi
People who work to make peace are blessed; they will be called God's children.
%3 Nephi 12:10
\bverse \iffalse And blessed are all they who are persecuted for my name's sake, for theirs is the kingdom of heaven. \fi
People who are bullied and mistreated for being my followers are blessed; the kingdom of heaven is theirs.
%3 Nephi 12:11
\bverse \iffalse And blessed are ye when men shall revile you and persecute, and shall say all manner of evil against you falsely, for my sake; \fi
And you are blessed when people hate you and hurt you, and say all sorts of evil things that are not true about you, because you believe in me.
%3 Nephi 12:12
\bverse \iffalse For ye shall have great joy and be exceedingly glad, for great shall be your reward in heaven; for so persecuted they the prophets who were before you. \fi
You will be full of joy and very glad, and you will have a reward waiting for you in heaven. The prophets who came before you were also treated badly and hated.
%3 Nephi 12:13
\bverse \iffalse Verily, verily, I say unto you, I give unto you to be the salt of the earth; but if the salt shall lose its savor wherewith shall the earth be salted? The salt shall be thenceforth good for nothing, but to be cast out and to be trodden under foot of men. \fi
I tell you the truth. I have made you like the salt of the earth. But if the salt loses its flavor, it isn't good for anything because it can't be made salty again. It is thrown out of the house, and trampled on by the people outside.
%3 Nephi 12:14
\bverse \iffalse Verily, verily, I say unto you, I give unto you to be the light of this people. A city that is set on a hill cannot be hid. \fi
I tell you the truth. I have made you like lights among this people. A city that is built on a hill cannot be hidden.
%3 Nephi 12:15
\bverse \iffalse Behold, do men light a candle and put it under a bushel? Nay, but on a candlestick, and it giveth light to all that are in the house; \fi
When a candle is lit, does the one who lit it put it under a basket? No! They put it on a candlestick, so that it can give light to everyone in the house.
%3 Nephi 12:16
\bverse \iffalse Therefore let your light so shine before this people, that they may see your good works and glorify your Father who is in heaven. \fi
So let your light shine among this people. Let them see the good things you do, so that they glorify your Father and Mother who are in heaven.
%3 Nephi 12:17
\bverse \iffalse Think not that I am come to destroy the law or the prophets.  I am not come to destroy but to fulfil; \fi
Do not think that I have come to destroy the law and the prophets. I have not come to destroy those things; I have come to do the things that they said.
%3 Nephi 12:18
\bverse \iffalse For verily I say unto you, one jot nor tittle hath not passed away from the law, but in me it hath all been fulfilled. \fi
I tell you the truth: Through me, every bit of the law and prophecy in the law of Moses has come true.
%3 Nephi 12:19
\bverse \iffalse And behold, I have given you the law and the commandments of my Father, that ye shall believe in me, and that ye shall repent of your sins, and come unto me with a broken heart and a contrite spirit. Behold, ye have the commandments before you, and the law is fulfilled. \fi
And I have given you my Father's commandments, so that you will believe in me and repent of your sins, and come to me with a heart that is broken and sorry for your sins. You have the commandments in front of you, and the law has come true.
%3 Nephi 12:20
\bverse \iffalse Therefore come unto me and be ye saved; for verily I say unto you, that except ye shall keep my commandments, which I have commanded you at this time, ye shall in no case enter into the kingdom of heaven. \fi
So come to me, and be saved. I tell you the truth: if you do not keep the commandments that I am giving you, it will be impossible for you to come into the kingdom of heaven.
%3 Nephi 12:21
\bverse \iffalse Ye have heard that it hath been said by them of old time, and it is also written before you, that thou shalt not kill, and whosoever shall kill shall be in danger of the judgment of God; \fi
You have heard the old saying, and you have seen in the scriptures, that you shall not murder. And anyone who murders is in danger of God's punishment.
%3 Nephi 12:22
\bverse \iffalse But I say unto you, that whosoever is angry with his brother shall be in danger of his judgment. And whosoever shall say to his brother, Raca, shall be in danger of the council; and whosoever shall say, Thou fool, shall be in danger of hell fire. \fi
But I tell you, that anyone who is angry with their sister or brother is in danger of God's punishment. And anyone who says bad things to their sisters and brothers will be judged by the council. And anyone who says to another, \lq You are a fool!\rq{} will be in danger of the fire of hell.
%3 Nephi 12:23
\bverse \iffalse Therefore, if ye shall come unto me, or shall desire to come unto me, and rememberest that thy brother hast aught against thee-- \fi
So if you come to me---or if you want to come to me---and you remember that you have done something to upset your friend,
%3 Nephi 12:24
\bverse \iffalse Go thy way unto thy brother, and first be reconciled to thy brother, and then come unto me with full purpose of heart, and I will receive you. \fi
First go to your friend, and make up for what you have done, and make peace with them. Then come to me with all of your heart, and I will welcome you.
%3 Nephi 12:25
\bverse \iffalse Agree with thine adversary quickly while thou art in the way with him, lest at any time he shall get thee, and thou shalt be cast into prison. \fi
When your enemy is taking you to court, make friends with them quickly. If you do not, they may turn you over to the judge and have you thrown into prison.
%3 Nephi 12:26
\bverse \iffalse Verily, verily, I say unto thee, thou shalt by no means come out thence until thou hast paid the uttermost senine. And while ye are in prison can ye pay even one senine? Verily, verily, I say unto you, Nay. \fi
And once you are in prison, you will not get out until you have paid the very last cent. 
%3 Nephi 12:27
\bverse \iffalse Behold, it is written by them of old time, that thou shalt not commit adultery; \fi
You have seen the saying in the scriptures, that you shall not commit adultery.
%3 Nephi 12:28
\bverse \iffalse But I say unto you, that whosoever looketh on a woman, to lust after her, hath committed adultery already in his heart. \fi
But I tell you that if a person looks at another person with lust, that person has already committed adultery in their heart.
%3 Nephi 12:29
\bverse \iffalse Behold, I give unto you a commandment, that ye suffer none of these things to enter into your heart; \fi
I command you: Do not let any of these things come into your heart.
%3 Nephi 12:30
\bverse \iffalse For it is better that ye should deny yourselves of these things, wherein ye will take up your cross, than that ye should be cast into hell. \fi
It is better for you to keep away from these things, and even die on a cross, than to be thrown into hell.
%3 Nephi 12:31
\bverse \iffalse It hath been written, that whosoever shall put away his wife, let him give her a writing of divorcement. \fi
In the scriptures it says that any man who wants to get a divorce must give his wife a divorce paper.
%3 Nephi 12:32
\bverse \iffalse Verily, verily, I say unto you, that whosoever shall put away his wife, saving for the cause of fornication, causeth her to commit adultery; and whoso shall marry her who is divorced committeth adultery. \fi
But I tell you, that anyone who divorces their husband or wife causes them to be guilty of adultery. The only reason for a person to get divorced is if their husband or wife has been unfaithful. And anyone who marries that divorced person is guilty of adultery.
%3 Nephi 12:33
\bverse \iffalse And again it is written, thou shalt not forswear thyself, but shalt perform unto the Lord thine oaths; \fi
In the scriptures it says that you must do what you swear you will do, and you must keep the promises that you make to the Lord.
%3 Nephi 12:34
\bverse \iffalse But verily, verily, I say unto you, swear not at all; neither by heaven, for it is God's throne; \fi
But truly, I tell you, do not swear at all. Do not swear in the name of heaven, because heaven is God's throne.
%3 Nephi 12:35
\bverse \iffalse Nor by the earth, for it is his footstool; \fi
Do not swear in the name of the earth, because the earth also belongs to God.
%3 Nephi 12:36
\bverse \iffalse Neither shalt thou swear by thy head, because thou canst not make one hair black or white; \fi
Do not swear by your own head, because you do not even have power to make one of your own hairs black or white.
%3 Nephi 12:37
\bverse \iffalse But let your communication be Yea, yea; Nay, nay; for whatsoever cometh of more than these is evil. \fi
Instead, say \lq yes\rq{} when you mean \lq yes,\rq{} and say \lq no\rq{} when you mean \lq no.\rq{} Anything more than this is evil.
%3 Nephi 12:38
\bverse \iffalse And behold, it is written, an eye for an eye, and a tooth for a tooth; \fi
In the scriptures it says, \lq an eye for an eye, and a tooth for a tooth.\rq
%3 Nephi 12:39
\bverse \iffalse But I say unto you, that ye shall not resist evil, but whosoever shall smite thee on thy right cheek, turn to him the other also; \fi
But I tell you, don't stand up to fight an evil person. If someone slaps you on the cheek, turn and let them slap the other cheek also.
%3 Nephi 12:40
\bverse \iffalse And if any man will sue thee at the law and take away thy coat, let him have thy cloak also; \fi
And if anyone tries to sue you and take your shirt, give them your coat also.
%3 Nephi 12:41
\bverse \iffalse And whosoever shall compel thee to go a mile, go with him twain. \fi
And if someone forces you to walk a mile with them, walk two miles with them.
%3 Nephi 12:42
\bverse \iffalse Give to him that asketh thee, and from him that would borrow of thee turn not away. \fi
If someone asks you for help, give them help. When someone wants to borrow something from you, do not tell them to go away.
%3 Nephi 12:43
\bverse \iffalse And behold it is written also, that thou shalt love thy neighbor and hate thine enemy; \fi
It says in the scriptures that you should love your neighbor and hate your enemy.
%3 Nephi 12:44
\bverse \iffalse But I say unto you, love your enemies, bless them that curse you, do good to them that hate you, and pray for them who despitefully use you and persecute you; \fi
But I tell you, love your enemies. Bless the person who curses you. Do good to the person that hates you. Pray for the person who uses you and treats you badly.
%3 Nephi 12:45
\bverse \iffalse That ye may be the children of your Father who is in heaven; for he maketh his sun to rise on the evil and on the good. \fi
Then you will be the children of your Father and Mother who are in heaven. They let the sun shine on both evil people and good people.
%3 Nephi 12:46
\bverse \iffalse Therefore those things which were of old time, which were under the law, in me are all fulfilled. \fi
The things that were said long ago, in the law of Moses, have all come true.
%3 Nephi 12:47
\bverse \iffalse Old things are done away, and all things have become new. \fi
Old things go away, and everything has become like new.
%3 Nephi 12:48
\bverse \iffalse Therefore I would that ye should be perfect even as I, or your Father who is in heaven is perfect. \fi
I want you to be perfect like I am. I want you to be perfect like your Father and Mother in heaven.
%Chapter 13
\bchapter
%3 Nephi 13:1
\bverse \iffalse Verily, verily, I say that I would that ye should do alms unto the poor; but take heed that ye do not your alms before men to be seen of them; otherwise ye have no reward of your Father who is in heaven. \fi
I want you to give to the poor. But watch yourself; do not give to the poor just so that other people can see you. If you do, you will not get any reward from your Father and Mother in heaven.\footnote{Many of the verses in this chapter are paraphrased or closely quoted from Matthew 6 in the Internation Children's Bible.}
%3 Nephi 13:2
\bverse \iffalse Therefore, when ye shall do your alms do not sound a trumpet before you, as will hypocrites do in the synagogues and in the streets, that they may have glory of men. Verily I say unto you, they have their reward. \fi
When you give to the poor, do not try to get the attention of other people. This is what hypocrites do in the synagogues and in the streets, so that other people will praise them. I tell you the truth, they already have their reward.
%3 Nephi 13:3
\bverse \iffalse But when thou doest alms let not thy left hand know what thy right hand doeth; \fi
But when you give to the poor, do not let your left hand know what your right hand is doing.
%3 Nephi 13:4
\bverse \iffalse That thine alms may be in secret; and thy Father who seeth in secret, himself shall reward thee openly. \fi
When you give, do it in secret. Your Father and Mother see what you do, and they will reward you.
%3 Nephi 13:5
\bverse \iffalse And when thou prayest thou shalt not do as the hypocrites, for they love to pray, standing in the synagogues and in the corners of the streets, that they may be seen of men. Verily I say unto you, they have their reward. \fi
When you pray, do not pray like the hypocrites. They love to pray while they are standing in the synagogues and on the street, so that other people will see them. I tell you the truth: they already have their reward.
%3 Nephi 13:6
\bverse \iffalse But thou, when thou prayest, enter into thy closet, and when thou hast shut thy door, pray to thy Father who is in secret; and thy Father, who seeth in secret, shall reward thee openly. \fi
But when you pray, go into your room. When you have shut your door, pray to your Father in heaven. He sees you in your room. And he will reward you.
%3 Nephi 13:7
\bverse \iffalse But when ye pray, use not vain repetitions, as the heathen, for they think that they shall be heard for their much speaking. \fi
When you pray, do not say the same thing every time you pray. This is what the people who do not really know God do. They think that their prayers will be answered because they say the same things over and over again.
%3 Nephi 13:8
\bverse \iffalse Be not ye therefore like unto them, for your Father knoweth what things ye have need of before ye ask him. \fi
But do not do what they do. Your Father and Mother know which things you need before you ask for them.
%3 Nephi 13:9
\bverse \iffalse After this manner therefore pray ye: Our Father who art in heaven, hallowed be thy name. \fi
Pray like this: Our Father who is in heaven, may your name always be kept holy.
%3 Nephi 13:10
\bverse \iffalse Thy will be done on earth as it is in heaven. \fi
May the things that you want done, be done here on earth, just as they are in heaven.
%3 Nephi 13:11
\bverse \iffalse And forgive us our debts, as we forgive our debtors. \fi
Forgive us for the things we have done wrong, just as we forgive the people who have wronged us.
%3 Nephi 13:12
\bverse \iffalse And lead us not into temptation, but deliver us from evil. \fi
Help us to avoid temptation, and to avoid evil.
%3 Nephi 13:13
\bverse \iffalse For thine is the kingdom, and the power, and the glory, forever. Amen. \fi
The kingdom of heaven, and all power, and all glory, is yours forever. Amen.
%3 Nephi 13:14
\bverse \iffalse For, if ye forgive men their trespasses your heavenly Father will also forgive you; \fi
If you forgive others for the bad things that they do, your heavenly Father will also forgive you.
%3 Nephi 13:15
\bverse \iffalse But if ye forgive not men their trespasses neither will your Father forgive your trespasses. \fi
But if you do not forgive others, your Father will not forgive you either.
%3 Nephi 13:16
\bverse \iffalse Moreover, when ye fast be not as the hypocrites, of a sad countenance, for they disfigure their faces that they may appear unto men to fast. Verily I say unto you, they have their reward. \fi
When you fast, do not be like the hypocrites and make sad faces. They frown and make themselves look sad so that others will notice that they are fasting. I tell you the truth: they already have their reward.
%3 Nephi 13:17
\bverse \iffalse But thou, when thou fastest, anoint thy head, and wash thy face; \fi
But when you fast, wash your hair and your face.
%3 Nephi 13:18
\bverse \iffalse That thou appear not unto men to fast, but unto thy Father, who is in secret; and thy Father, who seeth in secret, shall reward thee openly. \fi
Do not try to get others to notice that you are fasting. But your Father and Mother see when you fast, and they will reward you.
%3 Nephi 13:19
\bverse \iffalse Lay not up for yourselves treasures upon earth, where moth and rust doth corrupt, and thieves break through and steal; \fi
Do not save treasures for yourselves here on earth. Moths and rust destroy those treasures, and thieves can break in and steal them.
%3 Nephi 13:20
\bverse \iffalse But lay up for yourselves treasures in heaven, where neither moth nor rust doth corrupt, and where thieves do not break through nor steal. \fi
Save treasures for yourselves in heaven. Moths and rust cannot destroy them, and thieves cannot break in and steal them.
%3 Nephi 13:21
\bverse \iffalse For where your treasure is, there will your heart be also. \fi
And the place where you put your treasures is where your heart will also be.
%3 Nephi 13:22
\bverse \iffalse The light of the body is the eye; if, therefore, thine eye be single, thy whole body shall be full of light. \fi
Your eye is the lamp of your body. And if your eyes only look for what is good and right, your whole body will be full of light.
%3 Nephi 13:23
\bverse \iffalse But if thine eye be evil, thy whole body shall be full of darkness. If, therefore, the light that is in thee be darkness, how great is that darkness! \fi
But if your eyes search for evil, your whole body will be full of darkness. And if the light in you is gone, how great is the darkness that is left behind!
%3 Nephi 13:24
\bverse \iffalse No man can serve two masters; for either he will hate the one and love the other, or else he will hold to the one and despise the other. Ye cannot serve God and Mammon. \fi
You cannot serve two masters. Either you will hate one master and love the other, or else you will obey one master and disobey the other. You cannot serve God when you only care about your money!''
%3 Nephi 13:25
\bverse \iffalse And now it came to pass that when Jesus had spoken these words he looked upon the twelve whom he had chosen, and said unto them: Remember the words which I have spoken. For behold, ye are they whom I have chosen to minister unto this people. Therefore I say unto you, take no thought for your life, what ye shall eat, or what ye shall drink; nor yet for your body, what ye shall put on. Is not the life more than meat, and the body than raiment? \fi
When Jesus had said these words he looked at the twelve that he had chosen, and he said to them: ``Remember what I have said. You are the ones that I have chosen to serve and teach this people. Do not worry about your lives. Do not worry about where you will get food or drink. Do not worry about where you will get clothes to wear. Isn't life more than just food and clothing?
%3 Nephi 13:26
\bverse \iffalse Behold the fowls of the air, for they sow not, neither do they reap nor gather into barns; yet your heavenly Father feedeth them. Are ye not much better than they? \fi
Look at the birds in the air. They do not plant seeds. They do not harvest and store food in barns either. But your heavenly Father feeds them. You know that you are worth much more than birds. 
%3 Nephi 13:27
\bverse \iffalse Which of you by taking thought can add one cubit unto his stature? \fi
You can't add any time to your life by worrying about it.\footnote{See Matthew 6:27 in the Internation Children's Bible.}
%3 Nephi 13:28
\bverse \iffalse And why take ye thought for raiment? Consider the lilies of the field how they grow; they toil not, neither do they spin; \fi
Why do you worry about clothes? Look at the lilies that grow in the field. They do not work or make cloth.
%3 Nephi 13:29
\bverse \iffalse And yet I say unto you, that even Solomon, in all his glory, was not arrayed like one of these. \fi
But I tell you that even Solomon, with all of his money, was not dressed as beautifully as one of these lilies.
%3 Nephi 13:30
\bverse \iffalse Wherefore, if God so clothe the grass of the field, which today is, and tomorrow is cast into the oven, even so will he clothe you, if ye are not of little faith. \fi
God clothes the grass and grains in the field like that. The grass and grains are alive today, and tomorrow they will be thrown into the oven. God does this for the grass and grains, and God also has power to clothe you and give you the things you need, if you have faith.
%3 Nephi 13:31
\bverse \iffalse Therefore take no thought, saying, What shall we eat? or, What shall we drink? or, Wherewithal shall we be clothed? \fi
So do not worry and say, \lq What are we doing to eat?\rq{} or, \lq What are we going to drink?\rq{} or, \lq Where are we going to get clothing?\rq
%3 Nephi 13:32
\bverse \iffalse For your heavenly Father knoweth that ye have need of all these things. \fi
Your heavenly Father knows that you need these things. 
%3 Nephi 13:33
\bverse \iffalse But seek ye first the kingdom of God and his righteousness, and all these things shall be added unto you. \fi
What you should want most is to build God's kingdom and to keep God's commandments. Then God will give you the other things that you need.
%3 Nephi 13:34
\bverse \iffalse Take therefore no thought for the morrow, for the morrow shall take thought for the things of itself. Sufficient is the day unto the evil thereof. \fi
So do not worry about tomorrow. Each day has enough of its own troubles. There will be other things to worry about tomorrow.''
%Chapter 14
\bchapter
%3 Nephi 14:1
\bverse \iffalse And now it came to pass that when Jesus had spoken these words he turned again to the multitude, and did open his mouth unto them again, saying: Verily, verily, I say unto you, Judge not, that ye be not judged. \fi
When Jesus finished talking to his twelve disciples he turned again to face the crowds. He opened his mouth and said to them: ``I say to you, do not judge other people. Then you will not be judged.
%3 Nephi 14:2
\bverse \iffalse For with what judgment ye judge, ye shall be judged; and with what measure ye mete, it shall be measured to you again. \fi
The way that you judge other people is the way that you will be judged. And the forgiveness and kindness you give to others will be given back to you.
%3 Nephi 14:3
\bverse \iffalse And why beholdest thou the mote that is in thy brother's eye, but considerest not the beam that is in thine own eye? \fi
Why do you notice a speck of dust in your friend's eye, but you do not notice the piece of wood in your own eye?
%3 Nephi 14:4
\bverse \iffalse Or how wilt thou say to thy brother: Let me pull the mote out of thine eye--and behold, a beam is in thine own eye? \fi
How are you going to say to your friend, \lq Let me get the speck of dust out of your eye,\rq{} when you have a big piece of wood in your own eye?
%3 Nephi 14:5
\bverse \iffalse Thou hypocrite, first cast the beam out of thine own eye; and then shalt thou see clearly to cast the mote out of thy brother's eye. \fi
You hypocrite. First get the piece of wood out of your own eye. Then you will be able to see clearly enough to get the speck of dust out of your friend's eye.
%3 Nephi 14:6
\bverse \iffalse Give not that which is holy unto the dogs, neither cast ye your pearls before swine, lest they trample them under their feet, and turn again and rend you. \fi
Do not give things that are holy to dogs. Do not throw your pearls out to pigs. They will just trample the pearls under their feet, and the dogs will come to attack you.
%3 Nephi 14:7
\bverse \iffalse Ask, and it shall be given unto you; seek, and ye shall find; knock, and it shall be opened unto you. \fi
Ask, and you will get what you ask for. Look, and you will find what you look for. Knock, and the door will open.
%3 Nephi 14:8
\bverse \iffalse For every one that asketh, receiveth; and he that seeketh, findeth; and to him that knocketh, it shall be opened. \fi
Everyone that keeps asking will receive something. Everyone that keeps looking will find something. A door will open for everyone who keeps knocking.
%3 Nephi 14:9
\bverse \iffalse Or what man is there of you, who, if his son ask bread, will give him a stone? \fi
When your child asks you for bread, do you give your child a rock?
%3 Nephi 14:10
\bverse \iffalse Or if he ask a fish, will he give him a serpent? \fi
Or if your children want fish, do you give them a snake?
%3 Nephi 14:11
\bverse \iffalse If ye then, being evil, know how to give good gifts unto your children, how much more shall your Father who is in heaven give good things to them that ask him? \fi
Even a person like you, who sins, gives good things to your children. So your Father and Mother in heaven will surely give good things to the person who asks.
%3 Nephi 14:12
\bverse \iffalse Therefore, all things whatsoever ye would that men should do to you, do ye even so to them, for this is the law and the prophets. \fi
Anything that you would want someone to do for you, you should also do for them. This is what the law and the prophets say.
%3 Nephi 14:13
\bverse \iffalse Enter ye in at the strait gate; for wide is the gate, and broad is the way, which leadeth to destruction, and many there be who go in thereat; \fi
Go in through the narrow gate. The other gate that is wide, and the road that it leads to, will lead you to hell. There are many people who go through the wide gate.
%3 Nephi 14:14
\bverse \iffalse Because strait is the gate, and narrow is the way, which leadeth unto life, and few there be that find it. \fi
But the gate that leads to true life is narrow, and the road that leads to true life is difficult. Not very many people find that road.
%3 Nephi 14:15
\bverse \iffalse Beware of false prophets, who come to you in sheep's clothing, but inwardly they are ravening wolves. \fi
Look out for false prophets. They come to you dressed like sheep, but inside they are hungry wolves.
%3 Nephi 14:16
\bverse \iffalse Ye shall know them by their fruits. Do men gather grapes of thorns, or figs of thistles? \fi
You will recognize them by the things they do. Good things don't come from bad people, just like grapes don't grow on thorny bushes, and figs don't grow on thorny weeds.
%3 Nephi 14:17
\bverse \iffalse Even so every good tree bringeth forth good fruit; but a corrupt tree bringeth forth evil fruit. \fi
Good trees produce good fruit. And bad trees produce bad fruit.
%3 Nephi 14:18
\bverse \iffalse A good tree cannot bring forth evil fruit, neither a corrupt tree bring forth good fruit. \fi
A good tree will not produce bad fruit; a bad tree will not produce good fruit.
%3 Nephi 14:19
\bverse \iffalse Every tree that bringeth not forth good fruit is hewn down, and cast into the fire. \fi
And every tree that does not produce good fruit is chopped down and thrown into the fire.
%3 Nephi 14:20
\bverse \iffalse Wherefore, by their fruits ye shall know them. \fi
So you will know what kind of person someone is by the things that they do.
%3 Nephi 14:21
\bverse \iffalse Not every one that saith unto me, Lord, Lord, shall enter into the kingdom of heaven; but he that doeth the will of my Father who is in heaven. \fi
Not everyone who says to me, \lq You are my Lord\rq{} can come into the kingdom of heaven. Only the ones who do the things that my Father in heaven wants them to.
%3 Nephi 14:22
\bverse \iffalse Many will say to me in that day: Lord, Lord, have we not prophesied in thy name, and in thy name have cast out devils, and in thy name done many wonderful works? \fi
Many people will say to me on the last day: \lq Lord, haven't we prophesied in your name, and forced out demons and done miracles in your name?\rq
%3 Nephi 14:23
\bverse \iffalse And then will I profess unto them: I never knew you; depart from me, ye that work iniquity. \fi
Then I will say to them: \lq Go away from me, you evil people. I never knew you.\rq
%3 Nephi 14:24
\bverse \iffalse Therefore, whoso heareth these sayings of mine and doeth them, I will liken him unto a wise man, who built his house upon a rock-- \fi
So the person who hears my teachings, and obeys them, is like a wise person who built their house on a rock.
%3 Nephi 14:25
\bverse \iffalse And the rain descended, and the floods came, and the winds blew, and beat upon that house; and it fell not, for it was founded upon a rock. \fi
And when the rain and floods came, and the winds blew against that house, it did not fall, because it was built on a rock.
%3 Nephi 14:26
\bverse \iffalse And every one that heareth these sayings of mine and doeth them not shall be likened unto a foolish man, who built his house upon the sand-- \fi
And the person who hears my teachings, but does not obey them, is like a foolish person who built their house in the sand.
%3 Nephi 14:27
\bverse \iffalse And the rain descended, and the floods came, and the winds blew, and beat upon that house; and it fell, and great was the fall of it. \fi
And when the rain and floods came, and the winds blew against that house, it fell down with a crash.''
%Chapter 15
\bchapter
%3 Nephi 15:1
\bverse \iffalse And now it came to pass that when Jesus had ended these sayings he cast his eyes round about on the multitude, and said unto them: Behold, ye have heard the things which I taught before I ascended to my Father; therefore, whoso remembereth these sayings of mine and doeth them, him will I raise up at the last day. \fi
When Jesus finished saying these things he looked around at the crowds, and said to them: ``Now you have heard the things that I taught before I went into heaven with my Father. On the last day, I will lift up the person who remembers my teachings and obeys them to be with me.''
%3 Nephi 15:2
\bverse \iffalse And it came to pass that when Jesus had said these words he perceived that there were some among them who marveled, and wondered what he would concerning the law of Moses; for they understood not the saying that old things had passed away, and that all things had become new. \fi
When Jesus said this he could tell that some of the people were amazed and confused. They wondered if Jesus wanted them to keep following the law of Moses, because they didn't understand when Jesus said that old things go away, and everything has become like new.
%3 Nephi 15:3
\bverse \iffalse And he said unto them: Marvel not that I said unto you that old things had passed away, and that all things had become new. \fi
He said to them: ``Do not be amazed that I told you that old things have gone away, and everything has become like new.
%3 Nephi 15:4
\bverse \iffalse Behold, I say unto you that the law is fulfilled that was given unto Moses. \fi
I told you that the law of Moses is now finished and complete.
%3 Nephi 15:5
\bverse \iffalse Behold, I am he that gave the law, and I am he who covenanted with my people Israel; therefore, the law in me is fulfilled, for I have come to fulfil the law; therefore it hath an end. \fi
I am the one who gave that law, and I am the one who covenanted with my people, the people of Israel. The law of Moses is complete because of me. I have come to do everything that is in the law of Moses, so the time for living that law has come to an end.
%3 Nephi 15:6
\bverse \iffalse Behold, I do not destroy the prophets, for as many as have not been fulfilled in me, verily I say unto you, shall all be fulfilled. \fi
Do not be confused. I am not here to destroy the things that the prophets have said. Anything that they prophesied that has not yet come true through me, will still come true.
%3 Nephi 15:7
\bverse \iffalse And because I said unto you that old things have passed away, I do not destroy that which hath been spoken concerning things which are to come. \fi
I told you that old things have gone away, but this does not mean that the things they prophesied will not happen. I have not come to destroy the things that they said.
%3 Nephi 15:8
\bverse \iffalse For behold, the covenant which I have made with my people is not all fulfilled; but the law which was given unto Moses hath an end in me. \fi
I am not finished keeping the promises that I made to my people, but the time for living the law that I gave to Moses is over, because I have come.
%3 Nephi 15:9
\bverse \iffalse Behold, I am the law, and the light. Look unto me, and endure to the end, and ye shall live; for unto him that endureth to the end will I give eternal life. \fi
I am the law, and I am the light. Look to me, and be faithful to the end. Then you will live forever. I will give eternal life to the person who lives faithfully to the end.
%3 Nephi 15:10
\bverse \iffalse Behold, I have given unto you the commandments; therefore keep my commandments. And this is the law and the prophets, for they truly testified of me. \fi
I have given you the commandments, so keep these commandments. This is the meaning of the law of Moses and the things that the prophets said. They testified about me.''
%3 Nephi 15:11
\bverse \iffalse And now it came to pass that when Jesus had spoken these words, he said unto those twelve whom he had chosen: \fi
When Jesus had finished say this, he said to the twelve disciples he had chosen:
%3 Nephi 15:12
\bverse \iffalse Ye are my disciples; and ye are a light unto this people, who are a remnant of the house of Joseph. \fi
``You are my disciples. You should be a light for this people. They are descendants of Asenath and Joseph.
%3 Nephi 15:13
\bverse \iffalse And behold, this is the land of your inheritance; and the Father hath given it unto you. \fi
This land has been given to you for your inheritance. The Father is the one who has given it to you.
%3 Nephi 15:14
\bverse \iffalse And not at any time hath the Father given me commandment that I should tell it unto your brethren at Jerusalem. \fi
When I was with your relatives in Jerusalem, the Father never commanded me to tell them about this.
%3 Nephi 15:15
\bverse \iffalse Neither at any time hath the Father given me commandment that I should tell unto them concerning the other tribes of the house of Israel, whom the Father hath led away out of the land. \fi
The Father never commanded me to tell them about the other tribes of Israel that the Father led out of the land.
%3 Nephi 15:16
\bverse \iffalse This much did the Father command me, that I should tell unto them: \fi
This is the only thing that the Father commanded me to tell them:
%3 Nephi 15:17
\bverse \iffalse That other sheep I have which are not of this fold; them also I must bring, and they shall hear my voice; and there shall be one fold, and one shepherd. \fi
\lq I have other sheep that are not a part of this flock. But I will also lead them, and they will hear my voice. At the end there will only be one flock, and one shepherd.\rq
%3 Nephi 15:18
\bverse \iffalse And now, because of stiffneckedness and unbelief they understood not my word; therefore I was commanded to say no more of the Father concerning this thing unto them. \fi
But because they were proud and did not have faith, they did not understand what I said, and the Father commanded me not to say any more to them about this.
%3 Nephi 15:19
\bverse \iffalse But, verily, I say unto you that the Father hath commanded me, and I tell it unto you, that ye were separated from among them because of their iniquity; therefore it is because of their iniquity that they know not of you. \fi
But the Father did command me to tell you that you were separated from them because of their sins. It is because of their sins that they do not know about you.
%3 Nephi 15:20
\bverse \iffalse And verily, I say unto you again that the other tribes hath the Father separated from them; and it is because of their iniquity that they know not of them. \fi
I tell you again, the Father has also separated the other tribes from them. Because of their sins, they do not know that the Father has done this.
%3 Nephi 15:21
\bverse \iffalse And verily I say unto you, that ye are they of whom I said: Other sheep I have which are not of this fold; them also I must bring, and they shall hear my voice; and there shall be one fold, and one shepherd. \fi
But I tell you the truth---you are the ones I was talking about when I said: \lq I have other sheep that are not a part of this flock. I will also lead them, and they will follow my voice. And at the end there will be only one flock, and one shepherd.\rq
%3 Nephi 15:22
\bverse \iffalse And they understood me not, for they supposed it had been the Gentiles; for they understood not that the Gentiles should be converted through their preaching. \fi
They did not understand me; they thought that I was talking about the nonbelievers. They did not understand that the nonbelievers would be converted by their preaching.
%3 Nephi 15:23
\bverse \iffalse And they understood me not that I said they shall hear my voice; and they understood me not that the Gentiles should not at any time hear my voice--that I should not manifest myself unto them save it were by the Holy Ghost. \fi
They did not understand me when I said that they will hear my voice. They did not understand me, because the nonbelievers would never hear my voice. I would only show myself to the nonbelievers through the Holy Ghost.
%3 Nephi 15:24
\bverse \iffalse But behold, ye have both heard my voice, and seen me; and ye are my sheep, and ye are numbered among those whom the Father hath given me. \fi
But now you have heard my own voice, and you have seen me. You are my sheep, and you are a part of the people that the Father has given to me.
%Chapter 16
\bchapter
%3 Nephi 16:1
\bverse \iffalse And verily, verily, I say unto you that I have other sheep which are not of this land, neither of the land of Jerusalem, neither in any parts of that land round about whither I have been to minister. \fi
I tell you the truth; I have other sheep that are not in this land, and that are not in the land of Jerusalem. They are not in any part of that land where I have been teaching and serving.
%3 Nephi 16:2
\bverse \iffalse For they of whom I speak are they who have not as yet heard my voice; neither have I at any time manifested myself unto them. \fi
The sheep that I am talking about now have not heard my voice yet. And I have never yet shown myself to them.
%3 Nephi 16:3
\bverse \iffalse But I have received a commandment of the Father that I shall go unto them, and that they shall hear my voice, and shall be numbered among my sheep, that there may be one fold and one shepherd; therefore I go to show myself unto them. \fi
But the Father has commanded me to go to them so that they can hear my voice, and be counted as my sheep, so that there will be one flock and one shepherd. So I will go to show myself to them.
%3 Nephi 16:4
\bverse \iffalse And I command you that ye shall write these sayings after I am gone, that if it so be that my people at Jerusalem, they who have seen me and been with me in my ministry, do not ask the Father in my name, that they may receive a knowledge of you by the Holy Ghost, and also of the other tribes whom they know not of, that these sayings which ye shall write shall be kept and shall be manifested unto the Gentiles, that through the fulness of the Gentiles, the remnant of their seed, who shall be scattered forth upon the face of the earth because of their unbelief, may be brought in, or may be brought to a knowledge of me, their Redeemer. \fi
I command you to write down what I have said after I am gone. That way, if my people in the land of Jerusalem---the ones who have seen me and been with me while I taught---do not pray to the Father in my name, and have the Holy Ghost teach them about you and the other tribes that they do not know about, the things I have taught you will be kept safe, and revealed to the non-Israelites. Then the descendants of my people in Jerusalem will be scattered over the whole earth because of their unbelief. Then, when the time comes that the non-Israelites believe the gospel, they will bring the descendants of my people into the flock and teach them about me, their Redeemer.
%3 Nephi 16:5
\bverse \iffalse And then will I gather them in from the four quarters of the earth; and then will I fulfill the covenant which the Father hath made unto all the people of the house of Israel. \fi
Then I will gather them in from every place on earth. This is what I will do to keep the promises that the Father has made to all the descendants of Israel.
%3 Nephi 16:6
\bverse \iffalse And blessed are the Gentiles, because of their belief in me, in and of the Holy Ghost, which witnesses unto them of me and of the Father. \fi
And the non-Israelites are blessed because they believe in me through the Holy Ghost. The Holy Ghost bears testimony to them about me and the Father.
%3 Nephi 16:7
\bverse \iffalse Behold, because of their belief in me, saith the Father, and because of the unbelief of you, O house of Israel, in the latter day shall the truth come unto the Gentiles, that the fulness of these things shall be made known unto them. \fi
The Father says that because they believe in me, and because you do not believe, O descendants of Israel, in the last days the truth will go to the non-Israelites, and they will learn about all of these things.
%3 Nephi 16:8
\bverse \iffalse But wo, saith the Father, unto the unbelieving of the Gentiles--for notwithstanding they have come forth upon the face of this land, and have scattered my people who are of the house of Israel; and my people who are of the house of Israel have been cast out from among them, and have been trodden under feet by them; \fi
But how terrible it will be for the non-Israelites who do not believe, says the Father. They will come to this land, and scatter my people who are descendants of Israel. They will throw out my people who are descendants of Israel, and trample them.
%3 Nephi 16:9
\bverse \iffalse And because of the mercies of the Father unto the Gentiles, and also the judgments of the Father upon my people who are of the house of Israel, verily, verily, I say unto you, that after all this, and I have caused my people who are of the house of Israel to be smitten, and to be afflicted, and to be slain, and to be cast out from among them, and to become hated by them, and to become a hiss and a byword among them-- \fi
The Father will show mercy to the non-Israelites, and punish my people who are descendants of Israel. I will let the non-Israelites beat my people who are descendants of Israel, and make them suffer and kill them, and thrown them out, and hate them, and mock them.
%3 Nephi 16:10
\bverse \iffalse And thus commandeth the Father that I should say unto you: At that day when the Gentiles shall sin against my gospel, and shall reject the fulness of my gospel, and shall be lifted up in the pride of their hearts above all nations, and above all the people of the whole earth, and shall be filled with all manner of lyings, and of deceits, and of mischiefs, and all manner of hypocrisy, and murders, and priestcrafts, and whoredoms, and of secret abominations; and if they shall do all those things, and shall reject the fulness of my gospel, behold, saith the Father, I will bring the fulness of my gospel from among them. \fi
But this is what the Father commands me to say to you: The day will come when the non-Israelites will sin against my gospel, and reject the full truth of the gospel. They will be prouder than every other nation, and every other people on earth. They will lie and cause mischief. They will be hypocrites and murderers. They will puts themselves and their own teachings above God and God's words, so that they can get power and money. They will commit sexual sins, and do horrible things in secret. And when they do these things and reject the truth of my gospel, I will take my true gospel away from them. This is what the Father says.
%3 Nephi 16:11
\bverse \iffalse And then will I remember my covenant which I have made unto my people, O house of Israel, and I will bring my gospel unto them. \fi
Then I will remember the promise that I have made to my people, O people of Israel, and I will bring my gospel to them.
%3 Nephi 16:12
\bverse \iffalse And I will show unto thee, O house of Israel, that the Gentiles shall not have power over you; but I will remember my covenant unto you, O house of Israel, and ye shall come unto the knowledge of the fulness of my gospel. \fi
I will show you, O descendants of Israel, that the non-Israelites will not have power over you. I will remember the promise that I made to you, O people of Israel, and you will come to know the full truth of my gospel. 
%3 Nephi 16:13
\bverse \iffalse But if the Gentiles will repent and return unto me, saith the Father, behold they shall be numbered among my people, O house of Israel. \fi
But if the non-Israelites repent and come back to me, says the Father, they will be counted as my people, O descendants of Israel.
%3 Nephi 16:14
\bverse \iffalse And I will not suffer my people, who are of the house of Israel, to go through among them, and tread them down, saith the Father. \fi
I will not let my people, who are descendants of Israel, go among them and trample them to the ground, says the Father.
%3 Nephi 16:15
\bverse \iffalse But if they will not turn unto me, and hearken unto my voice, I will suffer them, yea, I will suffer my people, O house of Israel, that they shall go through among them, and shall tread them down, and they shall be as salt that hath lost its savor, which is thenceforth good for nothing but to be cast out, and to be trodden under foot of my people, O house of Israel. \fi
But if they will not come back to me and listen to my voice, I will let my people, the descendants of Israel, go among them and trample them to the ground. They will be like salt that has lost its flavor, which is good for nothing but to be thrown out and trampled under the feet of my people, O descendants of Israel.
%3 Nephi 16:16
\bverse \iffalse Verily, verily, I say unto you, thus hath the Father commanded me--that I should give unto this people this land for their inheritance. \fi
I speak the truth, and I tell you that the Father has commanded me to give this people this land for their inheritance.
%3 Nephi 16:17
\bverse \iffalse And then the words of the prophet Isaiah shall be fulfilled, which say: \fi
Then the words of the prophet Isaiah will come true. These are the words:
%3 Nephi 16:18
\bverse \iffalse Thy watchmen shall lift up the voice; with the voice together shall they sing, for they shall see eye to eye when the Lord shall bring again Zion. \fi
\lq Your guards will raise their voices. With one voice they will sing together. They will see eye to eye when the Lord brings back Zion.
%3 Nephi 16:19
\bverse \iffalse Break forth into joy, sing together, ye waste places of Jerusalem; for the Lord hath comforted his people, he hath redeemed Jerusalem. \fi
Shout for joy, and sing together, O ruins of Jerusalem! Because the Lord has comforted his people and redeemed Jerusalem.
%3 Nephi 16:20
\bverse \iffalse The Lord hath made bare his holy arm in the eye of all the nations; and all the ends of the earth shall see the salvation of God. \fi
The Lord has shown his holy arm in the eyes of all the nations, and people in the farthest corners of the earth will see the salvation of our God.\rq ''
%Chapter 17
\bchapter
%3 Nephi 17:1
\bverse \iffalse Behold, now it came to pass that when Jesus had spoken these words he looked round about again on the multitude, and he said unto them: Behold, my time is at hand. \fi
After Jesus said these things he looked around at the crowds again, and he said to them: ``It is time for me to go.
%3 Nephi 17:2
\bverse \iffalse I perceive that ye are weak, that ye cannot understand all my words which I am commanded of the Father to speak unto you at this time. \fi
I see that you can't understand all the words that the Father has commanded me to say to you at this time.
%3 Nephi 17:3
\bverse \iffalse Therefore, go ye unto your homes, and ponder upon the things which I have said, and ask of the Father, in my name, that ye may understand, and prepare your minds for the morrow, and I come unto you again. \fi
So go to your homes, and study and think about the things that I have said. Then ask the Father in my name to help you understand, so that you are ready to learn tomorrow when I come to you again.
%3 Nephi 17:4
\bverse \iffalse But now I go unto the Father, and also to show myself unto the lost tribes of Israel, for they are not lost unto the Father, for he knoweth whither he hath taken them. \fi
But now I go to the Father. I am also going to show myself to the lost tribes of Israel, because the Father knows where they are. He knows where he has taken them.''
%3 Nephi 17:5
\bverse \iffalse And it came to pass that when Jesus had thus spoken, he cast his eyes round about again on the multitude, and beheld they were in tears, and did look steadfastly upon him as if they would ask him to tarry a little longer with them. \fi
When Jesus finished talking he looked around at the crowds, and saw that there were tears in their eyes. They were looking at him as if they wanted him to stay with them a little longer.
%3 Nephi 17:6
\bverse \iffalse And he said unto them: Behold, my bowels are filled with compassion towards you. \fi
He said to them: ``My whole body is filled with love for you.
%3 Nephi 17:7
\bverse \iffalse Have ye any that are sick among you? Bring them hither. Have ye any that are lame, or blind, or halt, or maimed, or leprous, or that are withered, or that are deaf, or that are afflicted in any manner? Bring them hither and I will heal them, for I have compassion upon you; my bowels are filled with mercy. \fi
Is anyone sick among you? Bring them to me. Is anyone crippled, or blind, or deaf, or injured, or very sick, or troubled in any way? Bring them to me and I will heal them. I want to help you. My whole body is filled with mercy for you.
%3 Nephi 17:8
\bverse \iffalse For I perceive that ye desire that I should show unto you what I have done unto your brethren at Jerusalem, for I see that your faith is sufficient that I should heal you. \fi
I can see that you want me to show you what I did for your sisters and brothers in Jerusalem. I see that you have enough faith for me to heal you.''
%3 Nephi 17:9
\bverse \iffalse And it came to pass that when he had thus spoken, all the multitude, with one accord, did go forth with their sick and their afflicted, and their lame, and with their blind, and with their dumb, and with all them that were afflicted in any manner; and he did heal them every one as they were brought forth unto him. \fi
And after he finished talking, the whole crowd went to him, and brought those who were sick, or crippled, or blind, or could not speak, or who were troubled or injured in any way. And he healed every single person who was brought to him.
%3 Nephi 17:10
\bverse \iffalse And they did all, both they who had been healed and they who were whole, bow down at his feet, and did worship him; and as many as could come for the multitude did kiss his feet, insomuch that they did bathe his feet with their tears. \fi
And all of them---both the people who were healthy and the ones who had been healed---bowed down at his feet and worshipped him. It was so crowded that there was almost no room, but everyone who could came and kissed his feet. They washed his feet with their tears.
%3 Nephi 17:11
\bverse \iffalse And it came to pass that he commanded that their little children should be brought. \fi
And he commanded the crowds to bring their little children to him.
%3 Nephi 17:12
\bverse \iffalse So they brought their little children and set them down upon the ground round about him, and Jesus stood in the midst; and the multitude gave way till they had all been brought unto him. \fi
So they brought their little children to him and set them on the ground around him. Jesus was standing in the middle of the children. And the crowds made room so that all the children could be brought to Jesus.
%3 Nephi 17:13
\bverse \iffalse And it came to pass that when they had all been brought, and Jesus stood in the midst, he commanded the multitude that they should kneel down upon the ground. \fi
And after they had all been brought, Jesus stood in the middle of them and commanded the crowds to kneel down on the ground.
%3 Nephi 17:14
\bverse \iffalse And it came to pass that when they had knelt upon the ground, Jesus groaned within himself, and said: Father, I am troubled because of the wickedness of the people of the house of Israel. \fi
And after they knelt down on the ground, Jesus sighed and said: ``Father, I am sad because of the wickedness of the descendants of Israel.''
%3 Nephi 17:15
\bverse \iffalse And when he had said these words, he himself also knelt upon the earth; and behold he prayed unto the Father, and the things which he prayed cannot be written, and the multitude did bear record who heard him. \fi
After he said this, he also knelt down on the ground. He prayed to the Father, and the things that he said cannot be written. But the crowds heard what he said.
%3 Nephi 17:16
\bverse \iffalse And after this manner do they bear record: The eye hath never seen, neither hath the ear heard, before, so great and marvelous things as we saw and heard Jesus speak unto the Father; \fi
And this is the testimony that they gave of what he said: ``No eye has ever seen, and no ear has ever heard, such great and wonderful things as we saw and heard Jesus say to the Father.
%3 Nephi 17:17
\bverse \iffalse And no tongue can speak, neither can there be written by any man, neither can the hearts of men conceive so great and marvelous things as we both saw and heard Jesus speak; and no one can conceive of the joy which filled our souls at the time we heard him pray for us unto the Father. \fi
No one could speak or write down the great and wonderful things that we heard Jesus say. It is impossible for anyone to imagine such wonderful things. And no one can imagine the joy that filled our spirits when we heard him pray to the Father for us.''
%3 Nephi 17:18
\bverse \iffalse And it came to pass that when Jesus had made an end of praying unto the Father, he arose; but so great was the joy of the multitude that they were overcome. \fi
And when Jesus had finished praying to the Father, he stood up. But the people in the crowds were so filled with joy that they could not stand.
%3 Nephi 17:19
\bverse \iffalse And it came to pass that Jesus spake unto them, and bade them arise. \fi
And Jesus spoke to them, and asked them to stand.
%3 Nephi 17:20
\bverse \iffalse And they arose from the earth, and he said unto them: Blessed are ye because of your faith. And now behold, my joy is full. \fi
So they stood up on their feet, and he said to them: ``You are blessed because of your faith. And now my joy is full.''
%3 Nephi 17:21
\bverse \iffalse And when he had said these words, he wept, and the multitude bare record of it, and he took their little children, one by one, and blessed them, and prayed unto the Father for them. \fi
And when he said these words he cried. The crowds witnessed it for themselves. And he took each of their little children, one by one, and blesssed them. He prayed to the Father for them.
%3 Nephi 17:22
\bverse \iffalse And when he had done this he wept again; \fi
When he finished, he cried again.
%3 Nephi 17:23
\bverse \iffalse And he spake unto the multitude, and said unto them: Behold your little ones. \fi
He said to the crowds: ``Look at your little ones.''
%3 Nephi 17:24
\bverse \iffalse And as they looked to behold they cast their eyes towards heaven, and they saw the heavens open, and they saw angels descending out of heaven as it were in the midst of fire; and they came down and encircled those little ones about, and they were encircled about with fire; and the angels did minister unto them. \fi
When they looked to see, they looked up toward heaven. They saw the heavens open, and angels coming out of heaven. It looked like they were surrounded by fire. They came down and surrounded the children, and it looked like the children were surrounded by fire. And the angels spoke with them and taught them.
%3 Nephi 17:25
\bverse \iffalse And the multitude did see and hear and bear record; and they know that their record is true for they all of them did see and hear, every man for himself; and they were in number about two thousand and five hundred souls; and they did consist of men, women, and children. \fi
The whole crowd saw this, and gave their testimony that it happened. They know that their testimony is true, because all of them saw and heard this. Every person saw for themselves. And there were about two thousand and five hundred people who saw this. They were children, women, and men.
%Chapter 18
\bchapter
%3 Nephi 18:1
\bverse \iffalse And it came to pass that Jesus commanded his disciples that they should bring forth some bread and wine unto him. \fi
Then Jesus commanded his disciples to bring some bread and wine to him.
%3 Nephi 18:2
\bverse \iffalse And while they were gone for bread and wine, he commanded the multitude that they should sit themselves down upon the earth. \fi
And while they went to get bread and wine, he commanded the crowds to sit on the ground.
%3 Nephi 18:3
\bverse \iffalse And when the disciples had come with bread and wine, he took of the bread and brake and blessed it; and he gave unto the disciples and commanded that they should eat. \fi
When the disciples came back with bread and wine, he took the bread and broke it, and blessed it. He gave some to the disciples and commanded them to eat.
%3 Nephi 18:4
\bverse \iffalse And when they had eaten and were filled, he commanded that they should give unto the multitude. \fi
When they had eaten and were filled, he commanded them to give some to the crowds.
%3 Nephi 18:5
\bverse \iffalse And when the multitude had eaten and were filled, he said unto the disciples: Behold there shall one be ordained among you, and to him will I give power that he shall break bread and bless it and give it unto the people of my church, unto all those who shall believe and be baptized in my name. \fi
And when everyone in the crowds had eaten and were filled, he said to the disciples: ``One of you will be ordained, and to that person I will give power to break bread and bless it, and give it to the people who belong to my church, who believe in me and are baptized in my name.
%3 Nephi 18:6
\bverse \iffalse And this shall ye always observe to do, even as I have done, even as I have broken bread and blessed it and given it unto you. \fi
I have broken bread and blessed it and given it to you. Make sure that you always do this, just as I have done it.
%3 Nephi 18:7
\bverse \iffalse And this shall ye do in remembrance of my body, which I have shown unto you. And it shall be a testimony unto the Father that ye do always remember me. And if ye do always remember me ye shall have my Spirit to be with you. \fi
You will do this to remember my body. I have shown my body to you. And this will be a testimony to the Father that you always remember me. And if you always remember me, you will always have my Spirit to be with you.''
%3 Nephi 18:8
\bverse \iffalse And it came to pass that when he said these words, he commanded his disciples that they should take of the wine of the cup and drink of it, and that they should also give unto the multitude that they might drink of it. \fi
And when he said these words, he commanded his disciples to take the cup full of wine and drink from it. Then he commanded them to give the cup to the multitude so that they could also drink from it.
%3 Nephi 18:9
\bverse \iffalse And it came to pass that they did so, and did drink of it and were filled; and they gave unto the multitude, and they did drink, and they were filled. \fi
So they did. All of them drank from the cup, and they were filled. Then they gave some of the wine to the crowds, and they drank, and they were also filled.
%3 Nephi 18:10
\bverse \iffalse And when the disciples had done this, Jesus said unto them: Blessed are ye for this thing which ye have done, for this is fulfilling my commandments, and this doth witness unto the Father that ye are willing to do that which I have commanded you. \fi
After the disciples did this, Jesus said to them: ``You are blessed for what you have done. When you do this you are keeping my commandments, and this witnesses to the Father that you are willing to do what I have commanded you to do.
%3 Nephi 18:11
\bverse \iffalse And this shall ye always do to those who repent and are baptized in my name; and ye shall do it in remembrance of my blood, which I have shed for you, that ye may witness unto the Father that ye do always remember me. And if ye do always remember me ye shall have my Spirit to be with you. \fi
You shall always do this for the people who repent and are baptized in my name. You will do it to remember my blood. I have poured out my blood for you. When you do this you show the Father that you always remember me. And if you always remember me, you will have my Spirit to be with you.
%3 Nephi 18:12
\bverse \iffalse And I give unto you a commandment that ye shall do these things. And if ye shall always do these things blessed are ye, for ye are built upon my rock. \fi
I command you to do these things. And if you always do these things you are blessed, and you build on my rock.
%3 Nephi 18:13
\bverse \iffalse But whoso among you shall do more or less than these are not built upon my rock, but are built upon a sandy foundation; and when the rain descends, and the floods come, and the winds blow, and beat upon them, they shall fall, and the gates of hell are ready open to receive them. \fi
But anyone among you who does more or less than these things does not build on my rock. They are building on a sandy foundation. And when it begins to rain, and the floods come, and the winds blow on them and hit them, they will fall down. And the gates of hell are open, ready to let them in.
%3 Nephi 18:14
\bverse \iffalse Therefore blessed are ye if ye shall keep my commandments, which the Father hath commanded me that I should give unto you. \fi
So you are blessed if you keep my commandments. These are the commandments that the Father has commanded me to give to you.
%3 Nephi 18:15
\bverse \iffalse Verily, verily, I say unto you, ye must watch and pray always, lest ye be tempted by the devil, and ye be led away captive by him. \fi
I tell you the truth, and I say to you, you must always be careful and pray for strength. If you do not the devil will tempt you, and take you and lead you away.
%3 Nephi 18:16
\bverse \iffalse And as I have prayed among you even so shall ye pray in my church, among my people who do repent and are baptized in my name. Behold I am the light; I have set an example for you. \fi
You must also pray in my church, among my people who repent and are baptized in my name, just as I have prayed here with you. Look at me: I am the light, and I have set an example for you.''
%3 Nephi 18:17
\bverse \iffalse And it came to pass that when Jesus had spoken these words unto his disciples, he turned again unto the multitude and said unto them: \fi
And when Jessu had spoken these words to his disciples, he turned toward the crowds and said to them:
%3 Nephi 18:18
\bverse \iffalse Behold, verily, verily, I say unto you, ye must watch and pray always lest ye enter into temptation; for Satan desireth to have you, that he may sift you as wheat. \fi
``I tell you the truth, and I say to you, you must always watch and be careful, and pray for strength against temptation. Satan wants to sift you and test you, like a farmer separating the wheat from the husks.
%3 Nephi 18:19
\bverse \iffalse Therefore ye must always pray unto the Father in my name; \fi
So you must always pray to the Father in my name.
%3 Nephi 18:20
\bverse \iffalse And whatsoever ye shall ask the Father in my name, which is right, believing that ye shall receive, behold it shall be given unto you. \fi
And when you ask the Father for something, it will be given to you if it is right, and if you believe that he will give it you.
%3 Nephi 18:21
\bverse \iffalse Pray in your families unto the Father, always in my name, that your wives and your children may be blessed. \fi
Pray with your families to Father. Always pray in my name. Then your families, your children and your husband or wife, will be blessed.
%3 Nephi 18:22
\bverse \iffalse And behold, ye shall meet together oft; and ye shall not forbid any man from coming unto you when ye shall meet together, but suffer them that they may come unto you and forbid them not; \fi
You should meet together often. You should not tell anyone that they are not allowed to meet with you. You should welcome them when they come.
%3 Nephi 18:23
\bverse \iffalse But ye shall pray for them, and shall not cast them out; and if it so be that they come unto you oft ye shall pray for them unto the Father, in my name. \fi
Do not throw them out. Instead, you should pray for them. And if they come to meet with you often, you should pray in my name to the Father for them.
%3 Nephi 18:24
\bverse \iffalse Therefore, hold up your light that it may shine unto the world. Behold I am the light which ye shall hold up--that which ye have seen me do. Behold ye see that I have prayed unto the Father, and ye all have witnessed. \fi
Hold up your light so that it can shine for the world. I am the light that you should hold up. Show the things that you have seen me do. All of you saw for yourselves how I prayed to the Father.
%3 Nephi 18:25
\bverse \iffalse And ye see that I have commanded that none of you should go away, but rather have commanded that ye should come unto me, that ye might feel and see; even so shall ye do unto the world; and whosoever breaketh this commandment suffereth himself to be led into temptation. \fi
You saw how I did not command any of you to go away. Instead, I commanded you to come to me, so that you could come and feel, and see for yourselves. This is what you should do for the world. And anyone who breaks this commandment is letting themselves be tempted by the devil.''
%3 Nephi 18:26
\bverse \iffalse And now it came to pass that when Jesus had spoken these words, he turned his eyes again upon the disciples whom he had chosen, and said unto them: \fi
And now, when Jesus had spoken these words, he turned toward the disciples that he had chosen again, and he said to them:
%3 Nephi 18:27
\bverse \iffalse Behold verily, verily, I say unto you, I give unto you another commandment, and then I must go unto my Father that I may fulfil other commandments which he hath given me. \fi
``I tell you the truth, and now I give you another commandment. After this I must go to my Father so that I can do the other things that he has commanded me to do.
%3 Nephi 18:28
\bverse \iffalse And now behold, this is the commandment which I give unto you, that ye shall not suffer any one knowingly to partake of my flesh and blood unworthily, when ye shall minister it; \fi
Here is the commandment: If you know that someone is not worthy to have some of my body or my blood, you must not let them have any when you are serving it.
%3 Nephi 18:29
\bverse \iffalse For whoso eateth and drinketh my flesh and blood unworthily eateth and drinketh damnation to his soul; therefore if ye know that a man is unworthy to eat and drink of my flesh and blood ye shall forbid him. \fi
Anyone who eats my body or drinks my blood when they are not worthy is bringing punishment on themselves. So if you know that a person is not worthy to eat my body and drink my blood, you must forbid them.
%3 Nephi 18:30
\bverse \iffalse Nevertheless, ye shall not cast him out from among you, but ye shall minister unto him and shall pray for him unto the Father, in my name; and if it so be that he repenteth and is baptized in my name, then shall ye receive him, and shall minister unto him of my flesh and blood. \fi
But you should not throw that person out from among you. Instead, you should help them and pray to the Father for them, in my name. And if they repent and are baptized in my name, you should welcome them. Then you can give them my body and my blood.
%3 Nephi 18:31
\bverse \iffalse But if he repent not he shall not be numbered among my people, that he may not destroy my people, for behold I know my sheep, and they are numbered. \fi
But if they do not repent they will not be counted as my people. They should not be allowed to destroy my people. I know my sheep, and I know how many sheep I have.
%3 Nephi 18:32
\bverse \iffalse Nevertheless, ye shall not cast him out of your synagogues, or your places of worship, for unto such shall ye continue to minister; for ye know not but what they will return and repent, and come unto me with full purpose of heart, and I shall heal them; and ye shall be the means of bringing salvation unto them. \fi
But you should not throw that person out of your synagogues or church meetings. You should keep teaching them and helping them. You don't know---someday they might repent, and come back to me with all of their heart. Then I will heal them, and you will be the one that helped them to find me and be saved.
%3 Nephi 18:33
\bverse \iffalse Therefore, keep these sayings which I have commanded you that ye come not under condemnation; for wo unto him whom the Father condemneth. \fi
So keep these commandments that I have given you. If you do, you will not be punished. How terrible it will be for the person that the Father punishes!
%3 Nephi 18:34
\bverse \iffalse And I give you these commandments because of the disputations which have been among you. And blessed are ye if ye have no disputations among you. \fi
I give you these commandments because of the arguments that you have had among yourselves. You will be blessed if you do not allow arguments among yourselves.
%3 Nephi 18:35
\bverse \iffalse And now I go unto the Father, because it is expedient that I should go unto the Father for your sakes. \fi
Now I go to the Father. It is best for you if I go to the Father now.''
%3 Nephi 18:36
\bverse \iffalse And it came to pass that when Jesus had made an end of these sayings, he touched with his hand the disciples whom he had chosen, one by one, even until he had touched them all, and spake unto them as he touched them. \fi
And when Jesus finished saying all of this, he touched all of the disciples whom he had chosen with his hand, one by one, until he had touched them all. He spoke to each of them as he touched them.
%3 Nephi 18:37
\bverse \iffalse And the multitude heard not the words which he spake, therefore they did not bear record; but the disciples bare record that he gave them power to give the Holy Ghost. And I will show unto you hereafter that this record is true. \fi
The crowds did not hear the words that he said, so they could not bear testimony about it. But the disciples bore testimony that he gave them power to give the Holy Ghost. And I will show you later that their testimonies are true.
%3 Nephi 18:38
\bverse \iffalse And it came to pass that when Jesus had touched them all, there came a cloud and overshadowed the multitude that they could not see Jesus. \fi
And when Jesus had touched them all, a cloud came and covered the crowds so that they could not see Jesus.
%3 Nephi 18:39
\bverse \iffalse And while they were overshadowed he departed from them, and ascended into heaven. And the disciples saw and did bear record that he ascended again into heaven. \fi
And while they were covered, he left them and went up into heaven. The disciples saw this, and bore testimony that he had gone back into heaven.
%Chapter 19
\bchapter
%3 Nephi 19:1
\bverse \iffalse And now it came to pass that when Jesus had ascended into heaven, the multitude did disperse, and every man did take his wife and his children and did return to his own home. \fi
And when Jesus had gone back into heaven the crowds left, and every mother and father took their children and went back to their own homes.
%3 Nephi 19:2
\bverse \iffalse And it was noised abroad among the people immediately, before it was yet dark, that the multitude had seen Jesus, and that he had ministered unto them, and that he would also show himself on the morrow unto the multitude. \fi
The news began to spread immediately---even before it was dark---that Jesus had appeared to the crowds and blessed them and taught them, and that he would appear again the next day.
%3 Nephi 19:3
\bverse \iffalse Yea, and even all the night it was noised abroad concerning Jesus; and insomuch did they send forth unto the people that there were many, yea, an exceedingly great number, did labor exceedingly all that night, that they might be on the morrow in the place where Jesus should show himself unto the multitude. \fi
The news about Jesus spread the entire night. The news spread so quickly that there were very many people who worked and prepared the entire night so that they could be at the place where Jesus was going to appear the next day.
%3 Nephi 19:4
\bverse \iffalse And it came to pass that on the morrow, when the multitude was gathered together, behold, Nephi and his brother whom he had raised from the dead, whose name was Timothy, and also his son, whose name was Jonas, and also Mathoni, and Mathonihah, his brother, and Kumen, and Kumenonhi, and Jeremiah, and Shemnon, and Jonas, and Zedekiah, and Isaiah--now these were the names of the disciples whom Jesus had chosen--and it came to pass that they went forth and stood in the midst of the multitude. \fi
And the next day, when the crowds had gathered together, the twelve disciples that Jesus had chosen came forward and stood in the middle of the crowds. These are the names of the twelve disciples: Nephi; his brother that he raised from the dead, Timothy; Nephi's son, Jonas; Mathoni; Nephi's brother, Mathonihah; Kumen; Kumenonhi; Jeremiah; Shemnon; Jonas; Zedekiah; and Isaiah.
%3 Nephi 19:5
\bverse \iffalse And behold, the multitude was so great that they did cause that they should be separated into twelve bodies. \fi
There were so many people in the crowds that the disciples had the people separate into twelve different groups.
%3 Nephi 19:6
\bverse \iffalse And the twelve did teach the multitude; and behold, they did cause that the multitude should kneel down upon the face of the earth, and should pray unto the Father in the name of Jesus. \fi
And the twelve disciples taught the crowds. They had them kneel down on the ground, and pray to the Father in the name of Jesus.
%3 Nephi 19:7
\bverse \iffalse And the disciples did pray unto the Father also in the name of Jesus. And it came to pass that they arose and ministered unto the people. \fi
The disciples also prayed to the Father in the name of Jesus. Then they stood up and continued to teach the people.
%3 Nephi 19:8
\bverse \iffalse And when they had ministered those same words which Jesus had spoken--nothing varying from the words which Jesus had spoken--behold, they knelt again and prayed to the Father in the name of Jesus. \fi
They taught the same words that Jesus had taught, and said exactly the things that Jesus had said. After this, they knelt down again and prayed to the Father in the name of Jesus.
%3 Nephi 19:9
\bverse \iffalse And they did pray for that which they most desired; and they desired that the Holy Ghost should be given unto them. \fi
And they prayed for the thing that they wanted most of all: they wanted the Holy Ghost to be given to them.
%3 Nephi 19:10
\bverse \iffalse And when they had thus prayed they went down unto the water's edge, and the multitude followed them. \fi
And after they prayed they went down to the edge of the water, and the crowds followed them.
%3 Nephi 19:11
\bverse \iffalse And it came to pass that Nephi went down into the water and was baptized. \fi
Nephi went down into the water, and he was baptized.
%3 Nephi 19:12
\bverse \iffalse And he came up out of the water and began to baptize. And he baptized all those whom Jesus had chosen. \fi
Then he came up out of the water and began to baptize the others. He baptized all of the other disciples that Jesus had chosen. 
%3 Nephi 19:13
\bverse \iffalse And it came to pass when they were all baptized and had come up out of the water, the Holy Ghost did fall upon them, and they were filled with the Holy Ghost and with fire. \fi
And when they were all baptized and had come up out of the water, the Holy Ghost came to them. They were filled with the Holy Ghost and with fire.
%3 Nephi 19:14
\bverse \iffalse And behold, they were encircled about as if it were by fire; and it came down from heaven, and the multitude did witness it, and did bear record; and angels did come down out of heaven and did minister unto them. \fi
It looked as though they were surrounded by fire. It came down from heaven, and the crowds saw it and bore testimony that they saw it. And angels came down from heaven and spoke with them and taught them.
%3 Nephi 19:15
\bverse \iffalse And it came to pass that while the angels were ministering unto the disciples, behold, Jesus came and stood in the midst and ministered unto them. \fi
And while the angels were speaking with the disciples and teaching them, Jesus came and stood in the middle of them, and spoke with them, and taught them.
%3 Nephi 19:16
\bverse \iffalse And it came to pass that he spake unto the multitude, and commanded them that they should kneel down again upon the earth, and also that his disciples should kneel down upon the earth. \fi
Then he spoke to the crowds, and commanded them to kneel down on the ground. He also commanded his disciples to kneel down on the ground.
%3 Nephi 19:17
\bverse \iffalse And it came to pass that when they had all knelt down upon the earth, he commanded his disciples that they should pray. \fi
And when they had all knelt down on the ground, he commanded his disciples to pray.
%3 Nephi 19:18
\bverse \iffalse And behold, they began to pray; and they did pray unto Jesus, calling him their Lord and their God. \fi
So they began to pray. They prayed to Jesus, and called him their Lord and their God.
%3 Nephi 19:19
\bverse \iffalse And it came to pass that Jesus departed out of the midst of them, and went a little way off from them and bowed himself to the earth, and he said: \fi
Then Jesus went a little way off from the disciples, and kneeled down on the ground and said:
%3 Nephi 19:20
\bverse \iffalse Father, I thank thee that thou hast given the Holy Ghost unto these whom I have chosen; and it is because of their belief in me that I have chosen them out of the world. \fi
``Father, I thank you for giving the Holy Ghost to these disciples that I have chosen. I have chosen them because they believe in me.
%3 Nephi 19:21
\bverse \iffalse Father, I pray thee that thou wilt give the Holy Ghost unto all them that shall believe in their words. \fi
Father, I pray and ask you to give the Holy Ghost to everyone that believes what they teach.
%3 Nephi 19:22
\bverse \iffalse Father, thou hast given them the Holy Ghost because they believe in me; and thou seest that they believe in me because thou hearest them, and they pray unto me; and they pray unto me because I am with them. \fi
Father, you have given them the Holy Ghost because they believe in me. You see that they believe in me because you can hear them praying to me. They are praying to me because I am with them.
%3 Nephi 19:23
\bverse \iffalse And now Father, I pray unto thee for them, and also for all those who shall believe on their words, that they may believe in me, that I may be in them as thou, Father, art in me, that we may be one. \fi
And now, Father, I am praying to you for them, and for everyone who believes what they say. I want them to believe in me, and I want to be in them just as you, Father, are in me, so that we can be one.''
%3 Nephi 19:24
\bverse \iffalse And it came to pass that when Jesus had thus prayed unto the Father, he came unto his disciples, and behold, they did still continue, without ceasing, to pray unto him; and they did not multiply many words, for it was given unto them what they should pray, and they were filled with desire. \fi
And after Jesus had prayed to the Father, he came to his disciples, and they had not stopped praying to Jesus. They were not saying the same things over and over again, because the things that they should say were told to them by the Spirit, and they wanted with all of their hearts to pray.
%3 Nephi 19:25
\bverse \iffalse And it came to pass that Jesus blessed them as they did pray unto him; and his countenance did smile upon them, and the light of his countenance did shine upon them, and behold they were as white as the countenance and also the garments of Jesus; and behold the whiteness thereof did exceed all the whiteness, yea, even there could be nothing upon earth so white as the whiteness thereof. \fi
And Jesus blessed them while they were praying to him. He smiled as he looked at them, and the light of Jesus's face shone on them. They were as white as Jesus's face and clothing; they were whiter than anything on earth can be.
%3 Nephi 19:26
\bverse \iffalse And Jesus said unto them: Pray on; nevertheless they did not cease to pray. \fi
And Jesus said unto them: Pray on; nevertheless they did not cease to pray.
%3 Nephi 19:27
\bverse \iffalse And he turned from them again, and went a little way off and bowed himself to the earth; and he prayed again unto the Father, saying: \fi
And he turned from them again, and went a little way off and bowed himself to the earth; and he prayed again unto the Father, saying:
%3 Nephi 19:28
\bverse \iffalse Father, I thank thee that thou hast purified those whom I have chosen, because of their faith, and I pray for them, and also for them who shall believe on their words, that they may be purified in me, through faith on their words, even as they are purified in me. \fi
``Father, I thank you for purifying the ones that I have chosen, because of their faith. I pray for them. I pray that the people who believe in what they say can also become pure through me, by having faith in their words, just as my disciples have been made pure through me.
%3 Nephi 19:29
\bverse \iffalse Father, I pray not for the world, but for those whom thou hast given me out of the world, because of their faith, that they may be purified in me, that I may be in them as thou, Father, art in me, that we may be one, that I may be glorified in them. \fi
Father, I am not praying for the people in the world. I am praying for the people that you have given me, because of their faith. I pray that you will make them pure and clean through me, so that I can be in them just as you, Father, are in me, so that we can be one, and my glory can be shown through them.''
%3 Nephi 19:30
\bverse \iffalse And when Jesus had spoken these words he came again unto his disciples; and behold they did pray steadfastly, without ceasing, unto him; and he did smile upon them again; and behold they were white, even as Jesus. \fi
And when Jesus had said these words he came back to his disciples. They were praying with all of their hearts to him, and they did not stop. He smiled down at them again; and they were as white as Jesus's face and clothing.
%3 Nephi 19:31
\bverse \iffalse And it came to pass that he went again a little way off and prayed unto the Father; \fi
And he went again a little way off and prayed to the Father.
%3 Nephi 19:32
\bverse \iffalse And tongue cannot speak the words which he prayed, neither can be written by man the words which he prayed. \fi
No human can speak or write the words that he prayed.
%3 Nephi 19:33
\bverse \iffalse And the multitude did hear and do bear record; and their hearts were open and they did understand in their hearts the words which he prayed. \fi
But the crowds bore testimony that they heard what he said. Their hearts were open, and they understood in their hearts the words that he prayed.
%3 Nephi 19:34
\bverse \iffalse Nevertheless, so great and marvelous were the words which he prayed that they cannot be written, neither can they be uttered by man. \fi
But the words that he prayed were so great and wonderful that they cannot be written down. They cannot be spoken or written by any human.
%3 Nephi 19:35
\bverse \iffalse And it came to pass that when Jesus had made an end of praying he came again to the disciples, and said unto them: So great faith have I never seen among all the Jews; wherefore I could not show unto them so great miracles, because of their unbelief. \fi
When Jesus finished praying he came back to the disciples and said to them: ``Among all the Jews I never saw as much faith as you and these people have. I could never show them such great miracles, because they did not believe.
%3 Nephi 19:36
\bverse \iffalse Verily I say unto you, there are none of them that have seen so great things as ye have seen; neither have they heard so great things as ye have heard. \fi
I tell you the truth. None of them have seen things as wonderful as the things you have seen. They have not heard anything so great as the things you have heard.''
%Chapter 20
\bchapter
%3 Nephi 20:1
\bverse \iffalse And it came to pass that he commanded the multitude that they should cease to pray, and also his disciples. And he commanded them that they should not cease to pray in their hearts. \fi
Jesus commanded the crowds and his disciples to stop praying out loud. He commanded them to keep praying in their hearts.
%3 Nephi 20:2
\bverse \iffalse And he commanded them that they should arise and stand up upon their feet. And they arose up and stood upon their feet. \fi
Then he commanded them to stand up. So they stood up.
%3 Nephi 20:3
\bverse \iffalse And it came to pass that he brake bread again and blessed it, and gave to the disciples to eat. \fi
Then be broke bread again and blessed it, and gave some to the disciples to eat.
%3 Nephi 20:4
\bverse \iffalse And when they had eaten he commanded them that they should break bread, and give unto the multitude. \fi
When they were finished eating he commanded them to break more bread and give it to the crowds.
%3 Nephi 20:5
\bverse \iffalse And when they had given unto the multitude he also gave them wine to drink, and commanded them that they should give unto the multitude. \fi
After they had given bread to the crowds he also gave them wine to drink, and commanded them to give some to the crowds.
%3 Nephi 20:6
\bverse \iffalse Now, there had been no bread, neither wine, brought by the disciples, neither by the multitude; \fi
Now, no bread or wine had been brought by the disciples or the crowds.
%3 Nephi 20:7
\bverse \iffalse But he truly gave unto them bread to eat, and also wine to drink. \fi
But he really did give them bread to eat and wine to drink.
%3 Nephi 20:8
\bverse \iffalse And he said unto them: He that eateth this bread eateth of my body to his soul; and he that drinketh of this wine drinketh of my blood to his soul; and his soul shall never hunger nor thirst, but shall be filled. \fi
And he said to them: ``The person that eats this bread is eating my body, and giving nourishment to their soul. And the person that drinks this wine is drinking my blood, and giving nourishment to their soul. And their soul will never be hungry or thirsty; it will be full.''
%3 Nephi 20:9
\bverse \iffalse Now, when the multitude had all eaten and drunk, behold, they were filled with the Spirit; and they did cry out with one voice, and gave glory to Jesus, whom they both saw and heard. \fi
When the crowds finished eating and drinking, they were filled with the Spirit. They cried out with one voice and gave glory to Jesus. They saw him and heard him for themselves.
%3 Nephi 20:10
\bverse \iffalse And it came to pass that when they had all given glory unto Jesus, he said unto them: Behold now I finish the commandment which the Father hath commanded me concerning this people, who are a remnant of the house of Israel. \fi
And when they had all given glory to Jesus, he said to them: ``Now I have done what the Father commanded me to do for you. You are all descendants of Israel.
%3 Nephi 20:11
\bverse \iffalse Ye remember that I spake unto you, and said that when the words of Isaiah should be fulfilled--behold they are written, ye have them before you, therefore search them-- \fi
You remember what I told you about the words of Isaiah. They are written down. You have them in front of you, so search them. 
%3 Nephi 20:12
\bverse \iffalse And verily, verily, I say unto you, that when they shall be fulfilled then is the fulfilling of the covenant which the Father hath made unto his people, O house of Israel. \fi
And I tell you the truth: the time when those words come true is the time when the Father will keep the covenant that he has made to his people, O descendants of Israel.
%3 Nephi 20:13
\bverse \iffalse And then shall the remnants, which shall be scattered abroad upon the face of the earth, be gathered in from the east and from the west, and from the south and from the north; and they shall be brought to the knowledge of the Lord their God, who hath redeemed them. \fi
Then all the people who are left of them, who will be scattered all over the earth, will be gathered in from the east and the west, and from the south and the north. Then they will learn about the Lord their God, the one who redeemed them.
%3 Nephi 20:14
\bverse \iffalse And the Father hath commanded me that I should give unto you this land, for your inheritance. \fi
And the Father has commanded me to give you this land for your inheritance.
%3 Nephi 20:15
\bverse \iffalse And I say unto you, that if the Gentiles do not repent after the blessing which they shall receive, after they have scattered my people-- \fi
And I tell you, that if the non-believers do not repent after the blessing that they will receive, and after they have scattered my people,
%3 Nephi 20:16
\bverse \iffalse Then shall ye, who are a remnant of the house of Jacob, go forth among them; and ye shall be in the midst of them who shall be many; and ye shall be among them as a lion among the beasts of the forest, and as a young lion among the flocks of sheep, who, if he goeth through both treadeth down and teareth in pieces, and none can deliver. \fi
Then 
%3 Nephi 20:17
\bverse \iffalse Thy hand shall be lifted up upon thine adversaries, and all thine enemies shall be cut off. \fi
Thy hand shall be lifted up upon thine adversaries, and all thine enemies shall be cut off.
%3 Nephi 20:18
\bverse \iffalse And I will gather my people together as a man gathereth his sheaves into the floor. \fi
And I will gather my people together as a man gathereth his sheaves into the floor.
%3 Nephi 20:19
\bverse \iffalse For I will make my people with whom the Father hath covenanted, yea, I will make thy horn iron, and I will make thy hoofs brass. And thou shalt beat in pieces many people; and I will consecrate their gain unto the Lord, and their substance unto the Lord of the whole earth. And behold, I am he who doeth it. \fi
For I will make my people with whom the Father hath covenanted, yea, I will make thy horn iron, and I will make thy hoofs brass. And thou shalt beat in pieces many people; and I will consecrate their gain unto the Lord, and their substance unto the Lord of the whole earth. And behold, I am he who doeth it.
%3 Nephi 20:20
\bverse \iffalse And it shall come to pass, saith the Father, that the sword of my justice shall hang over them at that day; and except they repent it shall fall upon them, saith the Father, yea, even upon all the nations of the Gentiles. \fi
And it shall come to pass, saith the Father, that the sword of my justice shall hang over them at that day; and except they repent it shall fall upon them, saith the Father, yea, even upon all the nations of the Gentiles.
%3 Nephi 20:21
\bverse \iffalse And it shall come to pass that I will establish my people, O house of Israel. \fi
And it shall come to pass that I will establish my people, O house of Israel.
%3 Nephi 20:22
\bverse \iffalse And behold, this people will I establish in this land, unto the fulfilling of the covenant which I made with your father Jacob; and it shall be a New Jerusalem. And the powers of heaven shall be in the midst of this people; yea, even I will be in the midst of you. \fi
And behold, this people will I establish in this land, unto the fulfilling of the covenant which I made with your father Jacob; and it shall be a New Jerusalem. And the powers of heaven shall be in the midst of this people; yea, even I will be in the midst of you.
%3 Nephi 20:23
\bverse \iffalse Behold, I am he of whom Moses spake, saying: A prophet shall the Lord your God raise up unto you of your brethren, like unto me; him shall ye hear in all things whatsoever he shall say unto you. And it shall come to pass that every soul who will not hear that prophet shall be cut off from among the people. \fi
Behold, I am he of whom Moses spake, saying: A prophet shall the Lord your God raise up unto you of your brethren, like unto me; him shall ye hear in all things whatsoever he shall say unto you. And it shall come to pass that every soul who will not hear that prophet shall be cut off from among the people.
%3 Nephi 20:24
\bverse \iffalse Verily I say unto you, yea, and all the prophets from Samuel and those that follow after, as many as have spoken, have testified of me. \fi
Verily I say unto you, yea, and all the prophets from Samuel and those that follow after, as many as have spoken, have testified of me.
%3 Nephi 20:25
\bverse \iffalse And behold, ye are the children of the prophets; and ye are of the house of Israel; and ye are of the covenant which the Father made with your fathers, saying unto Abraham: And in thy seed shall all the kindreds of the earth be blessed. \fi
And behold, ye are the children of the prophets; and ye are of the house of Israel; and ye are of the covenant which the Father made with your fathers, saying unto Abraham: And in thy seed shall all the kindreds of the earth be blessed.
%3 Nephi 20:26
\bverse \iffalse The Father having raised me up unto you first, and sent me to bless you in turning away every one of you from his iniquities; and this because ye are the children of the covenant-- \fi
The Father having raised me up unto you first, and sent me to bless you in turning away every one of you from his iniquities; and this because ye are the children of the covenant--
%3 Nephi 20:27
\bverse \iffalse And after that ye were blessed then fulfilleth the Father the covenant which he made with Abraham, saying: In thy seed shall all the kindreds of the earth be blessed--unto the pouring out of the Holy Ghost through me upon the Gentiles, which blessing upon the Gentiles shall make them mighty above all, unto the scattering of my people, O house of Israel. \fi
And after that ye were blessed then fulfilleth the Father the covenant which he made with Abraham, saying: In thy seed shall all the kindreds of the earth be blessed--unto the pouring out of the Holy Ghost through me upon the Gentiles, which blessing upon the Gentiles shall make them mighty above all, unto the scattering of my people, O house of Israel.
%3 Nephi 20:28
\bverse \iffalse And they shall be a scourge unto the people of this land.  Nevertheless, when they shall have received the fulness of my gospel, then if they shall harden their hearts against me I will return their iniquities upon their own heads, saith the Father. \fi
And they shall be a scourge unto the people of this land.  Nevertheless, when they shall have received the fulness of my gospel, then if they shall harden their hearts against me I will return their iniquities upon their own heads, saith the Father.
%3 Nephi 20:29
\bverse \iffalse And I will remember the covenant which I have made with my people; and I have covenanted with them that I would gather them together in mine own due time, that I would give unto them again the land of their fathers for their inheritance, which is the land of Jerusalem, which is the promised land unto them forever, saith the Father. \fi
And I will remember the covenant which I have made with my people; and I have covenanted with them that I would gather them together in mine own due time, that I would give unto them again the land of their fathers for their inheritance, which is the land of Jerusalem, which is the promised land unto them forever, saith the Father.
%3 Nephi 20:30
\bverse \iffalse And it shall come to pass that the time cometh, when the fulness of my gospel shall be preached unto them; \fi
And it shall come to pass that the time cometh, when the fulness of my gospel shall be preached unto them;
%3 Nephi 20:31
\bverse \iffalse And they shall believe in me, that I am Jesus Christ, the Son of God, and shall pray unto the Father in my name. \fi
And they shall believe in me, that I am Jesus Christ, the Son of God, and shall pray unto the Father in my name.
%3 Nephi 20:32
\bverse \iffalse Then shall their watchmen lift up their voice, and with the voice together shall they sing; for they shall see eye to eye. \fi
Then shall their watchmen lift up their voice, and with the voice together shall they sing; for they shall see eye to eye.
%3 Nephi 20:33
\bverse \iffalse Then will the Father gather them together again, and give unto them Jerusalem for the land of their inheritance. \fi
Then will the Father gather them together again, and give unto them Jerusalem for the land of their inheritance.
%3 Nephi 20:34
\bverse \iffalse Then shall they break forth into joy--Sing together, ye waste places of Jerusalem; for the Father hath comforted his people, he hath redeemed Jerusalem. \fi
Then shall they break forth into joy--Sing together, ye waste places of Jerusalem; for the Father hath comforted his people, he hath redeemed Jerusalem.
%3 Nephi 20:35
\bverse \iffalse The Father hath made bare his holy arm in the eyes of all the nations; and all the ends of the earth shall see the salvation of the Father; and the Father and I are one. \fi
The Father hath made bare his holy arm in the eyes of all the nations; and all the ends of the earth shall see the salvation of the Father; and the Father and I are one.
%3 Nephi 20:36
\bverse \iffalse And then shall be brought to pass that which is written: Awake, awake again, and put on thy strength, O Zion; put on thy beautiful garments, O Jerusalem, the holy city, for henceforth there shall no more come into thee the uncircumcised and the unclean. \fi
And then shall be brought to pass that which is written: Awake, awake again, and put on thy strength, O Zion; put on thy beautiful garments, O Jerusalem, the holy city, for henceforth there shall no more come into thee the uncircumcised and the unclean.
%3 Nephi 20:37
\bverse \iffalse Shake thyself from the dust; arise, sit down, O Jerusalem; loose thyself from the bands of thy neck, O captive daughter of Zion. \fi
Shake thyself from the dust; arise, sit down, O Jerusalem; loose thyself from the bands of thy neck, O captive daughter of Zion.
%3 Nephi 20:38
\bverse \iffalse For thus saith the Lord: Ye have sold yourselves for naught, and ye shall be redeemed without money. \fi
For thus saith the Lord: Ye have sold yourselves for naught, and ye shall be redeemed without money.
%3 Nephi 20:39
\bverse \iffalse Verily, verily, I say unto you, that my people shall know my name; yea, in that day they shall know that I am he that doth speak. \fi
Verily, verily, I say unto you, that my people shall know my name; yea, in that day they shall know that I am he that doth speak.
%3 Nephi 20:40
\bverse \iffalse And then shall they say: How beautiful upon the mountains are the feet of him that bringeth good tidings unto them, that publisheth peace; that bringeth good tidings unto them of good, that publisheth salvation; that saith unto Zion: Thy God reigneth! \fi
And then shall they say: How beautiful upon the mountains are the feet of him that bringeth good tidings unto them, that publisheth peace; that bringeth good tidings unto them of good, that publisheth salvation; that saith unto Zion: Thy God reigneth!
%3 Nephi 20:41
\bverse \iffalse And then shall a cry go forth: Depart ye, depart ye, go ye out from thence, touch not that which is unclean; go ye out of the midst of her; be ye clean that bear the vessels of the Lord. \fi
And then shall a cry go forth: Depart ye, depart ye, go ye out from thence, touch not that which is unclean; go ye out of the midst of her; be ye clean that bear the vessels of the Lord.
%3 Nephi 20:42
\bverse \iffalse For ye shall not go out with haste nor go by flight; for the Lord will go before you, and the God of Israel shall be your rearward. \fi
For ye shall not go out with haste nor go by flight; for the Lord will go before you, and the God of Israel shall be your rearward.
%3 Nephi 20:43
\bverse \iffalse Behold, my servant shall deal prudently; he shall be exalted and extolled and be very high. \fi
Behold, my servant shall deal prudently; he shall be exalted and extolled and be very high.
%3 Nephi 20:44
\bverse \iffalse As many were astonished at thee--his visage was so marred, more than any man, and his form more than the sons of men-- \fi
As many were astonished at thee--his visage was so marred, more than any man, and his form more than the sons of men--
%3 Nephi 20:45
\bverse \iffalse So shall he sprinkle many nations; the kings shall shut their mouths at him, for that which had not been told them shall they see; and that which they had not heard shall they consider. \fi
So shall he sprinkle many nations; the kings shall shut their mouths at him, for that which had not been told them shall they see; and that which they had not heard shall they consider.
%3 Nephi 20:46
\bverse \iffalse Verily, verily, I say unto you, all these things shall surely come, even as the Father hath commanded me. Then shall this covenant which the Father hath covenanted with his people be fulfilled; and then shall Jerusalem be inhabited again with my people, and it shall be the land of their inheritance. \fi
Verily, verily, I say unto you, all these things shall surely come, even as the Father hath commanded me. Then shall this covenant which the Father hath covenanted with his people be fulfilled; and then shall Jerusalem be inhabited again with my people, and it shall be the land of their inheritance.
%Chapter 21
\bchapter
%3 Nephi 21:1
\bverse \iffalse And verily I say unto you, I give unto you a sign, that ye may know the time when these things shall be about to take place--that I shall gather in, from their long dispersion, my people, O house of Israel, and shall establish again among them my Zion; \fi
And verily I say unto you, I give unto you a sign, that ye may know the time when these things shall be about to take place--that I shall gather in, from their long dispersion, my people, O house of Israel, and shall establish again among them my Zion;
%3 Nephi 21:2
\bverse \iffalse And behold, this is the thing which I will give unto you for a sign--for verily I say unto you that when these things which I declare unto you, and which I shall declare unto you hereafter of myself, and by the power of the Holy Ghost which shall be given unto you of the Father, shall be made known unto the Gentiles that they may know concerning this people who are a remnant of the house of Jacob, and concerning this my people who shall be scattered by them; \fi
And behold, this is the thing which I will give unto you for a sign--for verily I say unto you that when these things which I declare unto you, and which I shall declare unto you hereafter of myself, and by the power of the Holy Ghost which shall be given unto you of the Father, shall be made known unto the Gentiles that they may know concerning this people who are a remnant of the house of Jacob, and concerning this my people who shall be scattered by them;
%3 Nephi 21:3
\bverse \iffalse Verily, verily, I say unto you, when these things shall be made known unto them of the Father, and shall come forth of the Father, from them unto you; \fi
Verily, verily, I say unto you, when these things shall be made known unto them of the Father, and shall come forth of the Father, from them unto you;
%3 Nephi 21:4
\bverse \iffalse For it is wisdom in the Father that they should be established in this land, and be set up as a free people by the power of the Father, that these things might come forth from them unto a remnant of your seed, that the covenant of the Father may be fulfilled which he hath covenanted with his people, O house of Israel; \fi
For it is wisdom in the Father that they should be established in this land, and be set up as a free people by the power of the Father, that these things might come forth from them unto a remnant of your seed, that the covenant of the Father may be fulfilled which he hath covenanted with his people, O house of Israel;
%3 Nephi 21:5
\bverse \iffalse Therefore, when these works and the works which shall be wrought among you hereafter shall come forth from the Gentiles, unto your seed which shall dwindle in unbelief because of iniquity; \fi
Therefore, when these works and the works which shall be wrought among you hereafter shall come forth from the Gentiles, unto your seed which shall dwindle in unbelief because of iniquity;
%3 Nephi 21:6
\bverse \iffalse For thus it behooveth the Father that it should come forth from the Gentiles, that he may show forth his power unto the Gentiles, for this cause that the Gentiles, if they will not harden their hearts, that they may repent and come unto me and be baptized in my name and know of the true points of my doctrine, that they may be numbered among my people, O house of Israel; \fi
For thus it behooveth the Father that it should come forth from the Gentiles, that he may show forth his power unto the Gentiles, for this cause that the Gentiles, if they will not harden their hearts, that they may repent and come unto me and be baptized in my name and know of the true points of my doctrine, that they may be numbered among my people, O house of Israel;
%3 Nephi 21:7
\bverse \iffalse And when these things come to pass that thy seed shall begin to know these things--it shall be a sign unto them, that they may know that the work of the Father hath already commenced unto the fulfilling of the covenant which he hath made unto the people who are of the house of Israel. \fi
And when these things come to pass that thy seed shall begin to know these things--it shall be a sign unto them, that they may know that the work of the Father hath already commenced unto the fulfilling of the covenant which he hath made unto the people who are of the house of Israel.
%3 Nephi 21:8
\bverse \iffalse And when that day shall come, it shall come to pass that kings shall shut their mouths; for that which had not been told them shall they see; and that which they had not heard shall they consider. \fi
And when that day shall come, it shall come to pass that kings shall shut their mouths; for that which had not been told them shall they see; and that which they had not heard shall they consider.
%3 Nephi 21:9
\bverse \iffalse For in that day, for my sake shall the Father work a work, which shall be a great and a marvelous work among them; and there shall be among them those who will not believe it, although a man shall declare it unto them. \fi
For in that day, for my sake shall the Father work a work, which shall be a great and a marvelous work among them; and there shall be among them those who will not believe it, although a man shall declare it unto them.
%3 Nephi 21:10
\bverse \iffalse But behold, the life of my servant shall be in my hand; therefore they shall not hurt him, although he shall be marred because of them. Yet I will heal him, for I will show unto them that my wisdom is greater than the cunning of the devil. \fi
But behold, the life of my servant shall be in my hand; therefore they shall not hurt him, although he shall be marred because of them. Yet I will heal him, for I will show unto them that my wisdom is greater than the cunning of the devil.
%3 Nephi 21:11
\bverse \iffalse Therefore it shall come to pass that whosoever will not believe in my words, who am Jesus Christ, which the Father shall cause him to bring forth unto the Gentiles, and shall give unto him power that he shall bring them forth unto the Gentiles, (it shall be done even as Moses said) they shall be cut off from among my people who are of the covenant. \fi
Therefore it shall come to pass that whosoever will not believe in my words, who am Jesus Christ, which the Father shall cause him to bring forth unto the Gentiles, and shall give unto him power that he shall bring them forth unto the Gentiles, (it shall be done even as Moses said) they shall be cut off from among my people who are of the covenant.
%3 Nephi 21:12
\bverse \iffalse And my people who are a remnant of Jacob shall be among the Gentiles, yea, in the midst of them as a lion among the beasts of the forest, as a young lion among the flocks of sheep, who, if he go through both treadeth down and teareth in pieces, and none can deliver. \fi
And my people who are a remnant of Jacob shall be among the Gentiles, yea, in the midst of them as a lion among the beasts of the forest, as a young lion among the flocks of sheep, who, if he go through both treadeth down and teareth in pieces, and none can deliver.
%3 Nephi 21:13
\bverse \iffalse Their hand shall be lifted up upon their adversaries, and all their enemies shall be cut off. \fi
Their hand shall be lifted up upon their adversaries, and all their enemies shall be cut off.
%3 Nephi 21:14
\bverse \iffalse Yea, wo be unto the Gentiles except they repent; for it shall come to pass in that day, saith the Father, that I will cut off thy horses out of the midst of thee, and I will destroy thy chariots; \fi
Yea, wo be unto the Gentiles except they repent; for it shall come to pass in that day, saith the Father, that I will cut off thy horses out of the midst of thee, and I will destroy thy chariots;
%3 Nephi 21:15
\bverse \iffalse And I will cut off the cities of thy land, and throw down all thy strongholds; \fi
And I will cut off the cities of thy land, and throw down all thy strongholds;
%3 Nephi 21:16
\bverse \iffalse And I will cut off witchcrafts out of thy land, and thou shalt have no more soothsayers; \fi
And I will cut off witchcrafts out of thy land, and thou shalt have no more soothsayers;
%3 Nephi 21:17
\bverse \iffalse Thy graven images I will also cut off, and thy standing images out of the midst of thee, and thou shalt no more worship the works of thy hands; \fi
Thy graven images I will also cut off, and thy standing images out of the midst of thee, and thou shalt no more worship the works of thy hands;
%3 Nephi 21:18
\bverse \iffalse And I will pluck up thy groves out of the midst of thee; so will I destroy thy cities. \fi
And I will pluck up thy groves out of the midst of thee; so will I destroy thy cities.
%3 Nephi 21:19
\bverse \iffalse And it shall come to pass that all lyings, and deceivings, and envyings, and strifes, and priestcrafts, and whoredoms, shall be done away. \fi
And it shall come to pass that all lyings, and deceivings, and envyings, and strifes, and priestcrafts, and whoredoms, shall be done away.
%3 Nephi 21:20
\bverse \iffalse For it shall come to pass, saith the Father, that at that day whosoever will not repent and come unto my Beloved Son, them will I cut off from among my people, O house of Israel; \fi
For it shall come to pass, saith the Father, that at that day whosoever will not repent and come unto my Beloved Son, them will I cut off from among my people, O house of Israel;
%3 Nephi 21:21
\bverse \iffalse And I will execute vengeance and fury upon them, even as upon the heathen, such as they have not heard. \fi
And I will execute vengeance and fury upon them, even as upon the heathen, such as they have not heard.
%3 Nephi 21:22
\bverse \iffalse But if they will repent and hearken unto my words, and harden not their hearts, I will establish my church among them, and they shall come in unto the covenant and be numbered among this the remnant of Jacob, unto whom I have given this land for their inheritance; \fi
But if they will repent and hearken unto my words, and harden not their hearts, I will establish my church among them, and they shall come in unto the covenant and be numbered among this the remnant of Jacob, unto whom I have given this land for their inheritance;
%3 Nephi 21:23
\bverse \iffalse And they shall assist my people, the remnant of Jacob, and also as many of the house of Israel as shall come, that they may build a city, which shall be called the New Jerusalem. \fi
And they shall assist my people, the remnant of Jacob, and also as many of the house of Israel as shall come, that they may build a city, which shall be called the New Jerusalem.
%3 Nephi 21:24
\bverse \iffalse And then shall they assist my people that they may be gathered in, who are scattered upon all the face of the land, in unto the New Jerusalem. \fi
And then shall they assist my people that they may be gathered in, who are scattered upon all the face of the land, in unto the New Jerusalem.
%3 Nephi 21:25
\bverse \iffalse And then shall the power of heaven come down among them; and I also will be in the midst. \fi
And then shall the power of heaven come down among them; and I also will be in the midst.
%3 Nephi 21:26
\bverse \iffalse And then shall the work of the Father commence at that day, even when this gospel shall be preached among the remnant of this people. Verily I say unto you, at that day shall the work of the Father commence among all the dispersed of my people, yea, even the tribes which have been lost, which the Father hath led away out of Jerusalem. \fi
And then shall the work of the Father commence at that day, even when this gospel shall be preached among the remnant of this people. Verily I say unto you, at that day shall the work of the Father commence among all the dispersed of my people, yea, even the tribes which have been lost, which the Father hath led away out of Jerusalem.
%3 Nephi 21:27
\bverse \iffalse Yea, the work shall commence among all the dispersed of my people, with the Father to prepare the way whereby they may come unto me, that they may call on the Father in my name. \fi
Yea, the work shall commence among all the dispersed of my people, with the Father to prepare the way whereby they may come unto me, that they may call on the Father in my name.
%3 Nephi 21:28
\bverse \iffalse Yea, and then shall the work commence, with the Father among all nations in preparing the way whereby his people may be gathered home to the land of their inheritance. \fi
Yea, and then shall the work commence, with the Father among all nations in preparing the way whereby his people may be gathered home to the land of their inheritance.
%3 Nephi 21:29
\bverse \iffalse And they shall go out from all nations; and they shall not go out in haste, nor go by flight, for I will go before them, saith the Father, and I will be their rearward. \fi
And they shall go out from all nations; and they shall not go out in haste, nor go by flight, for I will go before them, saith the Father, and I will be their rearward.
%Chapter 22
\bchapter
%3 Nephi 22:1
\bverse \iffalse And then shall that which is written come to pass: Sing, O barren, thou that didst not bear; break forth into singing, and cry aloud, thou that didst not travail with child; for more are the children of the desolate than the children of the married wife, saith the Lord. \fi
And then shall that which is written come to pass: Sing, O barren, thou that didst not bear; break forth into singing, and cry aloud, thou that didst not travail with child; for more are the children of the desolate than the children of the married wife, saith the Lord.
%3 Nephi 22:2
\bverse \iffalse Enlarge the place of thy tent, and let them stretch forth the curtains of thy habitations; spare not, lengthen thy cords and strengthen thy stakes; \fi
Enlarge the place of thy tent, and let them stretch forth the curtains of thy habitations; spare not, lengthen thy cords and strengthen thy stakes;
%3 Nephi 22:3
\bverse \iffalse For thou shalt break forth on the right hand and on the left, and thy seed shall inherit the Gentiles and make the desolate cities to be inhabited. \fi
For thou shalt break forth on the right hand and on the left, and thy seed shall inherit the Gentiles and make the desolate cities to be inhabited.
%3 Nephi 22:4
\bverse \iffalse Fear not, for thou shalt not be ashamed; neither be thou confounded, for thou shalt not be put to shame; for thou shalt forget the shame of thy youth, and shalt not remember the reproach of thy youth, and shalt not remember the reproach of thy widowhood any more. \fi
Fear not, for thou shalt not be ashamed; neither be thou confounded, for thou shalt not be put to shame; for thou shalt forget the shame of thy youth, and shalt not remember the reproach of thy youth, and shalt not remember the reproach of thy widowhood any more.
%3 Nephi 22:5
\bverse \iffalse For thy maker, thy husband, the Lord of Hosts is his name; and thy Redeemer, the Holy One of Israel--the God of the whole earth shall he be called. \fi
For thy maker, thy husband, the Lord of Hosts is his name; and thy Redeemer, the Holy One of Israel--the God of the whole earth shall he be called.
%3 Nephi 22:6
\bverse \iffalse For the Lord hath called thee as a woman forsaken and grieved in spirit, and a wife of youth, when thou wast refused, saith thy God. \fi
For the Lord hath called thee as a woman forsaken and grieved in spirit, and a wife of youth, when thou wast refused, saith thy God.
%3 Nephi 22:7
\bverse \iffalse For a small moment have I forsaken thee, but with great mercies will I gather thee. \fi
For a small moment have I forsaken thee, but with great mercies will I gather thee.
%3 Nephi 22:8
\bverse \iffalse In a little wrath I hid my face from thee for a moment, but with everlasting kindness will I have mercy on thee, saith the Lord thy Redeemer. \fi
In a little wrath I hid my face from thee for a moment, but with everlasting kindness will I have mercy on thee, saith the Lord thy Redeemer.
%3 Nephi 22:9
\bverse \iffalse For this, the waters of Noah unto me, for as I have sworn that the waters of Noah should no more go over the earth, so have I sworn that I would not be wroth with thee. \fi
For this, the waters of Noah unto me, for as I have sworn that the waters of Noah should no more go over the earth, so have I sworn that I would not be wroth with thee.
%3 Nephi 22:10
\bverse \iffalse For the mountains shall depart and the hills be removed, but my kindness shall not depart from thee, neither shall the covenant of my peace be removed, saith the Lord that hath mercy on thee. \fi
For the mountains shall depart and the hills be removed, but my kindness shall not depart from thee, neither shall the covenant of my peace be removed, saith the Lord that hath mercy on thee.
%3 Nephi 22:11
\bverse \iffalse O thou afflicted, tossed with tempest, and not comforted!  Behold, I will lay thy stones with fair colors, and lay thy foundations with sapphires. \fi
O thou afflicted, tossed with tempest, and not comforted!  Behold, I will lay thy stones with fair colors, and lay thy foundations with sapphires.
%3 Nephi 22:12
\bverse \iffalse And I will make thy windows of agates, and thy gates of carbuncles, and all thy borders of pleasant stones. \fi
And I will make thy windows of agates, and thy gates of carbuncles, and all thy borders of pleasant stones.
%3 Nephi 22:13
\bverse \iffalse And all thy children shall be taught of the Lord; and great shall be the peace of thy children. \fi
And all thy children shall be taught of the Lord; and great shall be the peace of thy children.
%3 Nephi 22:14
\bverse \iffalse In righteousness shalt thou be established; thou shalt be far from oppression for thou shalt not fear, and from terror for it shall not come near thee. \fi
In righteousness shalt thou be established; thou shalt be far from oppression for thou shalt not fear, and from terror for it shall not come near thee.
%3 Nephi 22:15
\bverse \iffalse Behold, they shall surely gather together against thee, not by me; whosoever shall gather together against thee shall fall for thy sake. \fi
Behold, they shall surely gather together against thee, not by me; whosoever shall gather together against thee shall fall for thy sake.
%3 Nephi 22:16
\bverse \iffalse Behold, I have created the smith that bloweth the coals in the fire, and that bringeth forth an instrument for his work; and I have created the waster to destroy. \fi
Behold, I have created the smith that bloweth the coals in the fire, and that bringeth forth an instrument for his work; and I have created the waster to destroy.
%3 Nephi 22:17
\bverse \iffalse No weapon that is formed against thee shall prosper; and every tongue that shall revile against thee in judgment thou shalt condemn. This is the heritage of the servants of the Lord, and their righteousness is of me, saith the Lord. \fi
No weapon that is formed against thee shall prosper; and every tongue that shall revile against thee in judgment thou shalt condemn. This is the heritage of the servants of the Lord, and their righteousness is of me, saith the Lord.
%Chapter 23
\bchapter
%3 Nephi 23:1
\bverse \iffalse And now, behold, I say unto you, that ye ought to search these things. Yea, a commandment I give unto you that ye search these things diligently; for great are the words of Isaiah. \fi
And now, behold, I say unto you, that ye ought to search these things. Yea, a commandment I give unto you that ye search these things diligently; for great are the words of Isaiah.
%3 Nephi 23:2
\bverse \iffalse For surely he spake as touching all things concerning my people which are of the house of Israel; therefore it must needs be that he must speak also to the Gentiles. \fi
For surely he spake as touching all things concerning my people which are of the house of Israel; therefore it must needs be that he must speak also to the Gentiles.
%3 Nephi 23:3
\bverse \iffalse And all things that he spake have been and shall be, even according to the words which he spake. \fi
And all things that he spake have been and shall be, even according to the words which he spake.
%3 Nephi 23:4
\bverse \iffalse Therefore give heed to my words; write the things which I have told you; and according to the time and the will of the Father they shall go forth unto the Gentiles. \fi
Therefore give heed to my words; write the things which I have told you; and according to the time and the will of the Father they shall go forth unto the Gentiles.
%3 Nephi 23:5
\bverse \iffalse And whosoever will hearken unto my words and repenteth and is baptized, the same shall be saved. Search the prophets, for many there be that testify of these things. \fi
And whosoever will hearken unto my words and repenteth and is baptized, the same shall be saved. Search the prophets, for many there be that testify of these things.
%3 Nephi 23:6
\bverse \iffalse And now it came to pass that when Jesus had said these words he said unto them again, after he had expounded all the scriptures unto them which they had received, he said unto them: Behold, other scriptures I would that ye should write, that ye have not. \fi
And now it came to pass that when Jesus had said these words he said unto them again, after he had expounded all the scriptures unto them which they had received, he said unto them: Behold, other scriptures I would that ye should write, that ye have not.
%3 Nephi 23:7
\bverse \iffalse And it came to pass that he said unto Nephi: Bring forth the record which ye have kept. \fi
And it came to pass that he said unto Nephi: Bring forth the record which ye have kept.
%3 Nephi 23:8
\bverse \iffalse And when Nephi had brought forth the records, and laid them before him, he cast his eyes upon them and said: \fi
And when Nephi had brought forth the records, and laid them before him, he cast his eyes upon them and said:
%3 Nephi 23:9
\bverse \iffalse Verily I say unto you, I commanded my servant Samuel, the Lamanite, that he should testify unto this people, that at the day that the Father should glorify his name in me that there were many saints who should arise from the dead, and should appear unto many, and should minister unto them. And he said unto them: Was it not so? \fi
Verily I say unto you, I commanded my servant Samuel, the Lamanite, that he should testify unto this people, that at the day that the Father should glorify his name in me that there were many saints who should arise from the dead, and should appear unto many, and should minister unto them. And he said unto them: Was it not so?
%3 Nephi 23:10
\bverse \iffalse And his disciples answered him and said: Yea, Lord, Samuel did prophesy according to thy words, and they were all fulfilled. \fi
And his disciples answered him and said: Yea, Lord, Samuel did prophesy according to thy words, and they were all fulfilled.
%3 Nephi 23:11
\bverse \iffalse And Jesus said unto them: How be it that ye have not written this thing, that many saints did arise and appear unto many and did minister unto them? \fi
And Jesus said unto them: How be it that ye have not written this thing, that many saints did arise and appear unto many and did minister unto them?
%3 Nephi 23:12
\bverse \iffalse And it came to pass that Nephi remembered that this thing had not been written. \fi
And it came to pass that Nephi remembered that this thing had not been written.
%3 Nephi 23:13
\bverse \iffalse And it came to pass that Jesus commanded that it should be written; therefore it was written according as he commanded. \fi
And it came to pass that Jesus commanded that it should be written; therefore it was written according as he commanded.
%3 Nephi 23:14
\bverse \iffalse And now it came to pass that when Jesus had expounded all the scriptures in one, which they had written, he commanded them that they should teach the things which he had expounded unto them. \fi
And now it came to pass that when Jesus had expounded all the scriptures in one, which they had written, he commanded them that they should teach the things which he had expounded unto them.
%Chapter 24
\bchapter
%3 Nephi 24:1
\bverse \iffalse And it came to pass that he commanded them that they should write the words which the Father had given unto Malachi, which he should tell unto them. And it came to pass that after they were written he expounded them. And these are the words which he did tell unto them, saying: Thus said the Father unto Malachi--Behold, I will send my messenger, and he shall prepare the way before me, and the Lord whom ye seek shall suddenly come to his temple, even the messenger of the covenant, whom ye delight in; behold, he shall come, saith the Lord of Hosts. \fi
And it came to pass that he commanded them that they should write the words which the Father had given unto Malachi, which he should tell unto them. And it came to pass that after they were written he expounded them. And these are the words which he did tell unto them, saying: Thus said the Father unto Malachi--Behold, I will send my messenger, and he shall prepare the way before me, and the Lord whom ye seek shall suddenly come to his temple, even the messenger of the covenant, whom ye delight in; behold, he shall come, saith the Lord of Hosts.
%3 Nephi 24:2
\bverse \iffalse But who may abide the day of his coming, and who shall stand when he appeareth? For he is like a refiner's fire, and like fuller's soap. \fi
But who may abide the day of his coming, and who shall stand when he appeareth? For he is like a refiner's fire, and like fuller's soap.
%3 Nephi 24:3
\bverse \iffalse And he shall sit as a refiner and purifier of silver; and he shall purify the sons of Levi, and purge them as gold and silver, that they may offer unto the Lord an offering in righteousness. \fi
And he shall sit as a refiner and purifier of silver; and he shall purify the sons of Levi, and purge them as gold and silver, that they may offer unto the Lord an offering in righteousness.
%3 Nephi 24:4
\bverse \iffalse Then shall the offering of Judah and Jerusalem be pleasant unto the Lord, as in the days of old, and as in former years. \fi
Then shall the offering of Judah and Jerusalem be pleasant unto the Lord, as in the days of old, and as in former years.
%3 Nephi 24:5
\bverse \iffalse And I will come near to you to judgment; and I will be a swift witness against the sorcerers, and against the adulterers, and against false swearers, and against those that oppress the hireling in his wages, the widow and the fatherless, and that turn aside the stranger, and fear not me, saith the Lord of Hosts. \fi
And I will come near to you to judgment; and I will be a swift witness against the sorcerers, and against the adulterers, and against false swearers, and against those that oppress the hireling in his wages, the widow and the fatherless, and that turn aside the stranger, and fear not me, saith the Lord of Hosts.
%3 Nephi 24:6
\bverse \iffalse For I am the Lord, I change not; therefore ye sons of Jacob are not consumed. \fi
For I am the Lord, I change not; therefore ye sons of Jacob are not consumed.
%3 Nephi 24:7
\bverse \iffalse Even from the days of your fathers ye are gone away from mine ordinances, and have not kept them. Return unto me and I will return unto you, saith the Lord of Hosts. But ye say: Wherein shall we return? \fi
Even from the days of your fathers ye are gone away from mine ordinances, and have not kept them. Return unto me and I will return unto you, saith the Lord of Hosts. But ye say: Wherein shall we return?
%3 Nephi 24:8
\bverse \iffalse Will a man rob God? Yet ye have robbed me. But ye say: Wherein have we robbed thee? In tithes and offerings. \fi
Will a man rob God? Yet ye have robbed me. But ye say: Wherein have we robbed thee? In tithes and offerings.
%3 Nephi 24:9
\bverse \iffalse Ye are cursed with a curse, for ye have robbed me, even this whole nation. \fi
Ye are cursed with a curse, for ye have robbed me, even this whole nation.
%3 Nephi 24:10
\bverse \iffalse Bring ye all the tithes into the storehouse, that there may be meat in my house; and prove me now herewith, saith the Lord of Hosts, if I will not open you the windows of heaven, and pour you out a blessing that there shall not be room enough to receive it. \fi
Bring ye all the tithes into the storehouse, that there may be meat in my house; and prove me now herewith, saith the Lord of Hosts, if I will not open you the windows of heaven, and pour you out a blessing that there shall not be room enough to receive it.
%3 Nephi 24:11
\bverse \iffalse And I will rebuke the devourer for your sakes, and he shall not destroy the fruits of your ground; neither shall your vine cast her fruit before the time in the fields, saith the Lord of Hosts. \fi
And I will rebuke the devourer for your sakes, and he shall not destroy the fruits of your ground; neither shall your vine cast her fruit before the time in the fields, saith the Lord of Hosts.
%3 Nephi 24:12
\bverse \iffalse And all nations shall call you blessed, for ye shall be a delightsome land, saith the Lord of Hosts. \fi
And all nations shall call you blessed, for ye shall be a delightsome land, saith the Lord of Hosts.
%3 Nephi 24:13
\bverse \iffalse Your words have been stout against me, saith the Lord. Yet ye say: What have we spoken against thee? \fi
Your words have been stout against me, saith the Lord. Yet ye say: What have we spoken against thee?
%3 Nephi 24:14
\bverse \iffalse Ye have said: It is vain to serve God, and what doth it profit that we have kept his ordinances and that we have walked mournfully before the Lord of Hosts? \fi
Ye have said: It is vain to serve God, and what doth it profit that we have kept his ordinances and that we have walked mournfully before the Lord of Hosts?
%3 Nephi 24:15
\bverse \iffalse And now we call the proud happy; yea, they that work wickedness are set up; yea, they that tempt God are even delivered. \fi
And now we call the proud happy; yea, they that work wickedness are set up; yea, they that tempt God are even delivered.
%3 Nephi 24:16
\bverse \iffalse Then they that feared the Lord spake often one to another, and the Lord hearkened and heard; and a book of remembrance was written before him for them that feared the Lord, and that thought upon his name. \fi
Then they that feared the Lord spake often one to another, and the Lord hearkened and heard; and a book of remembrance was written before him for them that feared the Lord, and that thought upon his name.
%3 Nephi 24:17
\bverse \iffalse And they shall be mine, saith the Lord of Hosts, in that day when I make up my jewels; and I will spare them as a man spareth his own son that serveth him. \fi
And they shall be mine, saith the Lord of Hosts, in that day when I make up my jewels; and I will spare them as a man spareth his own son that serveth him.
%3 Nephi 24:18
\bverse \iffalse Then shall ye return and discern between the righteous and the wicked, between him that serveth God and him that serveth him not. \fi
Then shall ye return and discern between the righteous and the wicked, between him that serveth God and him that serveth him not.
%Chapter 25
\bchapter
%3 Nephi 25:1
\bverse \iffalse For behold, the day cometh that shall burn as an oven; and all the proud, yea, and all that do wickedly, shall be stubble; and the day that cometh shall burn them up, saith the Lord of Hosts, that it shall leave them neither root nor branch. \fi
For behold, the day cometh that shall burn as an oven; and all the proud, yea, and all that do wickedly, shall be stubble; and the day that cometh shall burn them up, saith the Lord of Hosts, that it shall leave them neither root nor branch.
%3 Nephi 25:2
\bverse \iffalse But unto you that fear my name, shall the Son of Righteousness arise with healing in his wings; and ye shall go forth and grow up as calves in the stall. \fi
But unto you that fear my name, shall the Son of Righteousness arise with healing in his wings; and ye shall go forth and grow up as calves in the stall.
%3 Nephi 25:3
\bverse \iffalse And ye shall tread down the wicked; for they shall be ashes under the soles of your feet in the day that I shall do this, saith the Lord of Hosts. \fi
And ye shall tread down the wicked; for they shall be ashes under the soles of your feet in the day that I shall do this, saith the Lord of Hosts.
%3 Nephi 25:4
\bverse \iffalse Remember ye the law of Moses, my servant, which I commanded unto him in Horeb for all Israel, with the statutes and judgments. \fi
Remember ye the law of Moses, my servant, which I commanded unto him in Horeb for all Israel, with the statutes and judgments.
%3 Nephi 25:5
\bverse \iffalse Behold, I will send you Elijah the prophet before the coming of the great and dreadful day of the Lord; \fi
Behold, I will send you Elijah the prophet before the coming of the great and dreadful day of the Lord;
%3 Nephi 25:6
\bverse \iffalse And he shall turn the heart of the fathers to the children, and the heart of the children to their fathers, lest I come and smite the earth with a curse. \fi
And he shall turn the heart of the fathers to the children, and the heart of the children to their fathers, lest I come and smite the earth with a curse.
%Chapter 26
\bchapter
%3 Nephi 26:1
\bverse \iffalse And now it came to pass that when Jesus had told these things he expounded them unto the multitude; and he did expound all things unto them, both great and small. \fi
And now it came to pass that when Jesus had told these things he expounded them unto the multitude; and he did expound all things unto them, both great and small.
%3 Nephi 26:2
\bverse \iffalse And he saith: These scriptures, which ye had not with you, the Father commanded that I should give unto you; for it was wisdom in him that they should be given unto future generations. \fi
And he saith: These scriptures, which ye had not with you, the Father commanded that I should give unto you; for it was wisdom in him that they should be given unto future generations.
%3 Nephi 26:3
\bverse \iffalse And he did expound all things, even from the beginning until the time that he should come in his glory--yea, even all things which should come upon the face of the earth, even until the elements should melt with fervent heat, and the earth should be wrapt together as a scroll, and the heavens and the earth should pass away; \fi
And he did expound all things, even from the beginning until the time that he should come in his glory--yea, even all things which should come upon the face of the earth, even until the elements should melt with fervent heat, and the earth should be wrapt together as a scroll, and the heavens and the earth should pass away;
%3 Nephi 26:4
\bverse \iffalse And even unto the great and last day, when all people, and all kindreds, and all nations and tongues shall stand before God, to be judged of their works, whether they be good or whether they be evil-- \fi
And even unto the great and last day, when all people, and all kindreds, and all nations and tongues shall stand before God, to be judged of their works, whether they be good or whether they be evil--
%3 Nephi 26:5
\bverse \iffalse If they be good, to the resurrection of everlasting life; and if they be evil, to the resurrection of damnation; being on a parallel, the one on the one hand and the other on the other hand, according to the mercy, and the justice, and the holiness which is in Christ, who was before the world began. \fi
If they be good, to the resurrection of everlasting life; and if they be evil, to the resurrection of damnation; being on a parallel, the one on the one hand and the other on the other hand, according to the mercy, and the justice, and the holiness which is in Christ, who was before the world began.
%3 Nephi 26:6
\bverse \iffalse And now there cannot be written in this book even a hundredth part of the things which Jesus did truly teach unto the people; \fi
And now there cannot be written in this book even a hundredth part of the things which Jesus did truly teach unto the people;
%3 Nephi 26:7
\bverse \iffalse But behold the plates of Nephi do contain the more part of the things which he taught the people. \fi
But behold the plates of Nephi do contain the more part of the things which he taught the people.
%3 Nephi 26:8
\bverse \iffalse And these things have I written, which are a lesser part of the things which he taught the people; and I have written them to the intent that they may be brought again unto this people, from the Gentiles, according to the words which Jesus hath spoken. \fi
And these things have I written, which are a lesser part of the things which he taught the people; and I have written them to the intent that they may be brought again unto this people, from the Gentiles, according to the words which Jesus hath spoken.
%3 Nephi 26:9
\bverse \iffalse And when they shall have received this, which is expedient that they should have first, to try their faith, and if it shall so be that they shall believe these things then shall the greater things be made manifest unto them. \fi
And when they shall have received this, which is expedient that they should have first, to try their faith, and if it shall so be that they shall believe these things then shall the greater things be made manifest unto them.
%3 Nephi 26:10
\bverse \iffalse And if it so be that they will not believe these things, then shall the greater things be withheld from them, unto their condemnation. \fi
And if it so be that they will not believe these things, then shall the greater things be withheld from them, unto their condemnation.
%3 Nephi 26:11
\bverse \iffalse Behold, I was about to write them, all which were engraven upon the plates of Nephi, but the Lord forbade it, saying: I will try the faith of my people. \fi
Behold, I was about to write them, all which were engraven upon the plates of Nephi, but the Lord forbade it, saying: I will try the faith of my people.
%3 Nephi 26:12
\bverse \iffalse Therefore I, Mormon, do write the things which have been commanded me of the Lord. And now I, Mormon, make an end of my sayings, and proceed to write the things which have been commanded me. \fi
Therefore I, Mormon, do write the things which have been commanded me of the Lord. And now I, Mormon, make an end of my sayings, and proceed to write the things which have been commanded me.
%3 Nephi 26:13
\bverse \iffalse Therefore, I would that ye should behold that the Lord truly did teach the people, for the space of three days; and after that he did show himself unto them oft, and did break bread oft, and bless it, and give it unto them. \fi
Therefore, I would that ye should behold that the Lord truly did teach the people, for the space of three days; and after that he did show himself unto them oft, and did break bread oft, and bless it, and give it unto them.
%3 Nephi 26:14
\bverse \iffalse And it came to pass that he did teach and minister unto the children of the multitude of whom hath been spoken, and he did loose their tongues, and they did speak unto their fathers great and marvelous things, even greater than he had revealed unto the people; and he loosed their tongues that they could utter. \fi
And it came to pass that he did teach and minister unto the children of the multitude of whom hath been spoken, and he did loose their tongues, and they did speak unto their fathers great and marvelous things, even greater than he had revealed unto the people; and he loosed their tongues that they could utter.
%3 Nephi 26:15
\bverse \iffalse And it came to pass that after he had ascended into heaven--the second time that he showed himself unto them, and had gone unto the Father, after having healed all their sick, and their lame, and opened the eyes of their blind and unstopped the ears of the deaf, and even had done all manner of cures among them, and raised a man from the dead, and had shown forth his power unto them, and had ascended unto the Father-- \fi
And it came to pass that after he had ascended into heaven--the second time that he showed himself unto them, and had gone unto the Father, after having healed all their sick, and their lame, and opened the eyes of their blind and unstopped the ears of the deaf, and even had done all manner of cures among them, and raised a man from the dead, and had shown forth his power unto them, and had ascended unto the Father--
%3 Nephi 26:16
\bverse \iffalse Behold, it came to pass on the morrow that the multitude gathered themselves together, and they both saw and heard these children; yea, even babes did open their mouths and utter marvelous things; and the things which they did utter were forbidden that there should not any man write them. \fi
Behold, it came to pass on the morrow that the multitude gathered themselves together, and they both saw and heard these children; yea, even babes did open their mouths and utter marvelous things; and the things which they did utter were forbidden that there should not any man write them.
%3 Nephi 26:17
\bverse \iffalse And it came to pass that the disciples whom Jesus had chosen began from that time forth to baptize and to teach as many as did come unto them; and as many as were baptized in the name of Jesus were filled with the Holy Ghost. \fi
And it came to pass that the disciples whom Jesus had chosen began from that time forth to baptize and to teach as many as did come unto them; and as many as were baptized in the name of Jesus were filled with the Holy Ghost.
%3 Nephi 26:18
\bverse \iffalse And many of them saw and heard unspeakable things, which are not lawful to be written. \fi
And many of them saw and heard unspeakable things, which are not lawful to be written.
%3 Nephi 26:19
\bverse \iffalse And they taught, and did minister one to another; and they had all things common among them, every man dealing justly, one with another. \fi
And they taught, and did minister one to another; and they had all things common among them, every man dealing justly, one with another.
%3 Nephi 26:20
\bverse \iffalse And it came to pass that they did do all things even as Jesus had commanded them. \fi
And it came to pass that they did do all things even as Jesus had commanded them.
%3 Nephi 26:21
\bverse \iffalse And they who were baptized in the name of Jesus were called the church of Christ. \fi
And they who were baptized in the name of Jesus were called the church of Christ.
%Chapter 27
\bchapter
%3 Nephi 27:1
\bverse \iffalse And it came to pass that as the disciples of Jesus were journeying and were preaching the things which they had both heard and seen, and were baptizing in the name of Jesus, it came to pass that the disciples were gathered together and were united in mighty prayer and fasting. \fi
And it came to pass that as the disciples of Jesus were journeying and were preaching the things which they had both heard and seen, and were baptizing in the name of Jesus, it came to pass that the disciples were gathered together and were united in mighty prayer and fasting.
%3 Nephi 27:2
\bverse \iffalse And Jesus again showed himself unto them, for they were praying unto the Father in his name; and Jesus came and stood in the midst of them, and said unto them: What will ye that I shall give unto you? \fi
And Jesus again showed himself unto them, for they were praying unto the Father in his name; and Jesus came and stood in the midst of them, and said unto them: What will ye that I shall give unto you?
%3 Nephi 27:3
\bverse \iffalse And they said unto him: Lord, we will that thou wouldst tell us the name whereby we shall call this church; for there are disputations among the people concerning this matter. \fi
And they said unto him: Lord, we will that thou wouldst tell us the name whereby we shall call this church; for there are disputations among the people concerning this matter.
%3 Nephi 27:4
\bverse \iffalse And the Lord said unto them: Verily, verily, I say unto you, why is it that the people should murmur and dispute because of this thing? \fi
And the Lord said unto them: Verily, verily, I say unto you, why is it that the people should murmur and dispute because of this thing?
%3 Nephi 27:5
\bverse \iffalse Have they not read the scriptures, which say ye must take upon you the name of Christ, which is my name? For by this name shall ye be called at the last day; \fi
Have they not read the scriptures, which say ye must take upon you the name of Christ, which is my name? For by this name shall ye be called at the last day;
%3 Nephi 27:6
\bverse \iffalse And whoso taketh upon him my name, and endureth to the end, the same shall be saved at the last day. \fi
And whoso taketh upon him my name, and endureth to the end, the same shall be saved at the last day.
%3 Nephi 27:7
\bverse \iffalse Therefore, whatsoever ye shall do, ye shall do it in my name; therefore ye shall call the church in my name; and ye shall call upon the Father in my name that he will bless the church for my sake. \fi
Therefore, whatsoever ye shall do, ye shall do it in my name; therefore ye shall call the church in my name; and ye shall call upon the Father in my name that he will bless the church for my sake.
%3 Nephi 27:8
\bverse \iffalse And how be it my church save it be called in my name? For if a church be called in Moses' name then it be Moses' church; or if it be called in the name of a man then it be the church of a man; but if it be called in my name then it is my church, if it so be that they are built upon my gospel. \fi
And how be it my church save it be called in my name? For if a church be called in Moses' name then it be Moses' church; or if it be called in the name of a man then it be the church of a man; but if it be called in my name then it is my church, if it so be that they are built upon my gospel.
%3 Nephi 27:9
\bverse \iffalse Verily I say unto you, that ye are built upon my gospel; therefore ye shall call whatsoever things ye do call, in my name; therefore if ye call upon the Father, for the church, if it be in my name the Father will hear you; \fi
Verily I say unto you, that ye are built upon my gospel; therefore ye shall call whatsoever things ye do call, in my name; therefore if ye call upon the Father, for the church, if it be in my name the Father will hear you;
%3 Nephi 27:10
\bverse \iffalse And if it so be that the church is built upon my gospel then will the Father show forth his own works in it. \fi
And if it so be that the church is built upon my gospel then will the Father show forth his own works in it.
%3 Nephi 27:11
\bverse \iffalse But if it be not built upon my gospel, and is built upon the works of men, or upon the works of the devil, verily I say unto you they have joy in their works for a season, and by and by the end cometh, and they are hewn down and cast into the fire, from whence there is no return. \fi
But if it be not built upon my gospel, and is built upon the works of men, or upon the works of the devil, verily I say unto you they have joy in their works for a season, and by and by the end cometh, and they are hewn down and cast into the fire, from whence there is no return.
%3 Nephi 27:12
\bverse \iffalse For their works do follow them, for it is because of their works that they are hewn down; therefore remember the things that I have told you. \fi
For their works do follow them, for it is because of their works that they are hewn down; therefore remember the things that I have told you.
%3 Nephi 27:13
\bverse \iffalse Behold I have given unto you my gospel, and this is the gospel which I have given unto you--that I came into the world to do the will of my Father, because my Father sent me. \fi
Behold I have given unto you my gospel, and this is the gospel which I have given unto you--that I came into the world to do the will of my Father, because my Father sent me.
%3 Nephi 27:14
\bverse \iffalse And my Father sent me that I might be lifted up upon the cross; and after that I had been lifted up upon the cross, that I might draw all men unto me, that as I have been lifted up by men even so should men be lifted up by the Father, to stand before me, to be judged of their works, whether they be good or whether they be evil-- \fi
And my Father sent me that I might be lifted up upon the cross; and after that I had been lifted up upon the cross, that I might draw all men unto me, that as I have been lifted up by men even so should men be lifted up by the Father, to stand before me, to be judged of their works, whether they be good or whether they be evil--
%3 Nephi 27:15
\bverse \iffalse And for this cause have I been lifted up; therefore, according to the power of the Father I will draw all men unto me, that they may be judged according to their works. \fi
And for this cause have I been lifted up; therefore, according to the power of the Father I will draw all men unto me, that they may be judged according to their works.
%3 Nephi 27:16
\bverse \iffalse And it shall come to pass, that whoso repenteth and is baptized in my name shall be filled; and if he endureth to the end, behold, him will I hold guiltless before my Father at that day when I shall stand to judge the world. \fi
And it shall come to pass, that whoso repenteth and is baptized in my name shall be filled; and if he endureth to the end, behold, him will I hold guiltless before my Father at that day when I shall stand to judge the world.
%3 Nephi 27:17
\bverse \iffalse And he that endureth not unto the end, the same is he that is also hewn down and cast into the fire, from whence they can no more return, because of the justice of the Father. \fi
And he that endureth not unto the end, the same is he that is also hewn down and cast into the fire, from whence they can no more return, because of the justice of the Father.
%3 Nephi 27:18
\bverse \iffalse And this is the word which he hath given unto the children of men. And for this cause he fulfilleth the words which he hath given, and he lieth not, but fulfilleth all his words. \fi
And this is the word which he hath given unto the children of men. And for this cause he fulfilleth the words which he hath given, and he lieth not, but fulfilleth all his words.
%3 Nephi 27:19
\bverse \iffalse And no unclean thing can enter into his kingdom; therefore nothing entereth into his rest save it be those who have washed their garments in my blood, because of their faith, and the repentance of all their sins, and their faithfulness unto the end. \fi
And no unclean thing can enter into his kingdom; therefore nothing entereth into his rest save it be those who have washed their garments in my blood, because of their faith, and the repentance of all their sins, and their faithfulness unto the end.
%3 Nephi 27:20
\bverse \iffalse Now this is the commandment: Repent, all ye ends of the earth, and come unto me and be baptized in my name, that ye may be sanctified by the reception of the Holy Ghost, that ye may stand spotless before me at the last day. \fi
Now this is the commandment: Repent, all ye ends of the earth, and come unto me and be baptized in my name, that ye may be sanctified by the reception of the Holy Ghost, that ye may stand spotless before me at the last day.
%3 Nephi 27:21
\bverse \iffalse Verily, verily, I say unto you, this is my gospel; and ye know the things that ye must do in my church; for the works which ye have seen me do that shall ye also do; for that which ye have seen me do even that shall ye do; \fi
Verily, verily, I say unto you, this is my gospel; and ye know the things that ye must do in my church; for the works which ye have seen me do that shall ye also do; for that which ye have seen me do even that shall ye do;
%3 Nephi 27:22
\bverse \iffalse Therefore, if ye do these things blessed are ye, for ye shall be lifted up at the last day. \fi
Therefore, if ye do these things blessed are ye, for ye shall be lifted up at the last day.
%3 Nephi 27:23
\bverse \iffalse Write the things which ye have seen and heard, save it be those which are forbidden. \fi
Write the things which ye have seen and heard, save it be those which are forbidden.
%3 Nephi 27:24
\bverse \iffalse Write the works of this people, which shall be, even as hath been written, of that which hath been. \fi
Write the works of this people, which shall be, even as hath been written, of that which hath been.
%3 Nephi 27:25
\bverse \iffalse For behold, out of the books which have been written, and which shall be written, shall this people be judged, for by them shall their works be known unto men. \fi
For behold, out of the books which have been written, and which shall be written, shall this people be judged, for by them shall their works be known unto men.
%3 Nephi 27:26
\bverse \iffalse And behold, all things are written by the Father; therefore out of the books which shall be written shall the world be judged. \fi
And behold, all things are written by the Father; therefore out of the books which shall be written shall the world be judged.
%3 Nephi 27:27
\bverse \iffalse And know ye that ye shall be judges of this people, according to the judgment which I shall give unto you, which shall be just. \fi
And know ye that ye shall be judges of this people, according to the judgment which I shall give unto you, which shall be just.
Therefore, what manner of men ought ye to be? Verily I say unto you, even as I am.

%3 Nephi 27:28
\bverse \iffalse And now I go unto the Father. And verily I say unto you, whatsoever things ye shall ask the Father in my name shall be given unto you. \fi
And now I go unto the Father. And verily I say unto you, whatsoever things ye shall ask the Father in my name shall be given unto you.
%3 Nephi 27:29
\bverse \iffalse Therefore, ask, and ye shall receive; knock, and it shall be opened unto you; for he that asketh, receiveth; and unto him that knocketh, it shall be opened. \fi
Therefore, ask, and ye shall receive; knock, and it shall be opened unto you; for he that asketh, receiveth; and unto him that knocketh, it shall be opened.
%3 Nephi 27:30
\bverse \iffalse And now, behold, my joy is great, even unto fulness, because of you, and also this generation; yea, and even the Father rejoiceth, and also all the holy angels, because of you and this generation; for none of them are lost. \fi
And now, behold, my joy is great, even unto fulness, because of you, and also this generation; yea, and even the Father rejoiceth, and also all the holy angels, because of you and this generation; for none of them are lost.
%3 Nephi 27:31
\bverse \iffalse Behold, I would that ye should understand; for I mean them who are now alive of this generation; and none of them are lost; and in them I have fulness of joy. \fi
Behold, I would that ye should understand; for I mean them who are now alive of this generation; and none of them are lost; and in them I have fulness of joy.
%3 Nephi 27:32
\bverse \iffalse But behold, it sorroweth me because of the fourth generation from this generation, for they are led away captive by him even as was the son of perdition; for they will sell me for silver and for gold, and for that which moth doth corrupt and which thieves can break through and steal. And in that day will I visit them, even in turning their works upon their own heads. \fi
But behold, it sorroweth me because of the fourth generation from this generation, for they are led away captive by him even as was the son of perdition; for they will sell me for silver and for gold, and for that which moth doth corrupt and which thieves can break through and steal. And in that day will I visit them, even in turning their works upon their own heads.
%3 Nephi 27:33
\bverse \iffalse And it came to pass that when Jesus had ended these sayings he said unto his disciples: Enter ye in at the strait gate; for strait is the gate, and narrow is the way that leads to life, and few there be that find it; but wide is the gate, and broad the way which leads to death, and many there be that travel therein, until the night cometh, wherein no man can work. \fi
And it came to pass that when Jesus had ended these sayings he said unto his disciples: Enter ye in at the strait gate; for strait is the gate, and narrow is the way that leads to life, and few there be that find it; but wide is the gate, and broad the way which leads to death, and many there be that travel therein, until the night cometh, wherein no man can work.
%Chapter 28
\bchapter
%3 Nephi 28:1
\bverse \iffalse And it came to pass when Jesus had said these words, he spake unto his disciples, one by one, saying unto them: What is it that ye desire of me, after that I am gone to the Father? \fi
And it came to pass when Jesus had said these words, he spake unto his disciples, one by one, saying unto them: What is it that ye desire of me, after that I am gone to the Father?
%3 Nephi 28:2
\bverse \iffalse And they all spake, save it were three, saying: We desire that after we have lived unto the age of man, that our ministry, wherein thou hast called us, may have an end, that we may speedily come unto thee in thy kingdom. \fi
And they all spake, save it were three, saying: We desire that after we have lived unto the age of man, that our ministry, wherein thou hast called us, may have an end, that we may speedily come unto thee in thy kingdom.
%3 Nephi 28:3
\bverse \iffalse And he said unto them: Blessed are ye because ye desired this thing of me; therefore, after that ye are seventy and two years old ye shall come unto me in my kingdom; and with me ye shall find rest. \fi
And he said unto them: Blessed are ye because ye desired this thing of me; therefore, after that ye are seventy and two years old ye shall come unto me in my kingdom; and with me ye shall find rest.
%3 Nephi 28:4
\bverse \iffalse And when he had spoken unto them, he turned himself unto the three, and said unto them: What will ye that I should do unto you, when I am gone unto the Father? \fi
And when he had spoken unto them, he turned himself unto the three, and said unto them: What will ye that I should do unto you, when I am gone unto the Father?
%3 Nephi 28:5
\bverse \iffalse And they sorrowed in their hearts, for they durst not speak unto him the things which they desired. \fi
And they sorrowed in their hearts, for they durst not speak unto him the things which they desired.
%3 Nephi 28:6
\bverse \iffalse And he said unto them: Behold, I know your thoughts, and ye have desired the thing which John, my beloved, who was with me in my ministry, before that I was lifted up by the Jews, desired of me. \fi
And he said unto them: Behold, I know your thoughts, and ye have desired the thing which John, my beloved, who was with me in my ministry, before that I was lifted up by the Jews, desired of me.
%3 Nephi 28:7
\bverse \iffalse Therefore, more blessed are ye, for ye shall never taste of death; but ye shall live to behold all the doings of the Father unto the children of men, even until all things shall be fulfilled according to the will of the Father, when I shall come in my glory with the powers of heaven. \fi
Therefore, more blessed are ye, for ye shall never taste of death; but ye shall live to behold all the doings of the Father unto the children of men, even until all things shall be fulfilled according to the will of the Father, when I shall come in my glory with the powers of heaven.
%3 Nephi 28:8
\bverse \iffalse And ye shall never endure the pains of death; but when I shall come in my glory ye shall be changed in the twinkling of an eye from mortality to immortality; and then shall ye be blessed in the kingdom of my Father. \fi
And ye shall never endure the pains of death; but when I shall come in my glory ye shall be changed in the twinkling of an eye from mortality to immortality; and then shall ye be blessed in the kingdom of my Father.
%3 Nephi 28:9
\bverse \iffalse And again, ye shall not have pain while ye shall dwell in the flesh, neither sorrow save it be for the sins of the world; and all this will I do because of the thing which ye have desired of me, for ye have desired that ye might bring the souls of men unto me, while the world shall stand. \fi
And again, ye shall not have pain while ye shall dwell in the flesh, neither sorrow save it be for the sins of the world; and all this will I do because of the thing which ye have desired of me, for ye have desired that ye might bring the souls of men unto me, while the world shall stand.
%3 Nephi 28:10
\bverse \iffalse And for this cause ye shall have fulness of joy; and ye shall sit down in the kingdom of my Father; yea, your joy shall be full, even as the Father hath given me fulness of joy; and ye shall be even as I am, and I am even as the Father; and the Father and I are one; \fi
And for this cause ye shall have fulness of joy; and ye shall sit down in the kingdom of my Father; yea, your joy shall be full, even as the Father hath given me fulness of joy; and ye shall be even as I am, and I am even as the Father; and the Father and I are one;
%3 Nephi 28:11
\bverse \iffalse And the Holy Ghost beareth record of the Father and me; and the Father giveth the Holy Ghost unto the children of men, because of me. \fi
And the Holy Ghost beareth record of the Father and me; and the Father giveth the Holy Ghost unto the children of men, because of me.
%3 Nephi 28:12
\bverse \iffalse And it came to pass that when Jesus had spoken these words, he touched every one of them with his finger save it were the three who were to tarry, and then he departed. \fi
And it came to pass that when Jesus had spoken these words, he touched every one of them with his finger save it were the three who were to tarry, and then he departed.
%3 Nephi 28:13
\bverse \iffalse And behold, the heavens were opened, and they were caught up into heaven, and saw and heard unspeakable things. \fi
And behold, the heavens were opened, and they were caught up into heaven, and saw and heard unspeakable things.
%3 Nephi 28:14
\bverse \iffalse And it was forbidden them that they should utter; neither was it given unto them power that they could utter the things which they saw and heard; \fi
And it was forbidden them that they should utter; neither was it given unto them power that they could utter the things which they saw and heard;
%3 Nephi 28:15
\bverse \iffalse And whether they were in the body or out of the body, they could not tell; for it did seem unto them like a transfiguration of them, that they were changed from this body of flesh into an immortal state, that they could behold the things of God. \fi
And whether they were in the body or out of the body, they could not tell; for it did seem unto them like a transfiguration of them, that they were changed from this body of flesh into an immortal state, that they could behold the things of God.
%3 Nephi 28:16
\bverse \iffalse But it came to pass that they did again minister upon the face of the earth; nevertheless they did not minister of the things which they had heard and seen, because of the commandment which was given them in heaven. \fi
But it came to pass that they did again minister upon the face of the earth; nevertheless they did not minister of the things which they had heard and seen, because of the commandment which was given them in heaven.
%3 Nephi 28:17
\bverse \iffalse And now, whether they were mortal or immortal, from the day of their transfiguration, I know not; \fi
And now, whether they were mortal or immortal, from the day of their transfiguration, I know not;
%3 Nephi 28:18
\bverse \iffalse But this much I know, according to the record which hath been given--they did go forth upon the face of the land, and did minister unto all the people, uniting as many to the church as would believe in their preaching; baptizing them, and as many as were baptized did receive the Holy Ghost. \fi
But this much I know, according to the record which hath been given--they did go forth upon the face of the land, and did minister unto all the people, uniting as many to the church as would believe in their preaching; baptizing them, and as many as were baptized did receive the Holy Ghost.
%3 Nephi 28:19
\bverse \iffalse And they were cast into prison by them who did not belong to the church. And the prisons could not hold them, for they were rent in twain. \fi
And they were cast into prison by them who did not belong to the church. And the prisons could not hold them, for they were rent in twain.
%3 Nephi 28:20
\bverse \iffalse And they were cast down into the earth; but they did smite the earth with the word of God, insomuch that by his power they were delivered out of the depths of the earth; and therefore they could not dig pits sufficient to hold them. \fi
And they were cast down into the earth; but they did smite the earth with the word of God, insomuch that by his power they were delivered out of the depths of the earth; and therefore they could not dig pits sufficient to hold them.
%3 Nephi 28:21
\bverse \iffalse And thrice they were cast into a furnace and received no harm. \fi
And thrice they were cast into a furnace and received no harm.
%3 Nephi 28:22
\bverse \iffalse And twice were they cast into a den of wild beasts; and behold they did play with the beasts as a child with a suckling lamb, and received no harm. \fi
And twice were they cast into a den of wild beasts; and behold they did play with the beasts as a child with a suckling lamb, and received no harm.
%3 Nephi 28:23
\bverse \iffalse And it came to pass that thus they did go forth among all the people of Nephi, and did preach the gospel of Christ unto all people upon the face of the land; and they were converted unto the Lord, and were united unto the church of Christ, and thus the people of that generation were blessed, according to the word of Jesus. \fi
And it came to pass that thus they did go forth among all the people of Nephi, and did preach the gospel of Christ unto all people upon the face of the land; and they were converted unto the Lord, and were united unto the church of Christ, and thus the people of that generation were blessed, according to the word of Jesus.
%3 Nephi 28:24
\bverse \iffalse And now I, Mormon, make an end of speaking concerning these things for a time. \fi
And now I, Mormon, make an end of speaking concerning these things for a time.
%3 Nephi 28:25
\bverse \iffalse Behold, I was about to write the names of those who were never to taste of death, but the Lord forbade; therefore I write them not, for they are hid from the world. \fi
Behold, I was about to write the names of those who were never to taste of death, but the Lord forbade; therefore I write them not, for they are hid from the world.
%3 Nephi 28:26
\bverse \iffalse But behold, I have seen them, and they have ministered unto me. \fi
But behold, I have seen them, and they have ministered unto me.
%3 Nephi 28:27
\bverse \iffalse And behold they will be among the Gentiles, and the Gentiles shall know them not. \fi
And behold they will be among the Gentiles, and the Gentiles shall know them not.
%3 Nephi 28:28
\bverse \iffalse They will also be among the Jews, and the Jews shall know them not. \fi
They will also be among the Jews, and the Jews shall know them not.
%3 Nephi 28:29
\bverse \iffalse And it shall come to pass, when the Lord seeth fit in his wisdom that they shall minister unto all the scattered tribes of Israel, and unto all nations, kindreds, tongues and people, and shall bring out of them unto Jesus many souls, that their desire may be fulfilled, and also because of the convincing power of God which is in them. \fi
And it shall come to pass, when the Lord seeth fit in his wisdom that they shall minister unto all the scattered tribes of Israel, and unto all nations, kindreds, tongues and people, and shall bring out of them unto Jesus many souls, that their desire may be fulfilled, and also because of the convincing power of God which is in them.
%3 Nephi 28:30
\bverse \iffalse And they are as the angels of God, and if they shall pray unto the Father in the name of Jesus they can show themselves unto whatsoever man it seemeth them good. \fi
And they are as the angels of God, and if they shall pray unto the Father in the name of Jesus they can show themselves unto whatsoever man it seemeth them good.
%3 Nephi 28:31
\bverse \iffalse Therefore, great and marvelous works shall be wrought by them, before the great and coming day when all people must surely stand before the judgment-seat of Christ; \fi
Therefore, great and marvelous works shall be wrought by them, before the great and coming day when all people must surely stand before the judgment-seat of Christ;
%3 Nephi 28:32
\bverse \iffalse Yea even among the Gentiles shall there be a great and marvelous work wrought by them, before that judgment day. \fi
Yea even among the Gentiles shall there be a great and marvelous work wrought by them, before that judgment day.
%3 Nephi 28:33
\bverse \iffalse And if ye had all the scriptures which give an account of all the marvelous works of Christ, ye would, according to the words of Christ, know that these things must surely come. \fi
And if ye had all the scriptures which give an account of all the marvelous works of Christ, ye would, according to the words of Christ, know that these things must surely come.
%3 Nephi 28:34
\bverse \iffalse And wo be unto him that will not hearken unto the words of Jesus, and also to them whom he hath chosen and sent among them; for whoso receiveth not the words of Jesus and the words of those whom he hath sent receiveth not him; and therefore he will not receive them at the last day; \fi
And wo be unto him that will not hearken unto the words of Jesus, and also to them whom he hath chosen and sent among them; for whoso receiveth not the words of Jesus and the words of those whom he hath sent receiveth not him; and therefore he will not receive them at the last day;
%3 Nephi 28:35
\bverse \iffalse And it would be better for them if they had not been born.  For do ye suppose that ye can get rid of the justice of an offended God, who hath been trampled under feet of men, that thereby salvation might come? \fi
And it would be better for them if they had not been born.  For do ye suppose that ye can get rid of the justice of an offended God, who hath been trampled under feet of men, that thereby salvation might come?
%3 Nephi 28:36
\bverse \iffalse And now behold, as I spake concerning those whom the Lord hath chosen, yea, even three who were caught up into the heavens, that I knew not whether they were cleansed from mortality to immortality-- \fi
And now behold, as I spake concerning those whom the Lord hath chosen, yea, even three who were caught up into the heavens, that I knew not whether they were cleansed from mortality to immortality--
%3 Nephi 28:37
\bverse \iffalse But behold, since I wrote, I have inquired of the Lord, and he hath made it manifest unto me that there must needs be a change wrought upon their bodies, or else it needs be that they must taste of death; \fi
But behold, since I wrote, I have inquired of the Lord, and he hath made it manifest unto me that there must needs be a change wrought upon their bodies, or else it needs be that they must taste of death;
%3 Nephi 28:38
\bverse \iffalse Therefore, that they might not taste of death there was a change wrought upon their bodies, that they might not suffer pain nor sorrow save it were for the sins of the world. \fi
Therefore, that they might not taste of death there was a change wrought upon their bodies, that they might not suffer pain nor sorrow save it were for the sins of the world.
%3 Nephi 28:39
\bverse \iffalse Now this change was not equal to that which shall take place at the last day; but there was a change wrought upon them, insomuch that Satan could have no power over them, that he could not tempt them; and they were sanctified in the flesh, that they were holy, and that the powers of the earth could not hold them.  3 Nephi 28:40 40 And in this state they were to remain until the judgment day of Christ; and at that day they were to receive a greater change, and to be received into the kingdom of the Father to go no more out, but to dwell eternally in the heavens. \fi
Now this change was not equal to that which shall take place at the last day; but there was a change wrought upon them, insomuch that Satan could have no power over them, that he could not tempt them; and they were sanctified in the flesh, that they were holy, and that the powers of the earth could not hold them.  3 Nephi 28:40 40 And in this state they were to remain until the judgment day of Christ; and at that day they were to receive a greater change, and to be received into the kingdom of the Father to go no more out, but to dwell eternally in the heavens.
%Chapter 29
\bchapter
%3 Nephi 29:1
\bverse \iffalse And now behold, I say unto you that when the Lord shall see fit, in his wisdom, that these sayings shall come unto the Gentiles according to his word, then ye may know that the covenant which the Father hath made with the children of Israel, concerning their restoration to the lands of their inheritance, is already beginning to be fulfilled. \fi
And now behold, I say unto you that when the Lord shall see fit, in his wisdom, that these sayings shall come unto the Gentiles according to his word, then ye may know that the covenant which the Father hath made with the children of Israel, concerning their restoration to the lands of their inheritance, is already beginning to be fulfilled.
%3 Nephi 29:2
\bverse \iffalse And ye may know that the words of the Lord, which have been spoken by the holy prophets, shall all be fulfilled; and ye need not say that the Lord delays his coming unto the children of Israel. \fi
And ye may know that the words of the Lord, which have been spoken by the holy prophets, shall all be fulfilled; and ye need not say that the Lord delays his coming unto the children of Israel.
%3 Nephi 29:3
\bverse \iffalse And ye need not imagine in your hearts that the words which have been spoken are vain, for behold, the Lord will remember his covenant which he hath made unto his people of the house of Israel. \fi
And ye need not imagine in your hearts that the words which have been spoken are vain, for behold, the Lord will remember his covenant which he hath made unto his people of the house of Israel.
%3 Nephi 29:4
\bverse \iffalse And when ye shall see these sayings coming forth among you, then ye need not any longer spurn at the doings of the Lord, for the sword of his justice is in his right hand; and behold, at that day, if ye shall spurn at his doings he will cause that it shall soon overtake you. \fi
And when ye shall see these sayings coming forth among you, then ye need not any longer spurn at the doings of the Lord, for the sword of his justice is in his right hand; and behold, at that day, if ye shall spurn at his doings he will cause that it shall soon overtake you.
%3 Nephi 29:5
\bverse \iffalse Wo unto him that spurneth at the doings of the Lord; yea, wo unto him that shall deny the Christ and his works! \fi
Wo unto him that spurneth at the doings of the Lord; yea, wo unto him that shall deny the Christ and his works!
%3 Nephi 29:6
\bverse \iffalse Yea, wo unto him that shall deny the revelations of the Lord, and that shall say the Lord no longer worketh by revelation, or by prophecy, or by gifts, or by tongues, or by healings, or by the power of the Holy Ghost! \fi
Yea, wo unto him that shall deny the revelations of the Lord, and that shall say the Lord no longer worketh by revelation, or by prophecy, or by gifts, or by tongues, or by healings, or by the power of the Holy Ghost!
%3 Nephi 29:7
\bverse \iffalse Yea, and wo unto him that shall say at that day, to get gain, that there can be no miracle wrought by Jesus Christ; for he that doeth this shall become like unto the son of perdition, for whom there was no mercy, according to the word of Christ! \fi
Yea, and wo unto him that shall say at that day, to get gain, that there can be no miracle wrought by Jesus Christ; for he that doeth this shall become like unto the son of perdition, for whom there was no mercy, according to the word of Christ!
%3 Nephi 29:8
\bverse \iffalse Yea, and ye need not any longer hiss, nor spurn, nor make game of the Jews, nor any of the remnant of the house of Israel; for behold, the Lord remembereth his covenant unto them, and he will do unto them according to that which he hath sworn. \fi
Yea, and ye need not any longer hiss, nor spurn, nor make game of the Jews, nor any of the remnant of the house of Israel; for behold, the Lord remembereth his covenant unto them, and he will do unto them according to that which he hath sworn.
%3 Nephi 29:9
\bverse \iffalse Therefore ye need not suppose that ye can turn the right hand of the Lord unto the left, that he may not execute judgment unto the fulfilling of the covenant which he hath made unto the house of Israel. \fi
Therefore ye need not suppose that ye can turn the right hand of the Lord unto the left, that he may not execute judgment unto the fulfilling of the covenant which he hath made unto the house of Israel.
%Chapter 30
\bchapter
%3 Nephi 30:1
\bverse \iffalse Hearken, O ye Gentiles, and hear the words of Jesus Christ, the Son of the living God, which he hath commanded me that I should speak concerning you, for, behold he commandeth me that I should write, saying: \fi
Hearken, O ye Gentiles, and hear the words of Jesus Christ, the Son of the living God, which he hath commanded me that I should speak concerning you, for, behold he commandeth me that I should write, saying:
%3 Nephi 30:2
\bverse \iffalse Turn, all ye Gentiles, from your wicked ways; and repent of your evil doings, of your lyings and deceivings, and of your whoredoms, and of your secret abominations, and your idolatries, and of your murders, and your priestcrafts, and your envyings, and your strifes, and from all your wickedness and abominations, and come unto me, and be baptized in my name, that ye may receive a remission of your sins, and be filled with the Holy Ghost, that ye may be numbered with my people who are of the house of Israel. \fi
Turn, all ye Gentiles, from your wicked ways; and repent of your evil doings, of your lyings and deceivings, and of your whoredoms, and of your secret abominations, and your idolatries, and of your murders, and your priestcrafts, and your envyings, and your strifes, and from all your wickedness and abominations, and come unto me, and be baptized in my name, that ye may receive a remission of your sins, and be filled with the Holy Ghost, that ye may be numbered with my people who are of the house of Israel.