\twocolumn[{%
 \centering
\bbook{THE BOOK OF MORMON}{Morm.}
 \normalsize
\vspace{0.5cm} %Put space between opening paragraph and first chapter
}]
\renewcommand\curbook{Morm.}

\balance

%Chapter 1
\bchapter
%Mormon 1:1
\bverse \iffalse And now I, Mormon, make a record of the things which I have both seen and heard, and call it the Book of Mormon. \fi
And now I, Mormon, am making a record of what I have heard and seen. I call it the Book of Mormon.
%Mormon 1:2
\bverse \iffalse And about the time that Ammaron hid up the records unto the Lord, he came unto me, (I being about ten years of age, and I began to be learned somewhat after the manner of the learning of my people) and Ammaron said unto me: I perceive that thou art a sober child, and art quick to observe; \fi
Around the time that Ammaron hid the records and asked the Lord to protect them, he came to me. I was about ten years old, and I had started to learn the ways of my people. Ammaron said to me: ``I see that you are smart, and that you pay close attention to what happens around you.
%Mormon 1:3
\bverse \iffalse Therefore, when ye are about twenty and four years old I would that ye should remember the things that ye have observed concerning this people; and when ye are of that age go to the land Antum, unto a hill which shall be called Shim; and there have I deposited unto the Lord all the sacred engravings concerning this people. \fi
So, when you are twenty-four years old, I want you to remember the things that you have seen this people doing. Go to the land of Antum, to the hill that will be called Shim. I have buried all the sacred records about this people there, and dedicated them to the Lord.
%Mormon 1:4
\bverse \iffalse And behold, ye shall take the plates of Nephi unto yourself, and the remainder shall ye leave in the place where they are; and ye shall engrave on the plates of Nephi all the things that ye have observed concerning this people. \fi
Take the plates of Nephi, and leave the rest where they are. Engrave on the plates of Nephi everything that you have observed about this people.''
%Mormon 1:5
\bverse \iffalse And I, Mormon, being a descendant of Nephi, (and my father's name was Mormon) I remembered the things which Ammaron commanded me. \fi
And I, Mormon, remembered the things that Ammaron commanded me to do. I am a descendant of Nephi, and my father's name was Mormon.
%Mormon 1:6
\bverse \iffalse And it came to pass that I, being eleven years old, was carried by my father into the land southward, even to the land of Zarahemla. \fi
When I was eleven years old, my father took me into the land southward, to the land of Zarahemla.
%Mormon 1:7
\bverse \iffalse The whole face of the land had become covered with buildings, and the people were as numerous almost, as it were the sand of the sea. \fi
The entire land there had become covered with buildings. It seemed there were as many people as there are sands on the seashore.
%Mormon 1:8
\bverse \iffalse And it came to pass in this year there began to be a war between the Nephites, who consisted of the Nephites and the Jacobites and the Josephites and the Zoramites; and this war was between the Nephites, and the Lamanites and the Lemuelites and the Ishmaelites. \fi
In this year a war started between the Nephites and the Lamanites. The Nephites were the Nephites, Jacobites, Josephites, and Zoramites.
%Mormon 1:9
\bverse \iffalse Now the Lamanites and the Lemuelites and the Ishmaelites were called Lamanites, and the two parties were Nephites and Lamanites. \fi
The Lamanites were the Lamanites, Lemuelites, and Ishmaelites. So the two groups that were fighting were the Lamanites and the Nephites.
%Mormon 1:10
\bverse \iffalse And it came to pass that the war began to be among them in the borders of Zarahemla, by the waters of Sidon. \fi
The war between them started near the borders of Zarahemla, by the waters of Sidon.
%Mormon 1:11
\bverse \iffalse And it came to pass that the Nephites had gathered together a great number of men, even to exceed the number of thirty thousand. And it came to pass that they did have in this same year a number of battles, in which the Nephites did beat the Lamanites and did slay many of them. \fi
The Nephites had gathered together a great number of people, more than thirty thousand. And during this year they had many battles, and the Nephites beat the Lamanites and killed many of them.
%Mormon 1:12
\bverse \iffalse And it came to pass that the Lamanites withdrew their design, and there was peace settled in the land; and peace did remain for the space of about four years, that there was no bloodshed. \fi
So the Lamanites stopped trying to attack the Nephites. There was peace in the land for about four years, and no one died in war during that time.
%Mormon 1:13
\bverse \iffalse But wickedness did prevail upon the face of the whole land, insomuch that the Lord did take away his beloved disciples, and the work of miracles and of healing did cease because of the iniquity of the people. \fi
But there was still great wickedness all over the land. The Lord took away his disciples that he loved, and there were no more miracles or healings because of the sins of the people.
%Mormon 1:14
\bverse \iffalse And there were no gifts from the Lord, and the Holy Ghost did not come upon any, because of their wickedness and unbelief. \fi
There were no spiritual gifts from the Lord, and the gift of the Holy Ghost was not given to anyone, because of their wickedness and unbelief.
%Mormon 1:15
\bverse \iffalse And I, being fifteen years of age and being somewhat of a sober mind, therefore I was visited of the Lord, and tasted and knew of the goodness of Jesus. \fi
I was about fifteen years old, and I had a somewhat serious mind. So I was visited by the Lord, and I knew about the goodness of Jesus and the gospel.
%Mormon 1:16
\bverse \iffalse And I did endeavor to preach unto this people, but my mouth was shut, and I was forbidden that I should preach unto them; for behold they had wilfully rebelled against their God; and the beloved disciples were taken away out of the land, because of their iniquity. \fi
I wanted to preach to the people, but the Lord commanded me not to preach to them. They chose to rebel against God, even though they knew better. The disciples that the Lord loved were taken out of the land because of the sins of the people.
%Mormon 1:17
\bverse \iffalse But I did remain among them, but I was forbidden to preach unto them, because of the hardness of their hearts; and because of the hardness of their hearts the land was cursed for their sake. \fi
I stayed among the people, but I was commanded to not preach to them because their hearts were so hard. And because their hearts were so hard, the land was cursed for their sake.
%Mormon 1:18
\bverse \iffalse And these Gadianton robbers, who were among the Lamanites, did infest the land, insomuch that the inhabitants thereof began to hide up their treasures in the earth; and they became slippery, because the Lord had cursed the land, that they could not hold them, nor retain them again. \fi
And the Gadianton robbers, who were among the Lamanites, spread all over the land. The people who lived in the land started trying to bury and hide their treasures, but they lost them. The Lord had cursed the land so that they could not keep them, or find them again when they tried to hide them.
%Mormon 1:19
\bverse \iffalse And it came to pass that there were sorceries, and witchcrafts, and magics; and the power of the evil one was wrought upon all the face of the land, even unto the fulfilling of all the words of Abinadi, and also Samuel the Lamanite. \fi
And there was evil magic, witchcraft, and sorcery. The power of the devil was over all the land, and showed that the words of Abinadi and Samuel the Lamanite had come true.
%Chapter 2
\bchapter
%Mormon 2:1
\bverse \iffalse And it came to pass in that same year there began to be a war again between the Nephites and the Lamanites. And notwithstanding I being young, was large in stature; therefore the people of Nephi appointed me that I should be their leader, or the leader of their armies. \fi
And during this year a war between the Nephites and Lamanites started. And even though I was young, I was large for my age, so the Nephites made me the leader of their armies.
%Mormon 2:2
\bverse \iffalse Therefore it came to pass that in my sixteenth year I did go forth at the head of an army of the Nephites, against the Lamanites; therefore three hundred and twenty and six years had passed away. \fi
And so, when I was sixteen years old, I led a Nephite army against the Lamanites. Three hundred and twenty-six years had gone by.
%Mormon 2:3
\bverse \iffalse And it came to pass that in the three hundred and twenty and seventh year the Lamanites did come upon us with exceedingly great power, insomuch that they did frighten my armies; therefore they would not fight, and they began to retreat towards the north countries. \fi
During the three hundred and twenty-seventh year the Lamanites attacked us. Their attack was so powerful that it frightened my armies, and they refused to fight. They began to run away into the countries to the north.
%Mormon 2:4
\bverse \iffalse And it came to pass that we did come to the city of Angola, and we did take possession of the city, and make preparations to defend ourselves against the Lamanites. And it came to pass that we did fortify the city with our might; but notwithstanding all our fortifications the Lamanites did come upon us and did drive us out of the city. \fi
And we came to the city of Angola. We took control of the city, and prepared to defend ourselves against the Lamanites. We did everything we could to strengthen our defenses, but the Lamanites still came and drove us out of the city.
%Mormon 2:5
\bverse \iffalse And they did also drive us forth out of the land of David. \fi
They chased us out of the land of David.
%Mormon 2:6
\bverse \iffalse And we marched forth and came to the land of Joshua, which was in the borders west by the seashore. \fi
And we marched and came to the land of Joshua, which was on the western border by the seashore.
%Mormon 2:7
\bverse \iffalse And it came to pass that we did gather in our people as fast as it were possible, that we might get them together in one body. \fi
We gathered our people together as fast as we could, so that everyone could be gathered into one group.
%Mormon 2:8
\bverse \iffalse But behold, the land was filled with robbers and with Lamanites; and notwithstanding the great destruction which hung over my people, they did not repent of their evil doings; therefore there was blood and carnage spread throughout all the face of the land, both on the part of the Nephites and also on the part of the Lamanites; and it was one complete revolution throughout all the face of the land. \fi
But the land was filled with robbers and Lamanites. And even though my people were about to be destroyed, they did not repent of the evil things that they were doing. There was blood and death over all the land, among the Nephites and the Lamanites. The entire land was covered in war.
%Mormon 2:9
\bverse \iffalse And now, the Lamanites had a king, and his name was Aaron; and he came against us with an army of forty and four thousand. And behold, I withstood him with forty and two thousand. And it came to pass that I beat him with my army that he fled before me. And behold, all this was done, and three hundred and thirty years had passed away. \fi
The king of the Lamanites was named Aaron. He came to battle us with an army of forty-four thousand soldiers. I fought him back with an army of forty-two thousand soldiers. I beat him with my army, and he and his army ran from us. All of this happened, and three hundred and thirty years had gone by.
%Mormon 2:10
\bverse \iffalse And it came to pass that the Nephites began to repent of their iniquity, and began to cry even as had been prophesied by Samuel the prophet; for behold no man could keep that which was his own, for the thieves, and the robbers, and the murderers, and the magic art, and the witchcraft which was in the land. \fi
Now the Nephites started to repent of the evil they had done, and they started to cry just as Samuel had prophesied. You see, no one could keep the things that they owned, because of the thieves, robbers, murderers, evil magic, and witchcraft that was in the land.
%Mormon 2:11
\bverse \iffalse Thus there began to be a mourning and a lamentation in all the land because of these things, and more especially among the people of Nephi. \fi
People all over the land, and especially the Nephites, were sad and full of sorrow because of these things.
%Mormon 2:12
\bverse \iffalse And it came to pass that when I, Mormon, saw their lamentation and their mourning and their sorrow before the Lord, my heart did begin to rejoice within me, knowing the mercies and the long-suffering of the Lord, therefore supposing that he would be merciful unto them that they would again become a righteous people. \fi
And when I, Mormon, saw that they were full of sorrow and sadness before the Lord, my heart filled with joy. I knew that the Lord is merciful and full of patient love. I thought that the Lord would be merciful to them, and that they would become a righteous people again.
%Mormon 2:13
\bverse \iffalse But behold this my joy was vain, for their sorrowing was not unto repentance, because of the goodness of God; but it was rather the sorrowing of the damned, because the Lord would not always suffer them to take happiness in sin. \fi
But I found out that my joy was empty. They were not sorry because they wanted to repent, because of God's goodness. They were sorry like a person who has already been judged guilty. The Lord was not going to always let them have joy in living sinful lives.
%Mormon 2:14
\bverse \iffalse And they did not come unto Jesus with broken hearts and contrite spirits, but they did curse God, and wish to die.  Nevertheless they would struggle with the sword for their lives. \fi
They did not come to Jesus with broken hearts that were sorry for their sins. No; they cursed God, and wished they could die. They were going to spend the rest of their lives in war, fighting with their swords.
%Mormon 2:15
\bverse \iffalse And it came to pass that my sorrow did return unto me again, and I saw that the day of grace was passed with them, both temporally and spiritually; for I saw thousands of them hewn down in open rebellion against their God, and heaped up as dung upon the face of the land. And thus three hundred and forty and four years had passed away. \fi
Then I was filled with sadness again. I saw that God's grace was no longer with them, either in this life or the next. I saw thousands of them killed while their hearts were full of rebellion against God. All over the land their bodies were piled up like dung. And three hundred and forty-four years had gone by.
%Mormon 2:16
\bverse \iffalse And it came to pass that in the three hundred and forty and fifth year the Nephites did begin to flee before the Lamanites; and they were pursued until they came even to the land of Jashon, before it was possible to stop them in their retreat. \fi
And in the three hundred and forty-fifth year the Nephites began to run to escape from the Lamanites. The Lamanites chased them all the way to the land of Jashon; then they caught up to them.
%Mormon 2:17
\bverse \iffalse And now, the city of Jashon was near the land where Ammaron had deposited the records unto the Lord, that they might not be destroyed. And behold I had gone according to the word of Ammaron, and taken the plates of Nephi, and did make a record according to the words of Ammaron. \fi
The city of Jashon was near the land where Ammaron had buried the records and asked the Lord to protect them, so that they would not be destroyed. I had done what Ammaron told me to do: I took the plates of Nephi, and made a record according to the instructions that Ammaron had given me.
%Mormon 2:18
\bverse \iffalse And upon the plates of Nephi I did make a full account of all the wickedness and abominations; but upon these plates I did forbear to make a full account of their wickedness and abominations, for behold, a continual scene of wickedness and abominations has been before mine eyes ever since I have been sufficient to behold the ways of man. \fi
I gave the full story of all the wickedness and horrible things I had seen on the plates of Nephi. But I decided not to tell the full story of the wickedness and horrible things I saw on these plates. Ever since I was old enough to understand what people around me were doing, I have seen nothing but wickedness and horribly evil things happening.
%Mormon 2:19
\bverse \iffalse And wo is me because of their wickedness; for my heart has been filled with sorrow because of their wickedness, all my days; nevertheless, I know that I shall be lifted up at the last day. \fi
It fills me with great sadness to see how wicked they are. My heart has been full of pain my whole life because of their wickedness. But I know that I will be lifted up to be with God on the last day.
%Mormon 2:20
\bverse \iffalse And it came to pass that in this year the people of Nephi again were hunted and driven. And it came to pass that we were driven forth until we had come northward to the land which was called Shem. \fi
And during this year, the people of Nephi were hunted and scattered. We were driven out until we had come northward to the land that was called Shem.
%Mormon 2:21
\bverse \iffalse And it came to pass that we did fortify the city of Shem, and we did gather in our people as much as it were possible, that perhaps we might save them from destruction. \fi
We strengthened the defenses in the city of Shem, and we gathered our people together as much as it was possible, so that perhaps we could save them from being destroyed.
%Mormon 2:22
\bverse \iffalse And it came to pass in the three hundred and forty and sixth year they began to come upon us again. \fi
And during the three hundred and forty-sixth year, they began to attack us again.
%Mormon 2:23
\bverse \iffalse And it came to pass that I did speak unto my people, and did urge them with great energy, that they would stand boldly before the Lamanites and fight for their wives, and their children, and their houses, and their homes. \fi
I spoke to my people, and encouraged them with all of my energy, to fight against the Lamanites for their families, and their houses and homes.
%Mormon 2:24
\bverse \iffalse And my words did arouse them somewhat to vigor, insomuch that they did not flee from before the Lamanites, but did stand with boldness against them. \fi
My words did at least get them to fight, so that they did not run away from the Lamanites. They fought boldly against them.
%Mormon 2:25
\bverse \iffalse And it came to pass that we did contend with an army of thirty thousand against an army of fifty thousand. And it came to pass that we did stand before them with such firmness that they did flee from before us. \fi
We fought against their army of fifty thousand soldiers with our army of thirty thousand soldiers. We fought them so fiercely that they started to run from us.
%Mormon 2:26
\bverse \iffalse And it came to pass that when they had fled we did pursue them with our armies, and did meet them again, and did beat them; nevertheless the strength of the Lord was not with us; yea, we were left to ourselves, that the Spirit of the Lord did not abide in us; therefore we had become weak like unto our brethren. \fi
When they began to run from us we chased them with our armies, and caught up to them and beat them. But the Lord's strength was not with us. We were left to fight on our own, without the Spirit of the Lord inside us. We had become weak like our sisters and brothers, the Lamanites.
%Mormon 2:27
\bverse \iffalse And my heart did sorrow because of this the great calamity of my people, because of their wickedness and their abominations.  But behold, we did go forth against the Lamanites and the robbers of Gadianton, until we had again taken possession of the lands of our inheritance. \fi
My heart was full of sorrow because of the downfall of my people. They were being destroyed because of their wickedness and horrible sins. But we went and fought against the Lamanites and the Gadianton robbers until we had gotten control of our own lands again.
%Mormon 2:28
\bverse \iffalse And the three hundred and forty and ninth year had passed away. And in the three hundred and fiftieth year we made a treaty with the Lamanites and the robbers of Gadianton, in which we did get the lands of our inheritance divided. \fi
And the three hundred and forty-ninth year ended. And in the three hundred and fiftieth year we made an agreement of peace with the Lamanites and the robbers of Gadianton. We agreed to divide up the lands where we lived.
%Mormon 2:29
\bverse \iffalse And the Lamanites did give unto us the land northward, yea, even to the narrow passage which led into the land southward.  And we did give unto the Lamanites all the land southward. \fi
The Lamanites gave us the land in the north, all the way until the narrow part of the land that leads to the land in the south. And we gave the Lamanites all the land in the south.
%Chapter 3
\bchapter
%Mormon 3:1
\bverse \iffalse And it came to pass that the Lamanites did not come to battle again until ten years more had passed away. And behold, I had employed my people, the Nephites, in preparing their lands and their arms against the time of battle. \fi
The Lamanites did not come to battle again until ten more years had gone by. During that time I had told my people, the Nephites, to prepare their lands and gather weapons for the next war.
%Mormon 3:2
\bverse \iffalse And it came to pass that the Lord did say unto me: Cry unto this people--Repent ye, and come unto me, and be ye baptized, and build up again my church, and ye shall be spared. \fi
And the Lord said to me: ``Cry out to this people---Repent, and come to me, and be baptized. Build up my church again, and you will not be destroyed.''
%Mormon 3:3
\bverse \iffalse And I did cry unto this people, but it was in vain; and they did not realize that it was the Lord that had spared them, and granted unto them a chance for repentance. And behold they did harden their hearts against the Lord their God. \fi
I cried out as the Lord had commanded me, but it was no use. They did not realize that it was the Lord that had saved them, and given them a chance to repent. And they hardened their hearts against the Lord their God.
%Mormon 3:4
\bverse \iffalse And it came to pass that after this tenth year had passed away, making, in the whole, three hundred and sixty years from the coming of Christ, the king of the Lamanites sent an epistle unto me, which gave unto me to know that they were preparing to come again to battle against us. \fi
And after this tenth year had gone by, three hundred and sixty years had gone by since Christ had come. The king of the Lamanites sent a letter to me, to let me know that they were preparing to come to battle against us.
%Mormon 3:5
\bverse \iffalse And it came to pass that I did cause my people that they should gather themselves together at the land Desolation, to a city which was in the borders, by the narrow pass which led into the land southward. \fi
I had my people gather together at the land of The Ruins, in a city near the borders by the narrow part of the land that led into the land in the south.
%Mormon 3:6
\bverse \iffalse And there we did place our armies, that we might stop the armies of the Lamanites, that they might not get possession of any of our lands; therefore we did fortify against them with all our force. \fi
We placed our armies so that we could stop the armies of the Lamanites from taking any of our lands. We used all of our armies and defenses to stop them.
%Mormon 3:7
\bverse \iffalse And it came to pass that in the three hundred and sixty and first year the Lamanites did come down to the city of Desolation to battle against us; and it came to pass that in that year we did beat them, insomuch that they did return to their own lands again. \fi
And during the three hundred and sixty-first year, the Lamanites came down to the city of The Ruins to battle against us. We beat them in battle, and they went back to their own lands.
%Mormon 3:8
\bverse \iffalse And in the three hundred and sixty and second year they did come down again to battle. And we did beat them again, and did slay a great number of them, and their dead were cast into the sea. \fi
And in the three hundred and sixty-second year they came to battle against us again. We beat them again, and killed very many of them. The dead Lamanites were thrown into the sea.
%Mormon 3:9
\bverse \iffalse And now, because of this great thing which my people, the Nephites, had done, they began to boast in their own strength, and began to swear before the heavens that they would avenge themselves of the blood of their brethren who had been slain by their enemies. \fi
And because my people, the Nephites, had defeated them, they boasted about how strong they were. They began to swear before heaven that they would take revenge for the blood of their people who had been killed by the Lamanites.
%Mormon 3:10
\bverse \iffalse And they did swear by the heavens, and also by the throne of God, that they would go up to battle against their enemies, and would cut them off from the face of the land. \fi
They swore by the heavens, and by God's throne, that they would go to battle against their enemies and completely destroy them.
%Mormon 3:11
\bverse \iffalse And it came to pass that I, Mormon, did utterly refuse from this time forth to be a commander and a leader of this people, because of their wickedness and abomination. \fi
When this happened, I, Mormon, refused to be a commander and leader of this people, because they were wicked.
%Mormon 3:12
\bverse \iffalse Behold, I had led them, notwithstanding their wickedness I had led them many times to battle, and had loved them, according to the love of God which was in me, with all my heart; and my soul had been poured out in prayer unto my God all the day long for them; nevertheless, it was without faith, because of the hardness of their hearts. \fi
I had led them in battle many times before, even though they were very wicked, and I had loved them with all of the love of God that I had in me, with all of my heart. I had poured out my soul in prayer to my God all day long for them. But I did not have any faith when I prayed, because their hearts were so hard.
%Mormon 3:13
\bverse \iffalse And thrice have I delivered them out of the hands of their enemies, and they have repented not of their sins. \fi
I have saved them from their enemies three times, but they have not repented of their sins.
%Mormon 3:14
\bverse \iffalse And when they had sworn by all that had been forbidden them by our Lord and Savior Jesus Christ, that they would go up unto their enemies to battle, and avenge themselves of the blood of their brethren, behold the voice of the Lord came unto me saying: \fi
And when they swore by all of the things that our Lord and Savior Jesus Christ commanded them not to swear by, and said that they would go to battle their enemies to take revenge, the voice of the Lord came to me and said:
%Mormon 3:15
\bverse \iffalse Vengeance is mine, and I will repay; and because this people repented not after I had delivered them, behold, they shall be cut off from the face of the earth. \fi
``I am the one who punishes; I will pay people back. These people did not repent after I saved them. Because of this, they will be completely destroyed.''
%Mormon 3:16
\bverse \iffalse And it came to pass that I utterly refused to go up against mine enemies; and I did even as the Lord had commanded me; and I did stand as an idle witness to manifest unto the world the things which I saw and heard, according to the manifestations of the Spirit which had testified of things to come. \fi
I completely refused to go and fight my enemies, and I did what the Lord commanded me to do. I did not do anything but watch, so that I could tell the world about the things I saw and heard, according to the things that the Spirit had testified would happen.
%Mormon 3:17
\bverse \iffalse Therefore I write unto you, Gentiles, and also unto you, house of Israel, when the work shall commence, that ye shall be about to prepare to return to the land of your inheritance; \fi
This is why I am writing to you non-Israelites, and also you people of Israel---to tell you that when the work has begun, you will be about to prepare to go back to the lands you have inherited.
%Mormon 3:18
\bverse \iffalse Yea, behold, I write unto all the ends of the earth; yea, unto you, twelve tribes of Israel, who shall be judged according to your works by the twelve whom Jesus chose to be his disciples in the land of Jerusalem. \fi
I write to all people, all over the world. I write to you, twelve tribes of Israel. You will be judged for the things you have done by the twelve whom Jesus chose to be his disciples in the land of Jerusalem.
%Mormon 3:19
\bverse \iffalse And I write also unto the remnant of this people, who shall also be judged by the twelve whom Jesus chose in this land; and they shall be judged by the other twelve whom Jesus chose in the land of Jerusalem. \fi
I also write to the survivors of this people; they will be judged by the twelve whom Jesus chose in this land. And they will be judged by the other twelve whom Jesus chose when he was in the land of Jerusalem.
%Mormon 3:20
\bverse \iffalse And these things doth the Spirit manifest unto me; therefore I write unto you all. And for this cause I write unto you, that ye may know that ye must all stand before the judgment-seat of Christ, yea, every soul who belongs to the whole human family of Adam; and ye must stand to be judged of your works, whether they be good or evil; \fi
These are the things that the Spirit has shown to me. This is why I am writing to all of you: so that you can know that all of you will stand before Christ's judgment-seat---every single person who belongs to the family of Eve and Adam. You will be judged for the things you have done, whether they are good or evil.
%Mormon 3:21
\bverse \iffalse And also that ye may believe the gospel of Jesus Christ, which ye shall have among you; and also that the Jews, the covenant people of the Lord, shall have other witness besides him whom they saw and heard, that Jesus, whom they slew, was the very Christ and the very God. \fi
I am also writing so that you will believe the gospel of Jesus Christ, which you will have among you, and so that the Jews, the people with whom the Lord made a covenant, will have more witnesses than Jesus, the one they saw and heard themselves, and killed. He was the true Christ and the true God.
%Mormon 3:22
\bverse \iffalse And I would that I could persuade all ye ends of the earth to repent and prepare to stand before the judgment-seat of Christ. \fi
I wish I could convince all of you, all over the world, to repent and prepare to stand before Christ's judgment-seat.
%Chapter 4
\bchapter
%Mormon 4:1
\bverse \iffalse And now it came to pass that in the three hundred and sixty and third year the Nephites did go up with their armies to battle against the Lamanites, out of the land Desolation. \fi
And during the three-hundred and sixty-third year, the Nephites went out of the land of the Ruins with their armies to battle against the Lamanites.
%Mormon 4:2
\bverse \iffalse And it came to pass that the armies of the Nephites were driven back again to the land of Desolation. And while they were yet weary, a fresh army of the Lamanites did come upon them; and they had a sore battle, insomuch that the Lamanites did take possession of the city Desolation, and did slay many of the Nephites, and did take many prisoners. \fi
And the armies of the Nephites were driven back into the land of The Ruins. And while they were still tired, a fresh army of Lamanites came to attack them. They had a horrible battle, and the Lamanites took control of the city of The Ruins. They killed many of the Nephites, and took many prisoners.
%Mormon 4:3
\bverse \iffalse And the remainder did flee and join the inhabitants of the city Teancum. Now the city Teancum lay in the borders by the seashore; and it was also near the city Desolation. \fi
The rest of the Nephites ran away and joined the people in the city Teancum. The city Teancum was on the borders, near the seashore. It was also near the city of The Ruins.
%Mormon 4:4
\bverse \iffalse And it was because the armies of the Nephites went up unto the Lamanites that they began to be smitten; for were it not for that, the Lamanites could have had no power over them. \fi
The Nephites started to lose in battle because they had gone to attack the Lamanites. If they had not done this, the Lamanites could not have conquered them.
%Mormon 4:5
\bverse \iffalse But, behold, the judgments of God will overtake the wicked; and it is by the wicked that the wicked are punished; for it is the wicked that stir up the hearts of the children of men unto bloodshed. \fi
But the punishment of God will catch up with the wicked. Wicked people are punished by the wicked, because it is wicked people who stir up the hearts of other to get them to kill each other.
%Mormon 4:6
\bverse \iffalse And it came to pass that the Lamanites did make preparations to come against the city Teancum. \fi
And the Lamanites began to prepare to attack the city Teancum.
%Mormon 4:7
\bverse \iffalse And it came to pass in the three hundred and sixty and fourth year the Lamanites did come against the city Teancum, that they might take possession of the city Teancum also. \fi
And in the three hundred and sixty-fourth year the Lamanites came to attack the city Teancum, and take control of it.
%Mormon 4:8
\bverse \iffalse And it came to pass that they were repulsed and driven back by the Nephites. And when the Nephites saw that they had driven the Lamanites they did again boast of their own strength; and they went forth in their own might, and took possession again of the city Desolation. \fi
But the Nephites drove them back. And when the Nephites saw that they had driven the Lamanites back, they again boasted about how strong they were.
%Mormon 4:9
\bverse \iffalse And now all these things had been done, and there had been thousands slain on both sides, both the Nephites and the Lamanites. \fi
All of this happened, and thousands of people had been killed on both sides, both the Nephites and the Lamanites.
%Mormon 4:10
\bverse \iffalse And it came to pass that the three hundred and sixty and sixth year had passed away, and the Lamanites came again upon the Nephites to battle; and yet the Nephites repented not of the evil they had done, but persisted in their wickedness continually. \fi
And the three hundred and sixty-sixth year had ended, and the Lamanites came again to attack the Nephites. And the Nephites still would not repent of the evil things they had done; they continued to live very wickedly.
%Mormon 4:11
\bverse \iffalse And it is impossible for the tongue to describe, or for man to write a perfect description of the horrible scene of the blood and carnage which was among the people, both of the Nephites and of the Lamanites; and every heart was hardened, so that they delighted in the shedding of blood continually. \fi
It is impossible for anyone to describe the horrible scene of blood and death that I saw among the people, both the Nephites and the Lamanites. Every person's heart was hardened, and they loved to kill and murder.
%Mormon 4:12
\bverse \iffalse And there never had been so great wickedness among all the children of Lehi, nor even among all the house of Israel, according to the words of the Lord, as was among this people. \fi
There had never been such wicked people among all the descendants of Sariah and Lehi, or even among all the descendants of Israel. This is what the Lord said.
%Mormon 4:13
\bverse \iffalse And it came to pass that the Lamanites did take possession of the city Desolation, and this because their number did exceed the number of the Nephites. \fi
And the Lamanites took over the city of The Ruins, because there were more of them than there were Nephites.
%Mormon 4:14
\bverse \iffalse And they did also march forward against the city Teancum, and did drive the inhabitants forth out of her, and did take many prisoners both women and children, and did offer them up as sacrifices unto their idol gods. \fi
They also marched to attack the city Teancum, and they drove out the people who lived there. They took many prisoners, including women and children, and offered them as sacrifices to their idol gods.
%Mormon 4:15
\bverse \iffalse And it came to pass that in the three hundred and sixty and seventh year, the Nephites being angry because the Lamanites had sacrificed their women and their children, that they did go against the Lamanites with exceedingly great anger, insomuch that they did beat again the Lamanites, and drive them out of their lands. \fi
And in the three hundred and sixty-seventh year, the Nephites were very angry because the Lamanites had sacrificed Nephite women and children to their gods. They went to attack the Lamanites in their anger, and they beat the Lamanites and drove them out of their lands.
%Mormon 4:16
\bverse \iffalse And the Lamanites did not come again against the Nephites until the three hundred and seventy and fifth year. \fi
And the Lamanites did not come to fight against the Nephites until the three hundred and seventy-fifth year.
%Mormon 4:17
\bverse \iffalse And in this year they did come down against the Nephites with all their powers; and they were not numbered because of the greatness of their number. \fi
But this year they came to fight the Nephites with all of their power. There were so many Lamanites that no one counted how many there were.
%Mormon 4:18
\bverse \iffalse And from this time forth did the Nephites gain no power over the Lamanites, but began to be swept off by them even as a dew before the sun. \fi
And from that time onward, the Nephites could not win against the Lamanites. The Nephites started to be swept off the land like dew drying up in the sunlight.
%Mormon 4:19
\bverse \iffalse And it came to pass that the Lamanites did come down against the city Desolation; and there was an exceedingly sore battle fought in the land Desolation, in the which they did beat the Nephites. \fi
And the Lamanites came to attack the city of The Ruins. There was a horrible battle in the land of The Ruins, and the Lamanites beat the Nephites.
%Mormon 4:20
\bverse \iffalse And they fled again from before them, and they came to the city Boaz; and there they did stand against the Lamanites with exceeding boldness, insomuch that the Lamanites did not beat them until they had come again the second time. \fi
The Nephites ran from them, and came to the city Boaz. They fought so boldly against the Lamanites there, that the Lamanites did not beat them until they had come to fight the Nephites the second time.
%Mormon 4:21
\bverse \iffalse And when they had come the second time, the Nephites were driven and slaughtered with an exceedingly great slaughter; their women and their children were again sacrificed unto idols. \fi
And when the Lamanites came the second time, the Nephites were driven out and almost completely killed. And the Lamanites again sacrificed Nephite women and children to idols.
%Mormon 4:22
\bverse \iffalse And it came to pass that the Nephites did again flee from before them, taking all the inhabitants with them, both in towns and villages. \fi
And the Nephites ran from them, and took all the people in all the towns and villages with them along the way.
%Mormon 4:23
\bverse \iffalse And now I, Mormon, seeing that the Lamanites were about to overthrow the land, therefore I did go to the hill Shim, and did take up all the records which Ammaron had hid up unto the Lord. \fi
And I, Mormon, saw that the Lamanites were about to overthrow the land. I went to the hill Shim with all of the records that Ammaron had hidden and asked the Lord to protect.
%Chapter 5
\bchapter
%Mormon 5:1
\bverse \iffalse And it came to pass that I did go forth among the Nephites, and did repent of the oath which I had made that I would no more assist them; and they gave me command again of their armies, for they looked upon me as though I could deliver them from their afflictions. \fi
And I went among the Nephites. I changed my mind, and took back the oath I had made to not help them anymore. They made me the commander of their armies, because they thought that I could save them from destruction.
%Mormon 5:2
\bverse \iffalse But behold, I was without hope, for I knew the judgments of the Lord which should come upon them; for they repented not of their iniquities, but did struggle for their lives without calling upon that Being who created them. \fi
But I did not have any hope for them. I knew that the Lord was punishing them, because they would not repent of their sins. They struggled to save their lives, but they refused to pray to the One who created them for help.
%Mormon 5:3
\bverse \iffalse And it came to pass that the Lamanites did come against us as we had fled to the city of Jordan; but behold, they were driven back that they did not take the city at that time. \fi
And the Lamanites came to fight us as we ran away to the city of Jordan. But we beat them back, and they did not capture the city at that time.
%Mormon 5:4
\bverse \iffalse And it came to pass that they came against us again, and we did maintain the city. And there were also other cities which were maintained by Nephites, which strongholds did cut them off that they could not get into the country which lay before us, to destroy the inhabitants of our land. \fi
They came to attack us again, but we kept control of the city. There were also other cities that the Nephites had control of. The Nephites in those places stopped the Lamanites from coming into the country to destroy the people there.
%Mormon 5:5
\bverse \iffalse And it came to pass that whatsoever lands we had passed by, and the inhabitants thereof were not gathered in, were destroyed by the Lamanites, and their towns, and villages, and cities were burned with fire; and thus three hundred and seventy and nine years passed away. \fi
And all of the lands, towns, villages, and cities that we passed by were burned and destroyed by the Lamanites. If the people living in those places did not come with us when we passed by, they were also destroyed. And three hundred and seventy-nine years had gone by.
%Mormon 5:6
\bverse \iffalse And it came to pass that in the three hundred and eightieth year the Lamanites did come again against us to battle, and we did stand against them boldly; but it was all in vain, for so great were their numbers that they did tread the people of the Nephites under their feet. \fi
And during the three hundred and eightieth year the Lamanites again came to battle against us. We fought against them boldly, but it was no use: there were so many of them that they trampled the Nephites under their feet.
%Mormon 5:7
\bverse \iffalse And it came to pass that we did again take to flight, and those whose flight was swifter than the Lamanites' did escape, and those whose flight did not exceed the Lamanites' were swept down and destroyed. \fi
We again tried to escape. Anyone who could run faster than the Lamanites escaped, but anyone who could not was killed.
%Mormon 5:8
\bverse \iffalse And now behold, I, Mormon, do not desire to harrow up the souls of men in casting before them such an awful scene of blood and carnage as was laid before mine eyes; but I, knowing that these things must surely be made known, and that all things which are hid must be revealed upon the house-tops-- \fi
I, Mormon, do not want to fill anyone with horror by describing the horrible scene of blood, war, and death that I saw. But I know that these things will be revealed, and that anything that is hidden will one day be shouted from housetops.
%Mormon 5:9
\bverse \iffalse And also that a knowledge of these things must come unto the remnant of these people, and also unto the Gentiles, who the Lord hath said should scatter this people, and this people should be counted as naught among them--therefore I write a small abridgment, daring not to give a full account of the things which I have seen, because of the commandment which I have received, and also that ye might not have too great sorrow because of the wickedness of this people. \fi
I know that the record of these things must come to the descendants of these people, and to the non-Israelites---the ones whom the Lord has said would scatter this people, and who would consider this people worthless. So I am only writing a small part; I do not dare to tell the full story of the things I have seen, because of the commandment the Lord gave me. I also do not want to fill anyone with overwhelming sadness because of the sins of this people.
%Mormon 5:10
\bverse \iffalse And now behold, this I speak unto their seed, and also to the Gentiles who have care for the house of Israel, that realize and know from whence their blessings come. \fi
Now I speak to their descendants, and to the non-Israelites who are supposed to take care of the descendants of Israel, and who understand where their blessings come from.
%Mormon 5:11
\bverse \iffalse For I know that such will sorrow for the calamity of the house of Israel; yea, they will sorrow for the destruction of this people; they will sorrow that this people had not repented that they might have been clasped in the arms of Jesus. \fi
I know that these people will be filled with sadness because of the downfall of the descendants of Israel. They will be filled with sadness when they find out about the destruction of this people, because this people would not repent when Jesus was willing to hold them in his arms.
%Mormon 5:12
\bverse \iffalse Now these things are written unto the remnant of the house of Jacob; and they are written after this manner, because it is known of God that wickedness will not bring them forth unto them; and they are to be hid up unto the Lord that they may come forth in his own due time. \fi
I am writing these things for the descendants of the people of Israel. I am writing them in this way, because God knows that no one that is wicked can find them and show them to the descendants of the Israelites. These records will be hidden and protected by the Lord until he decides that it is right for them to be revealed.
%Mormon 5:13
\bverse \iffalse And this is the commandment which I have received; and behold, they shall come forth according to the commandment of the Lord, when he shall see fit, in his wisdom. \fi
This is the commandment that I have been given. They will come out and be shown according to the Lord's commandments, when he decides with his wisdom that the time is right.
%Mormon 5:14
\bverse \iffalse And behold, they shall go unto the unbelieving of the Jews; and for this intent shall they go--that they may be persuaded that Jesus is the Christ, the Son of the living God; that the Father may bring about, through his most Beloved, his great and eternal purpose, in restoring the Jews, or all the house of Israel, to the land of their inheritance, which the Lord their God hath given them, unto the fulfilling of his covenant; \fi
These records will go to the Jews who do not believe. And this is why: To convince the Jews that Jesus is the Christ, the Son of the living God. Then, through the most Beloved Son of the Father, the Father can keep his great and eternal promise to give the Jews and all the descendants of Israel the land that they have inherited. The Lord their God has given this land to them, and he will keep his promise.
%Mormon 5:15
\bverse \iffalse And also that the seed of this people may more fully believe his gospel, which shall go forth unto them from the Gentiles; for this people shall be scattered, and shall become a dark, a filthy, and a loathsome people, beyond the description of that which ever hath been amongst us, yea, even that which hath been among the Lamanites, and this because of their unbelief and idolatry. \fi
These records will also come so that the descendants of this people will believe the gospel that the non-Israelites will bring to them. This people will be scattered and become a dark, filthy, and unwanted people, worse than anyone who has ever been among us or even among the Lamanites, because they do not believe and they worship idols.
%Mormon 5:16
\bverse \iffalse For behold, the Spirit of the Lord hath already ceased to strive with their fathers; and they are without Christ and God in the world; and they are driven about as chaff before the wind. \fi
The Spirit of the Lord has already left their ancestors. They live without Christ and without God in this world. And they are driven around like chaff in the wind.
%Mormon 5:17
\bverse \iffalse They were once a delightsome people, and they had Christ for their shepherd; yea, they were led even by God the Father. \fi
They were once a delightful people, and Christ was their shepherd. They were even led by God the Father.
%Mormon 5:18
\bverse \iffalse But now, behold, they are led about by Satan, even as chaff is driven before the wind, or as a vessel is tossed about upon the waves, without sail or anchor, or without anything wherewith to steer her; and even as she is, so are they. \fi
But now they are led by Satan, just like chaff that is driven by the wind, or like a ship that is tossed on the sea, with no sail and no anchor, and nothing to steer it. They are just like that ship.
%Mormon 5:19
\bverse \iffalse And behold, the Lord hath reserved their blessings, which they might have received in the land, for the Gentiles who shall possess the land. \fi
The Lord has saved the blessings that he would have given them, and he will give them to the non-Israelites who come into this land.
%Mormon 5:20
\bverse \iffalse But behold, it shall come to pass that they shall be driven and scattered by the Gentiles; and after they have been driven and scattered by the Gentiles, behold, then will the Lord remember the covenant which he made unto Abraham and unto all the house of Israel. \fi
And the descendants of this people will be driven and scattered by the non-Israelites. And after they have been driven and scattered by the non-Israelites, the Lord will remember the covenant that he made to Abraham and all the people of Israel.
%Mormon 5:21
\bverse \iffalse And also the Lord will remember the prayers of the righteous, which have been put up unto him for them. \fi
The Lord will remember the prayers that righteous people pray to him for them.
%Mormon 5:22
\bverse \iffalse And then, O ye Gentiles, how can ye stand before the power of God, except ye shall repent and turn from your evil ways? \fi
But, you non-Israelites, how are you going to stand up to the power of God if you do not repent and stop living sinful lives?
%Mormon 5:23
\bverse \iffalse Know ye not that ye are in the hands of God? Know ye not that he hath all power, and at his great command the earth shall be rolled together as a scroll? \fi
Don't you realize that you are in God's hands? Don't you know that God has all power? When he gives the command, the whole earth will be wrapped together like a scroll.
%Mormon 5:24
\bverse \iffalse Therefore, repent ye, and humble yourselves before him, lest he shall come out in justice against you--lest a remnant of the seed of Jacob shall go forth among you as a lion, and tear you in pieces, and there is none to deliver. \fi
So repent, and become humble before God. If you do not he will come to punish you, and the descendants of Israel will go among you like a lion and tear you in pieces. No one will be able to save you.
%Chapter 6
\bchapter
%Mormon 6:1
\bverse \iffalse And now I finish my record concerning the destruction of my people, the Nephites. And it came to pass that we did march forth before the Lamanites. \fi
Now I finish the record about the destruction of my people, the Nephites. We marched out in front of the Lamanites.
%Mormon 6:2
\bverse \iffalse And I, Mormon, wrote an epistle unto the king of the Lamanites, and desired of him that he would grant unto us that we might gather together our people unto the land of Cumorah, by a hill which was called Cumorah, and there we could give them battle. \fi
And I, Mormon, wrote a letter to the king of the Lamanites. I asked him if he would allow us gather our people together in the land of Cumorah, by a hill called Cumorah, and battle him there.
%Mormon 6:3
\bverse \iffalse And it came to pass that the king of the Lamanites did grant unto me the thing which I desired. \fi
And the king of the Lamanites allowed me to do this.
%Mormon 6:4
\bverse \iffalse And it came to pass that we did march forth to the land of Cumorah, and we did pitch our tents around about the hill Cumorah; and it was in a land of many waters, rivers, and fountains; and here we had hope to gain advantage over the Lamanites. \fi
So we marched to the land of Cumorah, and we pitched our tents around the hill Cumorah, in a land of many waters, rivers, and springs. We hoped that here we would have a chance to beat the Lamanites.
%Mormon 6:5
\bverse \iffalse And when three hundred and eighty and four years had passed away, we had gathered in all the remainder of our people unto the land of Cumorah. \fi
And when the three hundred and eighty-fourth year had ended, we had gathered together the rest of our people to the land of Cumorah.
%Mormon 6:6
\bverse \iffalse And it came to pass that when we had gathered in all our people in one to the land of Cumorah, behold I, Mormon, began to be old; and knowing it to be the last struggle of my people, and having been commanded of the Lord that I should not suffer the records which had been handed down by our fathers, which were sacred, to fall into the hands of the Lamanites, (for the Lamanites would destroy them) therefore I made this record out of the plates of Nephi, and hid up in the hill Cumorah all the records which had been entrusted to me by the hand of the Lord, save it were these few plates which I gave unto my son Moroni. \fi
When we had gathered all of our people together to the land of Cumorah, I, Mormon, began to be old. I knew that this would be my people's last battle, and the Lord had commanded me to not let the Lamanites gets the sacred records that had been handed down by our ancestors, because the Lamanites would destroy them. So I made this record our of the plates of Nephi, and I hid all the records that the Lord had trusted me with in the hill Cumorah, except for these few plates that I gave to my son, Moroni.
%Mormon 6:7
\bverse \iffalse And it came to pass that my people, with their wives and their children, did now behold the armies of the Lamanites marching towards them; and with that awful fear of death which fills the breasts of all the wicked, did they await to receive them. \fi
And my people---the mothers and fathers, and their children---saw the armies of the Lamanites marching toward them. My people waited for them to come with that horrible fear of death that fills the hearts of all wicked people.
%Mormon 6:8
\bverse \iffalse And it came to pass that they came to battle against us, and every soul was filled with terror because of the greatness of their numbers. \fi
And they came to battle against us. Every one of us was filled with terror because there were so many of them.
%Mormon 6:9
\bverse \iffalse And it came to pass that they did fall upon my people with the sword, and with the bow, and with the arrow, and with the ax, and with all manner of weapons of war. \fi
They attacked my people with swords, bows, arrows, axes, and every other kind of weapon.
%Mormon 6:10
\bverse \iffalse And it came to pass that my men were hewn down, yea, even my ten thousand who were with me, and I fell wounded in the midst; and they passed by me that they did not put an end to my life. \fi
All of the ten thousand soldiers that I led were killed, and I fell wounded on the battlefield. And when they passed by me they did not kill me.
%Mormon 6:11
\bverse \iffalse And when they had gone through and hewn down all my people save it were twenty and four of us, (among whom was my son Moroni) and we having survived the dead of our people, did behold on the morrow, when the Lamanites had returned unto their camps, from the top of the hill Cumorah, the ten thousand of my people who were hewn down, being led in the front by me. \fi
They went through and killed all of my people except for twenty-four of us. My son Moroni was one of the survivors. The next day, when the Lamanites had gone back to their camps, we were on top of the hill Cumorah. We saw my ten thousand soldiers who had been killed.
%Mormon 6:12
\bverse \iffalse And we also beheld the ten thousand of my people who were led by my son Moroni. \fi
We also saw the ten thousand Nephites who were led by my son, Moroni.
%Mormon 6:13
\bverse \iffalse And behold, the ten thousand of Gidgiddonah had fallen, and he also in the midst. \fi
The ten thousand soldiers of Gidgiddonah, and Gidgiddonah himself, had also died.
%Mormon 6:14
\bverse \iffalse And Lamah had fallen with his ten thousand; and Gilgal had fallen with his ten thousand; and Limhah had fallen with his ten thousand; and Jeneum had fallen with his ten thousand; and Cumenihah, and Moronihah, and Antionum, and Shiblom, and Shem, and Josh, had fallen with their ten thousand each. \fi
We saw that Lamah had been killed with his ten thousand; and Gilgal with his ten thousand; and Limhah with his ten thousand; and Jeneum with his ten thousand; and Cumenihah with his ten thousand; and Moronihah, and Antionum, and Shiblom, and Shem, and Josh---they had all died with their ten thousand.
%Mormon 6:15
\bverse \iffalse And it came to pass that there were ten more who did fall by the sword, with their ten thousand each; yea, even all my people, save it were those twenty and four who were with me, and also a few who had escaped into the south countries, and a few who had deserted over unto the Lamanites, had fallen; and their flesh, and bones, and blood lay upon the face of the earth, being left by the hands of those who slew them to molder upon the land, and to crumble and to return to their mother earth. \fi
There were ten others who died with their ten thousand soldiers each. All of my people had been killed, except for the twenty-four people who were with me, a few who had escaped into the countries to the south, and a few who had abandoned us and joined the Lamanites. Their bodies, and bones, and blood, were left out on the ground by their enemies to rot and crumble, and return back into the earth.
%Mormon 6:16
\bverse \iffalse And my soul was rent with anguish, because of the slain of my people, and I cried: \fi
My soul was torn with sadness, because my people had died. I cried out:
%Mormon 6:17
\bverse \iffalse O ye fair ones, how could ye have departed from the ways of the Lord! O ye fair ones, how could ye have rejected that Jesus, who stood with open arms to receive you! \fi
``You dear, beautiful people! Why did you leave the ways of the Lord? O you dear, beautiful people! Why did you reject Jesus when he was standing with open arms to welcome you?
%Mormon 6:18
\bverse \iffalse Behold, if ye had not done this, ye would not have fallen.  But behold, ye are fallen, and I mourn your loss. \fi
If you had not done this, you would not have died. But now you have died, and it fills me with grief.
%Mormon 6:19
\bverse \iffalse O ye fair sons and daughters, ye fathers and mothers, ye husbands and wives, ye fair ones, how is it that ye could have fallen! \fi
You beautiful daughters and sons, you mothers and fathers, you wives and husbands---my dear people, why have you fallen?
%Mormon 6:20
\bverse \iffalse But behold, ye are gone, and my sorrows cannot bring your return. \fi
But now you are gone, and my sadness will not bring you back.
%Mormon 6:21
\bverse \iffalse And the day soon cometh that your mortal must put on immortality, and these bodies which are now moldering in corruption must soon become incorruptible bodies; and then ye must stand before the judgment-seat of Christ to be judged according to your works and if it so be that ye are righteous, then are ye blessed with your fathers who have gone before you. \fi
And the day is soon coming that your bodies will be changed so that they no longer die. These bodies that are now rotting must soon become unchangeable. Then you will stand before Christ's judgment-seat to be judged for the things you have done. And if you are righteous, you will be blessed with your ancestors that have died before you.
%Mormon 6:22
\bverse \iffalse O that ye had repented before this great destruction had come upon you. But behold, ye are gone, and the Father, yea, the Eternal Father of heaven, knoweth your state; and he doeth with you according to his justice and mercy. \fi
O how I wish that you had repented before you were destroyed. But now you are gone, and the Father---the Eternal Father of heaven---knows where you are. And he will deal with you according to his justice and mercy.
%Chapter 7
\bchapter
%Mormon 7:1
\bverse \iffalse And now, behold, I would speak somewhat unto the remnant of this people who are spared, if it so be that God may give unto them my words, that they may know of the things of their fathers; yea, I speak unto you, ye remnant of the house of Israel; and these are the words which I speak: \fi
Now I want to speak to you about the descendants of these people who are spared so they can know about their ancestors, if God decides to give what I have written to them. I speak to you, descendants of Israel, and here is what I say:
%Mormon 7:2
\bverse \iffalse Know ye that ye are of the house of Israel. \fi
Understand that you are descendants of Israel.
%Mormon 7:3
\bverse \iffalse Know ye that ye must come unto repentance, or ye cannot be saved. \fi
Understand that you must repent, or you cannot be saved.
%Mormon 7:4
\bverse \iffalse Know ye that ye must lay down your weapons of war, and delight no more in the shedding of blood, and take them not again, save it be that God shall command you. \fi
Understand that you must put down your weapons, and stop thinking it is good to kill others. Do not pick up your weapons again, unless God commands you.
%Mormon 7:5
\bverse \iffalse Know ye that ye must come to the knowledge of your fathers, and repent of all your sins and iniquities, and believe in Jesus Christ, that he is the Son of God, and that he was slain by the Jews, and by the power of the Father he hath risen again, whereby he hath gained the victory over the grave; and also in him is the sting of death swallowed up. \fi
Understand that you must learn about your ancestors, and repent of all the sins and evil things you have done, and believe in Jesus Christ. Believe that he is the Son of God, and that he was killed by the Jews, and that he has risen from the dead through the power of the Father. Believe that he has conquered death, and that because of him the pain of death is taken away.
%Mormon 7:6
\bverse \iffalse And he bringeth to pass the resurrection of the dead, whereby man must be raised to stand before his judgment-seat. \fi
Because of him, everyone will be resurrected. And because of the resurrection, everyone must be raised up and stand in front of his judgment-seat.
%Mormon 7:7
\bverse \iffalse And he hath brought to pass the redemption of the world, whereby he that is found guiltless before him at the judgment day hath it given unto him to dwell in the presence of God in his kingdom, to sing ceaseless praises with the choirs above, unto the Father, and unto the Son, and unto the Holy Ghost, which are one God, in a state of happiness which hath no end. \fi
Because of him the world is redeemed, and the person that is not guilty on the day of judgment will be allowed to live with God in God's kingdom in neverending happiness, and sing neverending songs of praise with the choirs in heaven to the Father, and to the Son, and to the Holy Ghost. And they are one God.
%Mormon 7:8
\bverse \iffalse Therefore repent, and be baptized in the name of Jesus, and lay hold upon the gospel of Christ, which shall be set before you, not only in this record but also in the record which shall come unto the Gentiles from the Jews, which record shall come from the Gentiles unto you. \fi
So repent, and be baptized in the name of Jesus, and take hold of the gospel of Christ. The gospel will be set in front of you---not only in this record, but also in the record that comes to the non-Israelites from the Jews. And that record will come from the non-Israelites to you.
%Mormon 7:9
\bverse \iffalse For behold, this is written for the intent that ye may believe that; and if ye believe that ye will believe this also; and if ye believe this ye will know concerning your fathers, and also the marvelous works which were wrought by the power of God among them. \fi
This record is written so that you will believe that record. And if you believe that record, you will believe this record, too. And if you believe this record you will know about your ancestors, and the wonderful things that were done among them by God's power.
%Mormon 7:10
\bverse \iffalse And ye will also know that ye are a remnant of the seed of Jacob; therefore ye are numbered among the people of the first covenant; and if it so be that ye believe in Christ, and are baptized, first with water, then with fire and with the Holy Ghost, following the example of our Savior, according to that which he hath commanded us, it shall be well with you in the day of judgment. Amen. \fi
You will also know that you are descendants of Israel, and that you are counted as the people that God first made a covenant with. And if you believe in Christ, and you follow the example of our Savior and are baptized---first with water, and then with the Holy Ghost---as he commanded us, you will be happy on the day of judgment. Amen.
%Chapter 8
\bchapter
%Mormon 8:1
\bverse \iffalse Behold I, Moroni, do finish the record of my father, Mormon.  Behold, I have but few things to write, which things I have been commanded by my father. \fi
Now I, Moroni, am finishing the record made by my father, Mormon. I have only a few things to write, and they are things that my father commanded me to write.
%Mormon 8:2
\bverse \iffalse And now it came to pass that after the great and tremendous battle at Cumorah, behold, the Nephites who had escaped into the country southward were hunted by the Lamanites, until they were all destroyed. \fi
After the great and terrible battle at Cumorah, the Nephites who had escaped into the land in the south were hunted by the Lamanites until they were all killed.
%Mormon 8:3
\bverse \iffalse And my father also was killed by them, and I even remain alone to write the sad tale of the destruction of my people. But behold, they are gone, and I fulfil the commandment of my father. And whether they will slay me, I know not. \fi
My father was also killed by them. Now I am by myself, writing the sad story of my people's destruction. They are all gone, and I am doing what my father commanded me to do. I do not know whether they will find me and kill me, too.
%Mormon 8:4
\bverse \iffalse Therefore I will write and hide up the records in the earth; and whither I go it mattereth not. \fi
This is why I am writing this record. After this I will bury the records in the ground. And then it does not matter where I go.
%Mormon 8:5
\bverse \iffalse Behold, my father hath made this record, and he hath written the intent thereof. And behold, I would write it also if I had room upon the plates, but I have not; and ore I have none, for I am alone. My father hath been slain in battle, and all my kinsfolk, and I have not friends nor whither to go; and how long the Lord will suffer that I may live I know not. \fi
My father made this record, and he explained why he was writing it. I would also write about it if I had room on the plates, but there is no more room. I don't have any ore to make more plates, because I am alone. My father was killed in battle, and so were all of my relatives and friends. I have no friends now, and nowhere to go. I do not know how long the Lord will keep me alive. 
%Mormon 8:6
\bverse \iffalse Behold, four hundred years have passed away since the coming of our Lord and Savior. \fi
Four hundred years have gone by since our Lord and Savior came. 
%Mormon 8:7
\bverse \iffalse And behold, the Lamanites have hunted my people, the Nephites, down from city to city and from place to place, even until they are no more; and great has been their fall; yea, great and marvelous is the destruction of my people, the Nephites. \fi
And the Lamanites have hunted my people, the Nephites, from city to city and from place to place, and now they are all dead. The fall of my people has been great. It is hard to believe that my people have been destroyed.
%Mormon 8:8
\bverse \iffalse And behold, it is the hand of the Lord which hath done it.  And behold also, the Lamanites are at war one with another; and the whole face of this land is one continual round of murder and bloodshed; and no one knoweth the end of the war. \fi
It is the Lord's hand that has done this. The Lamanites are also at war with each other. All over this entire land there is neverending murder and death. No one knows when the wars will end.
%Mormon 8:9
\bverse \iffalse And now, behold, I say no more concerning them, for there are none save it be the Lamanites and robbers that do exist upon the face of the land. \fi
But now I will not say anything else about them. No one is left in this land but Lamanites and robbers.
%Mormon 8:10
\bverse \iffalse And there are none that do know the true God save it be the disciples of Jesus, who did tarry in the land until the wickedness of the people was so great that the Lord would not suffer them to remain with the people; and whether they be upon the face of the land no man knoweth. \fi
No one knows the true God except for the disciples of Jesus, the ones who stayed in the land until the people became so wicked that the Lord would not let them stay with the people. And now no one knows where they are.
%Mormon 8:11
\bverse \iffalse But behold, my father and I have seen them, and they have ministered unto us. \fi
By my father and I have seen them; they have visited us and taught us.
%Mormon 8:12
\bverse \iffalse And whoso receiveth this record, and shall not condemn it because of the imperfections which are in it, the same shall know of greater things than these. Behold, I am Moroni; and were it possible, I would make all things known unto you. \fi
And anyone who reads this record, and who does not judge it because it is imperfect, will get to know about even greater things. I am Moroni, and if I could I would reveal everything to you.
%Mormon 8:13
\bverse \iffalse Behold, I make an end of speaking concerning this people. I am the son of Mormon, and my father was a descendant of Nephi. \fi
But now I stop speaking about my people. I am the son of Mormon, and my father was a descendant of Nephi.
%Mormon 8:14
\bverse \iffalse And I am the same who hideth up this record unto the Lord; the plates thereof are of no worth, because of the commandment of the Lord. For he truly saith that no one shall have them to get gain; but the record thereof is of great worth; and whoso shall bring it to light, him will the Lord bless. \fi
I am also the one who will hide this record and ask the Lord to protect it. The plates that I am writing on are not worth anything, because of the Lord's commandment. He says that no one who wants to get them for wealth and power will be able to get them. But the record on these plates is worthy very much. The person who shows it to the world will be blessed by the Lord.
%Mormon 8:15
\bverse \iffalse For none can have power to bring it to light save it be given him of God; for God wills that it shall be done with an eye single to his glory, or the welfare of the ancient and long dispersed covenant people of the Lord. \fi
No one can show it to the world unless God allows it. God wants the person who does this to do it because they only want to bring glory to God, and to help the ancient and scattered people of the Lord---the ones he has made a covenant with.
%Mormon 8:16
\bverse \iffalse And blessed be he that shall bring this thing to light; for it shall be brought out of darkness unto light, according to the word of God; yea, it shall be brought out of the earth, and it shall shine forth out of darkness, and come unto the knowledge of the people; and it shall be done by the power of God. \fi
The person that reveals this record to the world will be blessed. This record will be brought out of darkness and into the light, according to God's word. It will be brought out of the ground, and shine from the darkness, and people will know about it. All of this will be done by the power of God.
%Mormon 8:17
\bverse \iffalse And if there be faults they be the faults of a man. But behold, we know no fault; nevertheless God knoweth all things; therefore, he that condemneth, let him be aware lest he shall be in danger of hell fire. \fi
And if there are mistakes, they are mistakes made by humans. We do not think there are any mistakes in our record, but God is the one that knows all things. And the person that judges our record should be careful; they are in danger of being burned by the fires of hell.
%Mormon 8:18
\bverse \iffalse And he that saith: Show unto me, or ye shall be smitten--let him beware lest he commandeth that which is forbidden of the Lord. \fi
And the person who says ``Show me, or I will hurt you,'' should beware. They are asking for something that the Lord has forbidden.
%Mormon 8:19
\bverse \iffalse For behold, the same that judgeth rashly shall be judged rashly again; for according to his works shall his wages be; therefore, he that smiteth shall be smitten again, of the Lord. \fi
The person who judges too quickly, will also be judged quickly. They will be paid according to the things they do. And the person who strikes others will be struck back, by the Lord.
%Mormon 8:20
\bverse \iffalse Behold what the scripture says--man shall not smite, neither shall he judge; for judgment is mine, saith the Lord, and vengeance is mine also, and I will repay. \fi
Remember what the scripture says: ``People should not hit each other or judge each other. I am the one who judges, says the Lord, and the one who punishes. I will pay them back.''
%Mormon 8:21
\bverse \iffalse And he that shall breathe out wrath and strifes against the work of the Lord, and against the covenant people of the Lord who are the house of Israel, and shall say: We will destroy the work of the Lord, and the Lord will not remember his covenant which he hath made unto the house of Israel--the same is in danger to be hewn down and cast into the fire; \fi
Some people may be angry, and say threatening things and fight against the work of the Lord, and against the people of Israel---the Lord's people, whom he has made promises to. They might say: ``We will destroy the Lord's work; the Lord is not going to remember the promises that he made to the people of Israel.'' But those people are in danger of being cut down and thrown into the fire.
%Mormon 8:22
\bverse \iffalse For the eternal purposes of the Lord shall roll on, until all his promises shall be fulfilled. \fi
The Lord's work and his plans will continue, until he has kept all of his promises.
%Mormon 8:23
\bverse \iffalse Search the prophecies of Isaiah. Behold, I cannot write them. Yea, behold I say unto you, that those saints who have gone before me, who have possessed this land, shall cry, yea, even from the dust will they cry unto the Lord; and as the Lord liveth he will remember the covenant which he hath made with them. \fi
Study the prophecies of Isaiah. I cannot write them here. I tell you, the saints who have died before me and lived in this land will cry out to the Lord from the dust. And just as the Lord lives, he will remember the covenant that he made with them.
%Mormon 8:24
\bverse \iffalse And he knoweth their prayers, that they were in behalf of their brethren. And he knoweth their faith, for in his name could they remove mountains; and in his name could they cause the earth to shake; and by the power of his word did they cause prisons to tumble to the earth; yea, even the fiery furnace could not harm them, neither wild beasts nor poisonous serpents, because of the power of his word. \fi
He remembers their prayers. He remembers how they prayed for their sisters and brothers. He knows they are faithful, because they moved mountains in his name. They made the earth shake in his name. By the power of his word they made prisons tumble to the ground. Because of the power of his word burning furnaces could not hurt them; neither could wild beasts or poisonous snakes.
%Mormon 8:25
\bverse \iffalse And behold, their prayers were also in behalf of him that the Lord should suffer to bring these things forth. \fi
They also prayed for the person that the Lord will use to reveal these things to the world.
%Mormon 8:26
\bverse \iffalse And no one need say they shall not come, for they surely shall, for the Lord hath spoken it; for out of the earth shall they come, by the hand of the Lord, and none can stay it; and it shall come in a day when it shall be said that miracles are done away; and it shall come even as if one should speak from the dead. \fi
No one needs to say that these things won't happen, because they certainly will; the Lord has said that they will. These records will come out of the ground with the Lord's help, and no one can stop it. It will happen in a time when people say that miracles don't happen anymore. And these records will be like a person speaking from the dead.
%Mormon 8:27
\bverse \iffalse And it shall come in a day when the blood of saints shall cry unto the Lord, because of secret combinations and the works of darkness. \fi
It will come in a time when the blood of the saints cry out to the Lord, because of secret societies and dark and evil works.
%Mormon 8:28
\bverse \iffalse Yea, it shall come in a day when the power of God shall be denied, and churches become defiled and be lifted up in the pride of their hearts; yea, even in a day when leaders of churches and teachers shall rise in the pride of their hearts, even to the envying of them who belong to their churches. \fi
It will come in a time when people deny God's power, and churches are full of sin and pride. It will come in a day when leaders of churches and teachers have hearts  full of pride, and even envy the people who belong to their churches.
%Mormon 8:29
\bverse \iffalse Yea, it shall come in a day when there shall be heard of fires, and tempests, and vapors of smoke in foreign lands; \fi
It will come in a time when people speak of fires, and horrible storms, and smoke in foreign countries.
%Mormon 8:30
\bverse \iffalse And there shall also be heard of wars, rumors of wars, and earthquakes in divers places. \fi
People will also speak of wars, and rumors of wars, and earthquakes in different places.
%Mormon 8:31
\bverse \iffalse Yea, it shall come in a day when there shall be great pollutions upon the face of the earth; there shall be murders, and robbing, and lying, and deceivings, and whoredoms, and all manner of abominations; when there shall be many who will say, Do this, or do that, and it mattereth not, for the Lord will uphold such at the last day. But wo unto such for they are in the gall of bitterness and in the bonds of iniquity. \fi
It will come in a time when the face of the earth is covered in pollution. They will be murders and robberies. People will lie and trick one another, and commit sexual sins, and do all sorts of horrible things. There will be many people who say, ``Do what you want. It doesn't matter, because the Lord will not punish you on the last day.'' But how terrible it will be for these people! They are living in bitter poison, and in the chains of sin.
%Mormon 8:32
\bverse \iffalse Yea, it shall come in a day when there shall be churches built up that shall say: Come unto me, and for your money you shall be forgiven of your sins. \fi
It will come in a time when there are churches where people say: ``Come here, and you can be forgiven for your sins for your money.''
%Mormon 8:33
\bverse \iffalse O ye wicked and perverse and stiffnecked people, why have ye built up churches unto yourselves to get gain? Why have ye transfigured the holy word of God, that ye might bring damnation upon your souls? Behold, look ye unto the revelations of God; for behold, the time cometh at that day when all these things must be fulfilled. \fi
You wicked and evil people! Why have you built up churches for yourselves to get power and money? Why have you twisted the word of God? You are bringing punishment on your own souls! Look to the things that God has revealed, because the time is coming when all these things will come true.
%Mormon 8:34
\bverse \iffalse Behold, the Lord hath shown unto me great and marvelous things concerning that which must shortly come, at that day when these things shall come forth among you. \fi
The Lord has shown me great and wonderful things about what will happen soon after these records come among you.
%Mormon 8:35
\bverse \iffalse Behold, I speak unto you as if ye were present, and yet ye are not. But behold, Jesus Christ hath shown you unto me, and I know your doing. \fi
I am speaking to you as if you were here with me, even though you are not. But Jesus Christ has shown you to me, and I know the things you do.
%Mormon 8:36
\bverse \iffalse And I know that ye do walk in the pride of your hearts; and there are none save a few only who do not lift themselves up in the pride of their hearts, unto the wearing of very fine apparel, unto envying, and strifes, and malice, and persecutions, and all manner of iniquities; and your churches, yea, even every one, have become polluted because of the pride of your hearts. \fi
I know that you are full of proud thoughts. I know that all of you, except for a few, are full of pride and wear expensive clothes. You envy each other and fight with each other, and hate and bully other people. You do all sorts of evil things. And every single one of your churches have become polluted because of the pride in your hearts.
%Mormon 8:37
\bverse \iffalse For behold, ye do love money, and your substance, and your fine apparel, and the adorning of your churches, more than ye love the poor and the needy, the sick and the afflicted. \fi
You love money and the things you own. You love your clothing and the decorations in your churches more than you love the people who are poor and needy, and who are sick and need help.
%Mormon 8:38
\bverse \iffalse O ye pollutions, ye hypocrites, ye teachers, who sell yourselves for that which will canker, why have ye polluted the holy church of God? Why are ye ashamed to take upon you the name of Christ? Why do ye not think that greater is the value of an endless happiness than that misery which never dies--because of the praise of the world? \fi
You are full of pollution. You hypocrites and teachers! You sell yourselves for things that rot and rust. Why have you polluted God's holy church? Why are you ashamed to follow Christ and take his name on you? Why do you think that neverending misery is better than neverending happiness? Because you love the praise of the world.
%Mormon 8:39
\bverse \iffalse Why do ye adorn yourselves with that which hath no life, and yet suffer the hungry, and the needy, and the naked, and the sick and the afflicted to pass by you, and notice them not? \fi
Why do you dress in expensive clothing that has no life, and pass by the hungry, and needy, and naked, and sick, and the people who need your help, without even noticing them?
%Mormon 8:40
\bverse \iffalse Yea, why do ye build up your secret abominations to get gain, and cause that widows should mourn before the Lord, and also orphans to mourn before the Lord, and also the blood of their fathers and their husbands to cry unto the Lord from the ground, for vengeance upon your heads? \fi
Why do you build up your secret societies and evil works to get power and money? You make widows and orphans cry out to the Lord; you make the blood of their fathers and husbands cry out to the Lord from the ground, asking him to punish you.
%Mormon 8:41
\bverse \iffalse Behold, the sword of vengeance hangeth over you; and the time soon cometh that he avengeth the blood of the saints upon you, for he will not suffer their cries any longer. \fi
The sword of revenge is hanging over you. He will not wait forever while they cry out to him. Soon he will take revenge on you for the blood of the saints.
%Chapter 9
\bchapter
%Mormon 9:1
\bverse \iffalse And now, I speak also concerning those who do not believe in Christ. \fi
And now I will speak to people who do not believe in Christ.
%Mormon 9:2
\bverse \iffalse Behold, will ye believe in the day of your visitation--behold, when the Lord shall come, yea, even that great day when the earth shall be rolled together as a scroll, and the elements shall melt with fervent heat, yea, in that great day when ye shall be brought to stand before the Lamb of God--then will ye say that there is no God? \fi
Will you believe on the day when the Lord comes to judge you---the day when the earth will be rolled together like a scroll, and the elements melt with burning heat? That is the day when you will be brought to stand before God's Lamb. Then will you say that there is no God?
%Mormon 9:3
\bverse \iffalse Then will ye longer deny the Christ, or can ye behold the Lamb of God? Do ye suppose that ye shall dwell with him under a consciousness of your guilt? Do ye suppose that ye could be happy to dwell with that holy Being, when your souls are racked with a consciousness of guilt that ye have ever abused his laws? \fi
Then you will not reject Christ. You will see the Lamb of God for yourselves. Do you think you will live with him when you know that you are guilty? Do you think that you could be happy living with that holy Person, when you are full of pain and anguish because you know that you have always broken his laws?
%Mormon 9:4
\bverse \iffalse Behold, I say unto you that ye would be more miserable to dwell with a holy and just God, under a consciousness of your filthiness before him, than ye would to dwell with the damned souls in hell. \fi
You would be happier living with wicked people in hell than you would be living with a holy and fair God, knowing that you are not clean in God's eyes.
%Mormon 9:5
\bverse \iffalse For behold, when ye shall be brought to see your nakedness before God, and also the glory of God, and the holiness of Jesus Christ, it will kindle a flame of unquenchable fire upon you. \fi
When you see God's glory and the holiness of Jesus Christ, and you realize that you cannot hide anything from God, it will feel like a burning fire that cannot be put out.
%Mormon 9:6
\bverse \iffalse O then ye unbelieving, turn ye unto the Lord; cry mightily unto the Father in the name of Jesus, that perhaps ye may be found spotless, pure, fair, and white, having been cleansed by the blood of the Lamb, at that great and last day. \fi
You people who do not believe, turn to the Lord! Cry out to the Father in the name of Jesus so that on the last day you can be pure and white, and free of stains, because you have been cleaned by the blood of the Lamb.
%Mormon 9:7
\bverse \iffalse And again I speak unto you who deny the revelations of God, and say that they are done away, that there are no revelations, nor prophecies, nor gifts, nor healing, nor speaking with tongues, and the interpretation of tongues; \fi
I also speak to you people who deny God's revelations and say that there are no more revelations or prophecies, or spiritual gifts, or healing, or the gifts of speaking in other languages and interpreting other languages.
%Mormon 9:8
\bverse \iffalse Behold I say unto you, he that denieth these things knoweth not the gospel of Christ; yea, he has not read the scriptures; if so, he does not understand them. \fi
People who deny these things do not know the gospel of Christ. They have not read the scriptures; if they have, they do not understand them.
%Mormon 9:9
\bverse \iffalse For do we not read that God is the same yesterday, today, and forever, and in him there is no variableness neither shadow of changing? \fi
Don't the scriptures say that God is the same yesterday, and today, and forever? Don't they say that God does not change at all?
%Mormon 9:10
\bverse \iffalse And now if ye have imagined up unto yourselves a god who doth vary, and in whom there is shadow of changing, then have ye imagined up unto yourselves a god who is not God of miracles. \fi
And if you imagine that God changes at all, you are imagining a god who is not a God who does miracles.
%Mormon 9:11
\bverse \iffalse But behold, I will show unto you a God of miracles, even the God of Abraham, and the God of Isaac, and the God of Jacob; and it is that same God who created the heavens and the earth, and all things that in them are. \fi
But I want to show you a God who does miracles---the God of Sarah and Abraham, and Rebekah and Isaac, and Rachel and Jacob---the same God who created the heavens and the earth, and everything in them.
%Mormon 9:12
\bverse \iffalse Behold he created Adam, and by Adam came the fall of man.  And because of the fall of man came Jesus Christ, even the Father and the Son; and because of Jesus Christ came the redemption of man. \fi
God created Eve and Adam, and humankind fell because of Eve and Adam. And because humankind fell, Jesus Christ came. He is the Father and the Son. And because of Jesus Christ, humankind is redeemed.
%Mormon 9:13
\bverse \iffalse And because of the redemption of man, which came by Jesus Christ, they are brought back into the presence of the Lord; yea, this is wherein all men are redeemed, because the death of Christ bringeth to pass the resurrection, which bringeth to pass a redemption from an endless sleep, from which sleep all men shall be awakened by the power of God when the trump shall sound; and they shall come forth, both small and great, and all shall stand before his bar, being redeemed and loosed from this eternal band of death, which death is a temporal death. \fi
Humankind is redeemed through Jesus Christ. Because of this, they will be brought back to stand in front of the Lord. This is what it means that all humankind is redeemed: Everyone will be resurrected because of Christ's death. They will be resurrected from a neverending sleep. All people will wake up from this sleep by the power of God when the trumpet blows. They will all come forward, both the small and the great, and everyone will stand before God's judgment bar. They will be redeemed and set free from the eternal chains of the death of their bodies.
%Mormon 9:14
\bverse \iffalse And then cometh the judgment of the Holy One upon them; and then cometh the time that he that is filthy shall be filthy still; and he that is righteous shall be righteous still; he that is happy shall be happy still; and he that is unhappy shall be unhappy still. \fi
Then the Holy One will judge them. That is the time when a person who is filthy will still be filthy, and a person who is righteous will still be righteous. The person who is happy will remain happy, and the person who is unhappy will remain unhappy.
%Mormon 9:15
\bverse \iffalse And now, O all ye that have imagined up unto yourselves a god who can do no miracles, I would ask of you, have all these things passed, of which I have spoken? Has the end come yet? Behold I say unto you, Nay; and God has not ceased to be a God of miracles. \fi
All of you people who have imagined a god who does not do miracles, I want to ask you: Have the things that I am talking about happened yet? Has the end come yet? No, it has not. And God has not stopped being a God of miracles.
%Mormon 9:16
\bverse \iffalse Behold, are not the things that God hath wrought marvelous in our eyes? Yea, and who can comprehend the marvelous works of God? \fi
Don't you think that the things God has done are wonderful? Who can understand the wonderful things that God does?
%Mormon 9:17
\bverse \iffalse Who shall say that it was not a miracle that by his word the heaven and the earth should be; and by the power of his word man was created of the dust of the earth; and by the power of his word have miracles been wrought? \fi
Is there anyone who thinks it is not a miracle that heaven and earth were created by God's word? By the power of God's word humankind was created from the dust of the earth. By the power of God's word miracles have been done.
%Mormon 9:18
\bverse \iffalse And who shall say that Jesus Christ did not many mighty miracles? And there were many mighty miracles wrought by the hands of the apostles. \fi
Who would say that Jesus Christ did not do many mighty miracles? And the apostles also did many mighty miracles.
%Mormon 9:19
\bverse \iffalse And if there were miracles wrought then, why has God ceased to be a God of miracles and yet be an unchangeable Being?  And behold, I say unto you he changeth not; if so he would cease to be God; and he ceaseth not to be God, and is a God of miracles. \fi
And if they did miracles then, why would God stop being a God who does miracles? God does not change. If God did change, God would no longer be God. But God does not change, and God does miracles. 
%Mormon 9:20
\bverse \iffalse And the reason why he ceaseth to do miracles among the children of men is because that they dwindle in unbelief, and depart from the right way, and know not the God in whom they should trust. \fi
The reason that God stops doing miracles among humankind is that they stop believing, and they start to do wicked things. They do not know God, the one they should trust.
%Mormon 9:21
\bverse \iffalse Behold, I say unto you that whoso believeth in Christ, doubting nothing, whatsoever he shall ask the Father in the name of Christ it shall be granted him; and this promise is unto all, even unto the ends of the earth. \fi
I say to you, that anyone who believes in Christ and does not doubt, will be given whatever they ask the Father for in the name of Christ. God makes this promise with everyone, no matter who they are or where they live.
%Mormon 9:22
\bverse \iffalse For behold, thus said Jesus Christ, the Son of God, unto his disciples who should tarry, yea, and also to all his disciples, in the hearing of the multitude: Go ye into all the world, and preach the gospel to every creature; \fi
You remember that this is what Jesus Christ, the Son of God, said to his disciples that would continue to live on this earth, and to all his disciples, in front of the crowds: Go into all the world, and preach the gospel to every creature.
%Mormon 9:23
\bverse \iffalse And he that believeth and is baptized shall be saved, but he that believeth not shall be damned; \fi
The person that believes and is baptized will be saved. But the person who does not believe will be punished.
%Mormon 9:24
\bverse \iffalse And these signs shall follow them that believe--in my name shall they cast out devils; they shall speak with new tongues; they shall take up serpents; and if they drink any deadly thing it shall not hurt them; they shall lay hands on the sick and they shall recover; \fi
There will be miracles among the believers: they will force demons out of people; they will speak new languages; they will pick up snakes and not be harmed; if they drink something poisonous it will not hurt them; they will put their hands on people who are sick, and the sick will get better;
%Mormon 9:25
\bverse \iffalse And whosoever shall believe in my name, doubting nothing, unto him will I confirm all my words, even unto the ends of the earth. \fi
I will make these things come true for anyone who believes in my name and does not doubt, no matter where they are in the world.
%Mormon 9:26
\bverse \iffalse And now, behold, who can stand against the works of the Lord? Who can deny his sayings? Who will rise up against the almighty power of the Lord? Who will despise the works of the Lord? Who will despise the children of Christ? Behold, all ye who are despisers of the works of the Lord, for ye shall wonder and perish. \fi
Who is strong enough to fight against the Lord's works? Who dares to reject what the Lord says? Who will fight against the mighty power of the Lord? Beware, you people who hate and reject the works of the Lord. Beware, you people who hate the children of Christ. You will be amazed, and you will die.
%Mormon 9:27
\bverse \iffalse O then despise not, and wonder not, but hearken unto the words of the Lord, and ask the Father in the name of Jesus for what things soever ye shall stand in need. Doubt not, but be believing, and begin as in times of old, and come unto the Lord with all your heart, and work out your own salvation with fear and trembling before him. \fi
Do not hate or be amazed. Listen to the words of the Lord, and ask the Father in the name of Jesus for the things that you need. Do not doubt! Believe, and do as people did in the past: Come to the Lord with all your heart, and keep working to let Christ save you; do it with fear, and be serious.
%Mormon 9:28
\bverse \iffalse Be wise in the days of your probation; strip yourselves of all uncleanness; ask not, that ye may consume it on your lusts, but ask with a firmness unshaken, that ye will yield to no temptation, but that ye will serve the true and living God. \fi
Be wise while you are alive. Get rid of anything that makes you unclean. Do not ask God for things that you lust after, but ask with unshakable faith for help to not give in to temptation, and to serve the true and living God.
%Mormon 9:29
\bverse \iffalse See that ye are not baptized unworthily; see that ye partake not of the sacrament of Christ unworthily; but see that ye do all things in worthiness, and do it in the name of Jesus Christ, the Son of the living God; and if ye do this, and endure to the end, ye will in nowise be cast out. \fi
Do not be baptized if you are not worthy yet. Do not take the sacrament of Christ if you are not worthy. Do these things when you are worthy, and do them in the name of Jesus Christ, the Son of the living God. If you do this, and you are faithful until the end, you will not be thrown out.
%Mormon 9:30
\bverse \iffalse Behold, I speak unto you as though I spake from the dead; for I know that ye shall have my words. \fi
I am speaking to you like someone speaking from the dead, because I know that someday you will see the words that I am writing.
%Mormon 9:31
\bverse \iffalse Condemn me not because of mine imperfection, neither my father, because of his imperfection, neither them who have written before him; but rather give thanks unto God that he hath made manifest unto you our imperfections, that ye may learn to be more wise than we have been. \fi
Do not judge me because I am not perfect. Do not judge my father, or the people who wrote before him, for their imperfections. Instead, give thanks to God that God has shown you our imperfections, so that you can be wiser than we have been.
%Mormon 9:32
\bverse \iffalse And now, behold, we have written this record according to our knowledge, in the characters which are called among us the reformed Egyptian, being handed down and altered by us, according to our manner of speech. \fi
In this record we have written the things that we know. We have written in the language that we call reformed Egyptian. It has been handed down and changed by us, according to the way that we speak.
%Mormon 9:33
\bverse \iffalse And if our plates had been sufficiently large we should have written in Hebrew; but the Hebrew hath been altered by us also; and if we could have written in Hebrew, behold, ye would have had no imperfection in our record. \fi
If our plates were large enough we would write in Hebrew. But we have also changed the Hebrew language. And if we could have written in Hebrew, there would not be any imperfections in our record.
%Mormon 9:34
\bverse \iffalse But the Lord knoweth the things which we have written, and also that none other people knoweth our language; and because that none other people knoweth our language, therefore he hath prepared means for the interpretation thereof. \fi
But the Lord knows the things that we have written, and he knows that no one else knows our language. And because no one else knows our language, he has prepared a way for this record to be interpreted.
%Mormon 9:35
\bverse \iffalse And these things are written that we may rid our garments of the blood of our brethren, who have dwindled in unbelief. \fi
We write these things so that our clothes will not be stained by the blood of our sisters and brothers, who have fallen and who do not believe.
%Mormon 9:36
\bverse \iffalse And behold, these things which we have desired concerning our brethren, yea, even their restoration to the knowledge of Christ, are according to the prayers of all the saints who have dwelt in the land. \fi
We have a hope that our sisters and brothers will one day know about Christ again. And this hope is what all the saints who have lived in this land have prayed for.
%Mormon 9:37
\bverse \iffalse And may the Lord Jesus Christ grant that their prayers may be answered according to their faith; and may God the Father remember the covenant which he hath made with the house of Israel; and may he bless them forever, through faith on the name of Jesus Christ. Amen. \fi
May the Lord Jesus Christ answer their prayers according to their faith. May God the Father remember the covenant that he made with the people of Israel. May he bless them forever, through faith in the name of Jesus Christ. Amen.
